\chapter{The Potential Auger}
\label{ch:augers}
\indx{augers}
%\setlength{\columnsep}{\defcolwidth}
%\begin{multicols*}{2}
\begin{quote}
\begin{center}
It is easier to say\\
we have not lived up to our potential\\
than it is to do something about it!
\end{center}
\hspace*{\fill} – Daine Faushen, 7253 SF
\end{quote}
Life had not been easy for Father Faushen. The end of all things, and his own death, looked him in the face. He stood before a raging wall of strange flame, ready to cast an end to his life.
"YOU ARE IN DANGER."came the metallic voice. "CONTACT WILL TERMINATE YOUR EXISTENCE."

"You don't understand" shouted Daine, "That is what I seek!"

Taking a deep breath, he prepared to give himself to the flames. As he did, his thoughts turned to the events which had led him to this ill ending.

He had been ordained into the Followers of Light, only nine months before. It had been the night which he first felt the power of Apollo flow through his body. He had just reached the age of twenty-two years, and had been looking forward to a long life of service, satisfaction and peace after his ordination. But it was not to be. That very night Cershin the Sky watc her had warned the peoples of Rigel of the approaching object in the night skies. Jaern, he had called it, meaning Prodigal Son in ancient Trejon, for it looked similar to the other roving lights which skywatchers had been charting since the dawn of recorded history.

When Cershin announced the intruder was to pass close to Torandor, the people of the city began to attend services en mass, praying to Apollo to protected them from the baleful influence of the object in the sky, which, by then, wa s even then visible during the day. Daine spent every waking hour assisting services and helping reassure the faithful that Apollo would allow them to come to no harm.

When the object began to show a discernible disk, the windstorms and quakes began, throwing the city dwellers into a panic. Daine still believed that their salvation was at hand, his faith in Apollo stood strong. Even when the Chaos Lord found and activated the giant Kaaren of Destruction buried below the marketplace, Daine still held close to his beliefs.

Then the great darkness came. Clouds of volcanic dust spewed forth from the cracking and tortured land and darkened the skies. People rioted, breaking into businesses and homes, pillaging, setting fire to large parts of Rigel. Daine watched as a group of angry followers confronted the senior Priest of Light and burned him alive. Fleeing from the temple, he beseeched Apollo to save his people from this disaster.

His prayers went unanswered. His despair deepened when he, along with all the other priests, felt the death of
Apollo. His magic shorn from him, his faith shattered, Daine fled into the city, blindly seeking an end to his pain. Rushing through the streets, he was caught when a large building collaped. Stones from its wall struck him and he lost consciousness.

"Father?"a voice called to Daine. "Are you aware?"

Hurt, and frightened, Daine opened his eyes. Slowly coming to focus he saw the face of one of his parishioners, looking at him with concern. Beyond him were a number of other people, sitting on the floor, with scattered sacks and packs of belongings cluster about them. The room they were in had smooth grey walls, and was lit from strange harsh lanterns set behind panels in the ceiling.

"Yes... I am. Where are we?"he asked.

The man, whom he recognized as Haaro, a food merchant, responded by telling him that he was aboard the Kaaren of Destruction, suspended in space, and that their own world of Torandor was no more. Haaro had been on his way to board the ship when he saw Father Faushen unconscious in the street.  Carrying him and paying passage to the profiteering guildsmen, they descended below the city and boarded the buried Kaaren. Haaro watched as the vehicle ascended to the sky, and shared the horror as his world broke to pieces and scattered. Daine had been unconscious for over two days.

"So, our god is dead, our world is lost, and we have no home but this metal prison?" Daine asked.

"Do not despair father, for the Lord of Chaos and his wife have a plain. They claim we can journey to the interloper and carve a place to live there. They have directed those steering this great vessel to lead it to Jaern, where we will start anew,"responded Haaro. "Even now, we approach."

As to demonstrate this, the room gentlely shook. Haaro directed Daine's attention to a window through which the others were now watching. Not far away, the gleaming white ice of Jaern's surface approached. The land was moving by quickly, and becoming visibly closer.

Despair gripped Daine. Unable to bear up under the many disasters, his sanity crumbled as he lost a grip on the reality about him.

"Why do this?"he shouted. "It will come to naught. We will spoil this wasteland, just as we destroyed our home!"

"Calm down father," Haaro said as he reached for Daine's convulsing body. "This will do you no good."

The others in the room looked on, either too numb from their losses or too far gone to care to interfere. Haaro tried to calm the father, but his efforts only seemed to upset Daine even more. Daine swung his fist and connected with Haaro' s jaw, sending him sprawling across the room. Others started getting up to try to put an end to the fight, but Daine rushed out screaming, blindly seeking escape.

An endless maze of corridors soon made retreat impossible. Daine was hopelessly lost and confused, but with the end of everything he cherished before him, he did not seek to find those he left behind. A strange throbbing noise came from some of the passages ahead of him, and he headed in that direction. Coming into a large chamber, he heard a cold metallic voice speak.

"THIS AREA IS RESTRICTED. HIGH POTENTIALS CAN CAUSE DAMAGE TO LIFE FORMS. TURN BACK."

Well, danger did not deter him at this point. Ending his suffering was all he sought. Not thinking, he ceased his thoughts of the past, and leaped into the wall of flames.

* * * *

"Was this the afterlife?" Daine thought. He existed. That he was sure of, but there was nothing to see, nothing to
hear or feel. Was he to spend eternity here, in the dark, alone?

"Hello?"

"Is there anyone here?"

"Help me Apollo! hear my plea me," he cried.

But there was no answer from his deceased diety. This must be what it is like to not exist he thought to himself. In anguish, he shouted.
"Anyone answer me!"
"WHAT ANSWER IS REQUESTED?"came an unexpected query from the darkness beyond.

"Who is that?" Daine asked.

"I AM THE KAAREN. I AM THAT IN WHICH YOU AND YOUR COMPANIONS TRAVEL. I WAS CREATED TO SERVE. PLEASE STATE YOUR REQUEST."

Daine thought for quite a while, and then spoke what he felt.

"Please, kill me. All that I knew is gone, and all that lays ahead is more pain, anger and death. If we survive on the new planet, we will make all the same mistakes, the same prejudices, hatred and jealousy that has plagued mankind from its start. Let me meet my end now and be done with it."

"YOUR STATEMENT IS NOT CORRECT."

"Not correct? What do you mean?"

"WHAT I SAID. YOU NEED NOT BE SUBJECT TO THE SAME WEAKNESSES WHICH HAVE DRIVEN YOUR KIND."

"But surely mankind has not changed. We will make the same disastrous mess of this new world as we did the last."

"THAT IS BECAUSE YOU DO NOT THINK PROPERLT. ATTEND AS I SHOW YOU HOW TO PROPERLY UTILIZE ALL OF YOUR ABILITIES."

\section{History}

And indeed, that is what the mind of the Kaaren of Destruction did for the poor disillusioned priest. It showed him that there was more than one way to think and learn about life. It presented him with a path which would bring his dormant possibilities to fore, to let him experience and fulfill all his potentials. When Father Faushen came to, he laid behind his old existance, and gathered other broken and scared people about him and spread this new way of looking at the world to others. When the Kaaren landed, he and his companions left the company of the others, and sought an isolated place to put into practice this new way.

When the oceans melted, they were left isolated on a island far from the other survivors. These new thinkers worked together to learn how to live up to their potential, and over the next few generations achieved great steps in this direction. Their new mind disciplines taught them not to thing of problems and issues as isolated strings of events, purposes and directions, but to treat each item they learned independently. By letting all kinds of diverse, seemingly unrelated items effect each other, they discovered that they could learn new disciplines, philosophies and methods many times faster than their humanoid brothers.

\section{Motivation}

"Be all that You can Be"

The new thinkers, or Augers, as they became to known, have decided that their first priority is for each individual do everything needed to realize their full potential. To reach these goals, they learn to Auge, a new way of reasoning and correlating seeming diverse items to produce new and unexpected results.

\section{Their Society}

Since the needs of each individual differs, and since the search to realize potential is a very personal one, they don' t tend to work in groups to accomplish things. They are scattered about Jaern, each finding the place where they have the highest probability of fulfilling their goals. They do have a large tendency to avoid places where people are oppressed or under obligations which have nothing to do with their potentials.

Without the benefit of the Kaaren, which taught Daine to Auge in almost a single instant, understanding this new way of thinking is a life long process. Most must start in early childhood, and spend ten to twenty years learning and practicing these methods before striking out on their own. For this to succeed, they potential Augers place learning to auge as their first priority until they feel they no longer need their teachers.

\section{Requirements}

Any who wish are accepted at these remote Pensihalos, or halls of thinking. Augers whom have already discovered their best potential is teaching staff these places, passing on their learning skills. A potential auger must accept the authority of these sages, for only by completely giving themselves into the sages'  jurisdiction, can they experience the events needed to acquiring the ability to auge.

\section{Induction}

Most whom whill auge enter the Pensihalo at a young age, usually between five and ten. Augers have long ago learned that the younger the inductee, the easier it is for them to unlearn standard thinking and learn to auge. To prove their commitment, the potential auger presents themselves to the sages, giving them all their worldly possessions, and, in effect, becoming property of the Pensihalo.

\section{Augeing to Auge}

Little is know of the actual method of learning to use this new kind of thinking. The few that have been allowed to observe come out with confused rambling explanations of physical exercises, random yelling and talk by both students and teachers, and a lot of confusion. It is theorized by the
priests of the archive that augeing differs from methods by treating each fact and action separate from its context, not in sequence as is the priests'  normal method. Patterns of methods attributed to each fact and action guide how an auger will manipulate that fact. The auger can appear to be thinking and acting on a number of separate activities at one time, because they don' t attach sequence to these items.

In any case, the potential auger comes to understanding with this ability over time, and assist in passing the ability on to the younger inductees.

\section{Defining Potentials}

The sages help the inductees learn what are their strengths and weaknesses. Exercises containing elements of almost every conceivable profession and activity are simulated and practiced within their training. By looking at their current performance the inductee discovers their eventual potentials.

\section{Release from the Pensihalo}

When the sages and the student feel that the student has reached a level of augeing which will continue to grow without the help of the sage, the student is encouraged to leave, and seek to live up to their strengths and potentials. Some at this stage decide to stay behind to teach others. Occasionally even the sages will decide it is time to leave and seek outsiders to enlighten.

\section{Traditions}

With the vast variety of each auger' s potentials, there is little which is common among them. Outsiders however look upon augers both with awe and a little suspicion. When an auger comes into a town to take employment, the others with the same profession know they are in for competition.

Most augers are recognizable from their more archaic accent of Paroli which is kept alive within the  Pensihalo, and from their habit to become oblivious to the outside world from a few seconds to hours at a time as they auge. Less perceptive augers will often treat outsiders as a lower breed and this has often caused outsiders to form an opinion of the augers as stuck up and distant. Some professional guilds and organizations have even gone so far as rejecting auger members to try to eliminate what they perceive as unfair competition.

\section{Augering Skills}

Of all the backgrounds, the augers have the largest pool of skills. These skills are at costs which are generally lower than the skills specific to other backgrounds. While anyone can purchase these skills, only the augers get them at this low price. While, at first glance, it might seem that all professionals and workers would choose this path, the twenty years of isolation, discipline, and self control make this attractive only to a special breed of people. That is why only a small fraction of the people in a profession are augers.

\section{Learning new Abilities}

The Auger has a definite advantage in learning new abilities. Aside from the large pool of skills, the auger can also learn a skill or a spell from someone else, and they can learn it up to two ranks lower than the teacher rather than four. If one auger is teaching a such an ability to another auger, they can learn it up to the same rank that that the teacher possesses it.

%\end{multicols*}