\chapter{Earth Magic}
\label{ch:elemental-earth-magic}

Magicians of the Earth Element study the earth and stones, seeking to understand their properties and the matrices that hold the world together. As a result of their studies they are typically methodical, well-organized, and persistent.

Earth Magic typifies the power of the element, and the magician’s ability to manipulate that power.

Magnetism spells allow the caster to manipulate the natural attractions and repulsions that exist in various material objects. Items composed primarily of iron, nickel, or cobalt that are affected by magnetism are called “ferromagnetic.” Most spells in this group only affect ferromagnetic items. However, this group also contains spells to endow non-ferromagnetic items with that property.

Hindrances are used to pos e obstacles in another’s path, through the blocking nature of earth and stone.

The Necromancy/Time spells form a very powerful group capable of affecting the dimesion of time through the timelessness of Earth and dealing certain death to return living matter to the clay from whence it came. Because of the great power that must be channeled to use these spells, they are more difficult to learn, and drain their user. The base costs for these spells is 600 EP.

\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{tabular}{@{} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth}}
\begin{enumerate}
	\item Earth Magic
	\begin{enumerate}
		\item Rockfist
		\item Liquify
		\item Open
		\item Wane
		\item Trench
		\item Sculpt
		\item Wax
		\item Masonry
		\item Stretch
		\item EarthForce
		\item Transparency
		\item Flatten
		\item Chasm
		\item Adamant
		\item Terovoki
		\item Crumble
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{1}
	\item Magnetism
	\begin{enumerate}
		\item Compass
		\item Magnetize
		\item Iron Defense
		\item Ironize
		\item Hither
		\item Deflect Missile
		\item Flux Shield
		\item Ironwood
		\item Repulse Metal
		\item Dervish
		\item Monopole
		\item Magic Buckler
		\item Ironflesh
		\item Magic Matrix
		\item Rebound
		\item Conversion
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{2}
	\item Hindrances
	\begin{enumerate}
		\item Bar
		\item Springe
		\item Lock Arcane
		\item False Door
		\item Jam
		\item Caltrops
		\item Punchdoor
		\item Deadfall
		\item False Trails
		\item Camodoor
		\item Trigger
		\item Thorny Maze
		\item Blockade
		\item Mankiller
		\item No Trace
		\item Seal
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{3}
	\item Necromancy/Time
	\begin{enumerate}
		\item Shiver
		\item Mist
		\item Cold Touch
		\item The Voice
		\item Root
		\item Pause Time
		\item Soul Catcher
		\item Earthbone
		\item Suspend Time
		\item Soul Home
		\item Scythe
		\item Halt Time
		\item Soul Hold
		\item Obliterate
		\item Retreat Time
		\item Soul Banish
	\end{enumerate}
\end{enumerate} \\
\end{tabular}
\pagebreak
\section{Earth Magic}

\spellentry{Rockfist}{1}{1 Rounds}{None}{Touch}{Instantaneous}{Single target}{1d6 + 1/Finesse punch}{Empowers the caster to punch their target as if hitting them with a rock. The spell energy is only expended on a successfull attack roll.}

\spellentry{Liquify}{2}{2 Rounds}{None}{25 + 5/Finesse Feet}{15 + 5/Finesse Rounds}{10 + 2/Finesse Feet Radius}{Make mud}{Turns a horizontal surface of earth, sand or clay into a viscous mud. The mud cuts movement rates to 1/4 normal, and reduces all defense values by 2. When the spell expires, the mud turns into a soft loam from which ensnared creatures can easily escape.}

\spellentry{Open}{3}{2 Rounds}{None}{Touch}{1d6 + 1/Finesse Minutes}{0.5 + 0.25/Finesse Feet Radius}{Open passage}{Create a small cylindrical opening in earth or stone. This may only be a niche if the stone is very thick.}

\spellentry{Wane}{4}{1 Rounds}{3d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{1d6 + 1/Finesse Rounds}{1 creature}{Lower STR 1d6 + 2/Finesse}{Reduces the strength of living creatures as well as lowering damage from physical attacks. Any successful "to Strike" attacks by the affected target are automatically lowered by 1d6 + 2/Finesse DP for the duration of the spell.}

\spellentry{Trench}{5}{1 Rounds}{None}{LOS 10 + 3/Finesse Feet}{Instantaneous}{5 + 2/Finesse Feet Radius}{Dig hole in earth}{Dig a hole downward into earth, but not stone, of the indicated volume. The dimensions are as the caster desires, within the given area of effect. The trench is created by displacing the earth and is permanent until manually filled.}

\spellentry{Sculpt}{6}{1 Minutes}{None}{Touch}{Special}{5 + 1/Finesse lb stone}{Work stone}{By using this spell and the appropriate tools, the caster can quickly create finished objects (small statues, weapons, etc.) from stone within the casting time. Items normally made of stone last indefinitely, but things better made of metal (swords, etc.) will last only one day. The workmanship cannot exceed what the caster is capable of without the use of this spell.}

\spellentry{Wax}{7}{1 Rounds}{4d6 vs PWR negates}{Touch}{1d8 + 1/Finesse Rounds}{1 creature}{Raise STR 1d6 + 1/Finesse}{Increases the strength of its target as well as raising damage from physical attacks. Any successful "to Strike" attacks by the affected target are automatically raised by 1d6 + 2/Finesse DP for the duration of the spell.}

\spellentry{Masonry}{8}{3 Rounds}{None}{30 + 5/Finesse Feet}{30 + 10/Finesse Rounds}{10 + 5/Finesse Feet Radius}{Erect a wall}{Magically construct a stone wall. Stones rise up through the earth (or from whatever source seems reasonable) and are fitted and mortared together. The wall can be shaped as desired within the limits of the Area. At the end of the spell the mortar disintegrates and the whole construct can be easily knocked down.}

\spellentry{Stretch}{9}{2 Rounds}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{1d8 + 2/Finesse Rounds}{1 humanoid}{Double normal reach}{The target can stretch their arms as if they were made of rubber. The top of a wall may be reachable, opponents may be attacked while preventing them from striking, etc.}

\spellentry{Earth Force}{10}{1 Rounds}{None}{40 + 10/Finesse Feet}{1d8 + 2/Finesse Rounds}{6 + 2/Finesse Feet Radius}{0 to 2 Gs}{Creatures and objects in the area of effect can be made weightless so they rise from the ground if they push with their feet, or made up to 2 times heavier than normal, impeding all their actions by the factor of the G force.}

\spellentry{Transparency}{11}{1 Rounds}{None}{Touch}{10 + 3/Finesse Minutes}{1 + 0.5/Finesse Feet Radius}{See thru stone}{Create a cylindrical viewing portal through earth or stone. This portal allows viewing in only one direction of the caster's choice.}

\spellentry{Flatten}{12}{2 Rounds}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{8 + 2/Finesse Minutes}{1 being}{Make two-dimensional}{Make one target and their possessions as thin as a piece of paper. The target is still capable of movement, and can slip under doors or through cracks. Since a thin target is harder to hit, they receives a +6 defense bonus.}

\spellentry{Chasm}{13}{2 Rounds}{None}{150 + 50/Finesse Feet}{Instantaneous}{50 x 5 x 10 (LxWxD)}{Open crack in earth}{The earth trembles and produces a crack in the area of effect. Structures spanning the crack may be damaged or destroyed, depending upon the sturdiness and depth of their foundations (GM's discretion). Each finesse extends one dimension by 5 feet.}

\spellentry{Adamant}{14}{1 Minutes}{None}{Touch}{Instantaneous}{100 lb object}{Make unbreakable}{Endows normally breakable non-living objects (glass, pottery, etc.) with the durability of the hardest metal. It works on normal and magical objects alike.}

\spellentry{Terovoki}{15}{5 Rounds}{None}{100 + 20/Finesse}{4 + 1/Finesse Minutes}{Special}{Summon earth elemental}{Summon 1d6+1 earth elementals to any point within the caster's range. A source of earth must be available and must be part of the planet. The elementals will serve the caster, even in combat. They will not turn on them, but will make note of who summoned them. There is a 1d10 chance on each successive summoning that the elementials will attempt to take the caster with them at the end of the spell.}

\spellentry{Crumble}{16}{4 Rounds}{None}{100 + 30/Finesse Feet}{Instantaneous}{30 + 10/Finesse Feet Radius}{Turn stone to dust}{The indicated area of natural stone or stone work crumbles into dust.}

\section{Magnetism}

\spellentry{Compass}{1}{1 Round}{None}{Touch}{1d4 + 1/Finesse H}{1 foot radius object}{Find magnetic north}{Make any ferromagnetic item into a compass. The item points to magnetic north (not necessarily true north) for the duration of the spell.}

\spellentry{Magnetize}{2}{1 Round}{None}{Touch}{2 + 1/Finesse Minutes}{10 + 2/Finesse lb object}{Magnetizes object}{Magnetizes one ferromagnetic object.}

\spellentry{Iron Defense}{3}{1 Round}{None}{Caster}{2d6 + 2/Finesse Round}{Caster}{Inc. DV 2 + 1/Finesse}{Increase the defensive value of the caster's choice (combat, missile, or grapple) by temporarily adding a minute amount of the earth's magnetic power.}

\spellentry{Ironize}{4}{1 Round}{Special}{LOS 20 + 5/Finesse Feet}{14 + 4/Finesse Round}{100 lb object}{See text}{Make any metallic or stone object ferromagnetic. This spell is most useful as a preparatory step to other magnetism spells. Magical items get a 4d6 RC versus their PWR to avoid the effect.}

\spellentry{Hither}{5}{1 Round}{None}{LOS 20 + 5/Finesse Feet}{6 + 2/Finesse Round}{60 lb item}{Pull item to caster}{Pull a ferromagnetic item towards the caster. The object must be free standing and normally liftable by the caster. The caster gestures at the item and pulls and then the object moves at a rate of 4 feet per round in a straight line.}

\spellentry{Deflect Missile}{6}{1 Round}{None}{Caster}{2d6 + 2/Finesse Round}{Caster}{See text}{Cause any missile with a metal head to swerve harmlessly to one side of the magician (caster's choice). It is possible that someone else might be struck by the missile.}

\spellentry{Flux Shield}{7}{1 Round}{None}{Caster}{3d6 + 3/Finesse Minutes}{Caster}{See text}{Protect the caster from any natural or magically induced magnetic field, including the spells in this group.}

\spellentry{Ironwood}{8}{1 Round}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{14 + 4/Finesse Round}{5 + 2/Finesse Feet radius}{Magnetize plants}{Temporarily make plant material ferromagnetic. The area of effect is a 5 + 2/Finesse Feet radius for normal non-sentient plants, no RC. Only a single sentient plant lifeform can be affected, and it gets an RC of 4d6 vs PWR to resist.}

\spellentry{Repulse Metal}{9}{1 Round}{None}{LOS 20 + 5/Finesse Feet}{2d6 + 2/Finesse Minutes}{300 lb objects}{Push items away}{Push ferromagnetic items away from the caster at a rate of 4 feet per round.}

\spellentry{Dervish}{10}{2 Round}{5d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{2d6 + 2/Finesse Minutes}{200 + 50/Finesse lb object}{Spin 60 RPM}{Cause a ferromagnetic item to spin in an alternating magnetic field. If sentient, the target cannot perform any actions while enspelled.}

\spellentry{Monopole}{11}{1 Round}{None}{Touch}{2 + 1/Finesse Minutes}{100 + 20/Finesse lb object}{Anti-magnetism}{Creates a magnetic monopole from a single ferromagnetic object. Such monopoles are anti-magnetic in that they repel rather than attract other ferromagnetic items.}

\spellentry{Magic Buckler}{12}{1 Round}{None}{Caster}{2d6 + 2/Finesse Round}{Caster}{Deflects attacks}{Deflect all attacks from ferromagnetic weapons, regardless of the direction of the attack. Magical weapons get a 4d6 RC versus PWR to penetrate the effect.}

\spellentry{Ironflesh}{13}{1 Round}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{14 + 4/Finesse Minutes}{1 creature}{Magnetize flesh}{Turn flesh ferromagnetic, allowing it to be affected by either normal magnetism or magnetism spells.}

\spellentry{Magic Matrix}{14}{3 Round}{None}{Caster}{2d6 + 2/Finesse Round}{6 + 2/Finesse Feet radius}{Prevents magic}{Protects all those within it from the effect of all magician core spells, plus one type of elemental magic. The area is centered on the caster. The caster states the element to be excluded at the time of casting. Matrix spells can be "nested" to provide protection from more than one type of elemential magic.}

\spellentry{Rebound}{15}{1 Round}{4d6 vs STR negates}{Caster}{15 + 5/Finesse Round}{Caster}{Rebounds attacks}{Turn attacks upon the caster from ferromagnetic weapons back onto attackers. The counter attack is made using the attackers' melee modifiers and defense values. Once Rebound in place, the caster is free to perform other actions.}

\spellentry{Conversion}{16}{2 Round}{None}{Caster}{15 + 5/Finesse Round}{4 + 1/Finesse Feet radius}{Convert magic}{Convert any elemental spell directed at the caster into another elemental spell as if it were cast by the caster. Essentially, this allows the caster to cast a spell using someone else's power. The converted spell must use fewer elemental units than the original spell and must be a spell the caster knows.}


\section{Hindrances}

\spellentry{Bar}{1}{1 Round}{None}{Touch}{1d4 + 1/Finesse H}{1 door}{Bar a door}{Apply a magical force across the inside of an existing door just as if it were barred with a stout length of wood. STR checks to open the door are made with one additional die.}

\spellentry{Springe}{2}{4 Round}{4d6 vs PER finds}{Touch}{Until tripped}{5' radius}{Lay a snare}{Quickly lay a snare, the type that uses a bent sapling or other springy object to tighten a rope around a victim's feet and lift them into the air. The snare can be found and avoided by making the RC.}

\spellentry{Lock Arcane}{3}{1 Round}{None}{Touch}{1d4 + 1/Finesse H}{1 door}{Lock a door}{Lock any kind of door. Lockpicking attempts are of no use, and the only way to open it is to Revoke the spell or apply 2x as much force as would normally be necessary.}

\spellentry{False Door}{4}{4 Round}{4d6 vs PER finds}{Touch}{Permanent}{1 door}{Hide real door}{Create a door that leads nowhere and cannot be opened. Upon examination and a successful RC versus PER, it becomes apparent that this is someone Feets idea of a joke. False Door can be used in combination with Camodoor or other spells.}

\spellentry{Jam}{5}{2 Round}{None}{Touch}{1d4 + 1/Finesse H}{1 door}{Wedge door}{Wedges a door such that trying to open it simply jams it tighter. The door can be opened by Revoking the spell or applying 4x the necessary force, destroying the door in the process.}

\spellentry{Caltrops}{6}{1 Round}{None}{20 + 5/Finesse Feet}{1d4 + 1/Finesse Minutes}{6 + 2/Finesse Feet radius}{Prickly footing}{Deploy hundreds of four-headed spikes, one of which always points up. Moving through the area must be done at 1/4 normal speed or suffer 4d4 points of damage per round.}

\spellentry{Punchdoor}{7}{2 Round}{None}{Touch}{2d6 + 2/Finesse Minutes}{1 door}{Hit with door}{Place a portal that delivers the same number of dice and type of damage as that used to try to open it.}

\spellentry{Deadfall}{8}{5 Round}{4d6 vs PER finds}{Touch}{Until tripped}{3' radius}{Create deadfall}{Identify a good trap area and speedily set up a deadfall using available materials. The deadfall can be noticed by making the indicated RC, and tripped without damage. Stumbling into a stone deadfall inflicts 4d6 points of damage; a wooden one 2d6.}

\spellentry{False Trails}{9}{4 Round}{4d6 vs PER negates}{Multitouch}{Instantaneous}{Targets}{2 + 1/Finesse trails}{Create fakse trails. One trail is real, of course, but can only be positively identified by making the RC. These trails no longer radiate magic after being created, and will age naturally.}

\spellentry{Camodoor}{10}{4 Round}{4d6 vs PER finds}{Touch}{1d4 + 1/Finesse H}{1 door}{Blend door}{Disguise a door to blend in perfectly perfectly with its surroundings. It can only be found by making the RC (one attempt per adventurer).}

\spellentry{Trigger}{11}{1 Minutes}{None}{Touch}{30 + 10/Finesse Minutes}{6 + 2/Finesse Feet radius}{Triggers spell}{Create a circular plane that serves as a threshold over a doorway or opening. While casting Trigger, the caster casts a second spell that is held in suspension with the Trigger. If the threshold is breached, the second spell is set off. The triggered spell is subject to its normal resistance checks}

\spellentry{Thorny Maze}{12}{1 Minutes}{4d6 vs WIL negates}{Touch}{1d6 + 1/Finesse H}{1 normal door}{Create a maze}{Anyone who touches a door protected with Thorny Maze and fails the RC will find themselves in the midst of a maze of dense thorn bushes. They must roll 10 or above on 2d6 to extricate themselves (check each round). For every failed attempt they take 1d6 points of damage. On a successful check the maze disappears.}

\spellentry{Blockade}{13}{4 Round}{5d6 vs PWR negates}{Touch}{1d8 + 1/Finesse H}{1 door}{1 + 1/Finesse item}{Touch a normal door and name 1+1/Finesse people, creatures, or items. They cannot pass through the door unless the RC is made.}

\spellentry{Mankiller}{14}{1 Minutes}{5d6 vs AGI avoids}{Touch}{Until tripped}{10 + 2/Finesse Feet radius}{Make deadly trap}{Create a trap that uses swinging swords, thrusting spears, etc. to deal death and destruction. The setup is up to the caster, using up to four such weapons. The trap is so elegant that it cannot be discovered until it's too late. The victim dies unless the RC is made.}

\spellentry{No Trace}{15}{4 Round}{5d6 vs PER finds}{Multitouch}{3d6 + 2/Finesse hours}{Targets}{Obliterate trail}{Travel without leaving a trace. Anyone making the RC can find the trail, but only one attempt per seeker is allowed.}

\spellentry{Seal}{16}{1 Minutes}{Special}{Touch}{Permanent}{6' radius}{See text}{Create an area that cannot be passed. In addition, if the caster finesses the effect, anyone who touches the door and fails a 5d6 vs PWR RC, suffers one of these effects which the caster designates on setting the seal: 1 Finesse: Paralysis, 2 Finesses: Petrifaction, 3 Finesses: Death, 4 Finesses: Obliteration} 

\section{Necromancy/Time}

\spellentry{Shiver}{1}{1 Round}{None}{LOS 80 + 20/Finesse Feet}{1 Round}{1 target}{Causes a cold shiver}{Send a chilling shiver up the spine of its target. The spell requires little hand motion and no speech, so no one can tell the caster used a spell.}

\spellentry{Mist}{2}{1 Round}{None}{100 + 25/Finesse Feet}{4 + 2/Finesse Round}{10' radius}{Creates cloud}{Create a cloud of dust and mist. This is especially good for impressive entrances and exits and will obscure the caster. Finessing the effect will allow the caster to shape or sequence the cloud in any way they choose.}

\spellentry{Cold Touch}{3}{1 Round}{3d6 vs WIL negates}{Touch}{4 Round}{Target}{Produces chill}{Produces a very deep chilling sensation, as if the target was being gripped by the hand of death. No actual damage is done, but if the target fails his RC, they can perform no action while the spell exists and they are gripped by the caster.}

\spellentry{The Voice}{4}{1 Round}{3d6 vs WIL negates}{Hearing}{4 + 2/Finesse Round}{Targets}{Changes voice}{Change the caster's voice into a smooth and gripping tone which is difficult to ignore. Those who hear The Voice and fail their RC can perform no actions while the caster is speaking. Targets get an RC attempt each round of the duration.}

\spellentry{Root}{5}{2 Round}{3d6 vs PWR negates}{Touch}{Instantaneous}{1 being}{Change to plant}{If the target fails their RC, their feet become growing roots, rooting them to the spot. Each succeeding week sees another 10\% of their body turned into a plant; after 10 weeks the victim is a plant. During the transformation they gains nourishment from their roots. If Root is Revoked, before they are totally changed to a plant form, the effects can be reversed. Once the victim is 100\% a plant, they remain so forever.}

\spellentry{Pause Time}{6}{1 Round}{4d6 vs AGI negates}{Touch}{1d12 + 3/Finesse Round}{Targeted creature}{Halt passage of time}{Surround the target by a blue field of energy which suspends the passage of time. The field and the object appear immaterial and can not be affected by anything other than its revocation. When the duration expires or the spell is Revoked, the affected creatures will not know anything has happened to them.}

\spellentry{Soul Catcher}{7}{3 Round}{4d6 vs WIL negates}{Touch}{Permanent}{1 being}{Stores essence}{The living target of this spell must have a soul and be touching a stone with a weight of at least 10 pounds. If they fails their RC the essence of their being is drawn into the stone, leaving their body as a living empty husk.}

\spellentry{Earthbone}{8}{3 Round}{4d6 vs PWR negates}{Touch}{Permanent}{1 being}{Special}{Any being the caster touches, who fails the RC, is immediately turned to stone. This does not include the target's clothing or equipment. Successfully Revocating this spell returns the target to normal, with no knowledge of the intervening time, but a memory of being turned to stone.}

\spellentry{Suspend Time}{9}{1 Round}{5d6 vs PWR negates}{Touch}{Permanent}{Targeted creature}{Suspend life}{The target is surrounded by a blue field of energy which suspends the passage of time. The field and the object appear immaterial and can not be affected by anything other than its revocation. The target's body doesn't die nor require sustenance or air. They resume living when the spell is revoked with no concept of the passage of time nor knowledge of events that transpired during their magical sleep.}

\spellentry{Soul Home}{10}{1 Minutes}{3d6 vs WIL negates}{Touch}{Permanent}{Targeted creature}{Implants soul}{This spell allows the caster to take a stored soul and implant it into a living animal or humanoid body which is missing it's soul.}

\spellentry{Scythe}{11}{2 Round}{3d6 vs AGI negates}{20 + 5/Finesse Feet}{1 + 1/Finesse Round}{1 creature}{Instant death}{The target must be touching earth or stone to be affected by this spell. At the end of the casting time, a bloody scythe bursts from the ground and sweeps towards the target. If they makes a successful RC, the blade misses and the target is unharmed. If the RC fails, the blade hits and the victim is cut in half, decapitated, etc (GM's discretion).}

\spellentry{Halt Time}{12}{1 Round}{None}{80 + 20/Finesse Feet}{1 + 1/Finesse days}{30 + 10/Finesse Feet radius}{Stops time}{All creatures and objects with 50\% or more of their mass within the radius are surrounded by a blue field of energy which suspends the passage of time. The field and the objects appear immaterial and can not be affected by anything other than its Revocation. No people or creatures within the area will be aware that the spell occurred or that any time has passed.}

\spellentry{Soul Hold}{13}{1 H}{None}{Caster}{6 months}{Caster}{Locks soul}{Lock the caster's soul within their body, preventing it from leaving even if the body dies. The soul animates the body, which continues to function, past death. The body will decompose normally, and they will eventually end up a skeleton. Each time this spell is cast, the caster's HEA is reduced by one rank. When HEA reaches 0, the spell expires, the body collapses, and the soul leaves.}

\spellentry{Obliterate}{14}{3 Round}{5d6 vs PWR negates}{Touch}{Instantaneous}{1 creature}{Utter annihilation}{The target must be touching earth or stone. If they fail their RC they are immediately, irretrievably turned to dust.}

\spellentry{Retreat Time}{15}{1 Round}{None}{None}{Instantaneous}{Special}{Reverses time}{Cause the flow of time around the caster to reverse for one round. Thrusting himself into the past, at the conclusion of this spell they will find themselves at the place and time of the beginning of the previous round. No one else will have any knowledge that this has occurred, and the caster will only know because they will remember the events of the aborted round, and will have the spell energy for this spell expended.}

\spellentry{Soul Banish}{16}{1 Round}{5d6 vs WIL negates}{LOS 100 + 30/Finesse Feet}{Instantaneous}{Single creature}{Eliminates soul}{Totally and utterly obliterate the sould of the target, leaving their body an empty husk. No divine intervention is allowed the target.}
