\chapter{Elemental Core Magic}
\label{ch:elemental-core-magic}
Non-divine magic is derived from the natural forces present in the environment: \indy{Air}, \indy{Water}, \indy{Earth}, and \indy{Fire}. Magicians must learn how to control and shape these forces to create magical manifestations. Each type of force, or element, is controlled in very different and incompatible ways. When a magician begins to learn their art, they must choose an element with which to work. Once declared, the magician may only buy spell groups from that element, or the element it dominates (see \chpage{acquiring-spells-other-elements}.

There are seven spell groups common to all magicians, and four unique spell groups related to each element. The base cost of the \indy{Necromancy/Time}, \indy{Mind Twisters}, \indy{Charms} and the \indy{Dimensions groups} is \tcdefine{600 EP}. The base cost of the rest of the spell groups is \tcdefine{300 EP}.

\section{Magical Core Spells}
All magicians have access to certain “core” spells. These spells are created by the power of the caster’s element, but manifest themselves similarly regardless of the element. Thus, some form of the caster’s element must be present so that they may draw elemental power from it.
%\setlength{\columnsep}{\defcolwidth}\begin{multicols*}{2}
\indy{Defer} spells are the means by which a magician manufactures magical items and creates permanent or delayed spell effects.

\indy{Common Magic} is a group that makes a magician’s everyday and professional life easier.

\indy{Movement} spells enable the magician to move quickly and efficiently from place to place.

\indy{Revocation} spells are used to cancel another magician's magic. It has no effect on divine magic.

\indy{Tongues \& Scripts} provides the means to communicate by spoken, written, or cyphered word. All elemental mages understand and use the same basic concepts and ideas when casting, teaching and learning about their magics. These concepts transcend any normal human tongue, being the same for all elemental casters regardless of race, sex, or creed. This group of spells includes the ability to create, interpret and use
this elemental cypher.

\indy{Shadow Magic} enables the magician to manipulate shadow for useful and wondrous effects. They involve the use of both light and darkness. Light is used to delineate darkness, and thus these spells cease to function in the presence of extreme
light (greater than daylight) or absolute darkness. The shadows created have no physical substance, and cannot be harmed by any physical or magical attacks. They cannot, on the other hand, have any effect upon things in the physical world except as explicitly stated within each spell.
%\end{multicols*}

\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{tabular}{@{} p{0.33\linewidth} p{0.33\linewidth} p{0.33\linewidth}}
\begin{enumerate}
	\item Defer
	\begin{enumerate}
		\item Defer 1
		\item Defer 2
		\item Defer 3
		\item Defer 4
		\item Defer 5
		\item Defer 6
		\item Defer 7
		\item Defer 8
		\item Defer 9
		\item Defer 10
		\item Defer 11
		\item Defer 12
		\item Defer 13
		\item Defer 14
		\item Defer 15
		\item Defer 16
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\item Revocation
	\begin{enumerate}
		\item Revocation 1
		\item Revocation 2
		\item Revocation 3
		\item Revocation 4
		\item Revocation 5
		\item Revocation 6
		\item Revocation 7
		\item Revocation 8
		\item Revocation 9
		\item Revocation 10
		\item Revocation 11
		\item Revocation 12
		\item Revocation 13
		\item Revocation 14
		\item Revocation 15
		\item Revocation 16
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\item Common Magic
	\begin{enumerate}
		\item Lamp
		\item Change Temperature
		\item Renew Object
		\item Glue
		\item Mending
		\item Fetch
		\item Find
		\item Weapon Charm
		\item Reassemble
		\item Know Element
		\item Armor Charm
		\item Fasten
		\item Shelter
		\item Servants
		\item Know Magic
		\item Sunder
	\end{enumerate}
\end{enumerate} \\
\begin{enumerate}
	\item Movement
	\begin{enumerate}
		\item Tireless Walking
		\item Cushion
		\item Stickum
		\item Catspaw
		\item Quicken
		\item Leap
		\item Tireless Running
		\item Slow
		\item Grasshopper
		\item Speed
		\item Mire
		\item Puppet
		\item Hasten
		\item Lightfoot
		\item Snatch
		\item Emulate
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\item Tongues \& Scripts
	\begin{enumerate}
		\item Lexesoterica
		\item Message
		\item Cipher
		\item Name Tongue
		\item Encrypt
		\item Quiet
		\item Speed Read
		\item Fluency
		\item Scriptknow
		\item Lockjaw
		\item Interpolate
		\item Scribe
		\item Babble
		\item Need
		\item Translator
		\item Mindspeak
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\item Shadow Magic
	\begin{enumerate}
		\item Shadow Spawn
		\item Shadow Carry
		\item Contrast
		\item Shadow Object
		\item Shadow Servant
		\item Shadow Hide
		\item Shadow Horse
		\item Light Bane
		\item Shadow See
		\item Multishadow
		\item Shadow Leap
		\item Shadow Mount
		\item Shadow Life
		\item Shadow Target
		\item Shadow Guardian
		\item Shadow ‘Port
	\end{enumerate}
\end{enumerate}
\end{tabular}

\subsection{Defer}
\subsection{Revocation}
\subsection{Common Magic}
\spellentry{Lamp}{1}{1 Round}{None}{Touch}{\tcdieroll{1d6} + 1/Finesse Hours}{1 non-living object}{Illuminate}{Causes any non-living object to shed light equivalent to a normal torch.}

\spellentry{Change Temperature}{2}{1 Round}{None}{10 + 5/Finesse Feet}{\tcdieroll{1d8} + 2/Finesse Hours}{15 Foot Radius}{Set air temperature}{Sets and maintains any desired air temperature between \measure{0 and 100 degrees} Fahrenheit (-18 to 38 degrees Celsius) within an enclosed and moderately well insulated space without need for normal heating or cooling devices.}

\spellentry{Renew Object}{3}{1 Round}{None}{Touch}{Instantaneous}{30 + 15/Finesse pound object}{Restore appearance}{Restores one non-living object, within the given weight, to its new/clean appearance. Clothing, paintings, books, etc. are likely targets.}

\spellentry{Glue}{4}{1 Round}{None}{Touch}{Special}{1 surface/2 objects}{Glue together}{Glues two surfaces of any two items together. The items can not be separated by any physical means while this spell is in effect. Normally it last \measure{4 minutes}. One finesse allows it to last for \measure{1 hour}, two for \measure{1 day}, three for \measure{1 week} and four for \measure{1 month}. Needless to say, the caster’s finger is not stuck to the items while they target this spell.}

\spellentry{Mending}{5}{4 Rounds}{None}{Touch}{Instantaneous}{50 pound object}{Fix it}{Restores any broken, non-living item to its original condition if the object weighs no more than \measure{50 pounds}, the caster has handled the object before it was broken, and all the pieces are at hand.}

\spellentry{Fetch}{6}{1 Round}{None}{LOS 200 + 50/Finesse Feet}{Instantaneous}{20 + 5/Finesse pound object}{Retrieve item}{Causes any non-living, free standing item in the caster’s line of sight to be immediately brought to them.}

\spellentry{Find}{7}{1 Hour}{None}{caster}{Instantaneous}{300 + 150/Finesse met Radius}{Locate material}{The caster concentrates on a particular material and casts this spell. If that material is within the area of the spell, an image of the material will appear in the general direction of its location. This spell locates the material to the nearest met and is fairly useless at close range. The caster cannot use this spell to locate a particular object.}

\spellentry{Weapon Charm}{8}{1 Round}{None}{Touch}{4 + 1/Finesse Minutes}{1 weapon}{Enchant weapon}{Causes any weapon to function as if once enchanted for the duration of the spell. Using a weapon thus enchanted gives its bearer a \result{+1} on any roll “to \indy{strike}” or “to \indy{hit}” (\CM and \MM) and \result{+1 DP} on any
damage roll.}

\spellentry{Reassemble}{9}{1 Round}{None}{Touch}{Instantaneous}{100 pound object}{Make whole again}{Restores any broken, non-living object to its original condition. The caster must have the majority of the original pieces, but need not know the object’s original shape.}

\spellentry{Know Element}{10}{1 Minutes}{None}{Touch}{Instantaneous}{10 Foot radius object}{Reveal composition}{Reveals the elemental makeup of any object to the caster. Unfinnessed, the spell reveals the dominant element only; fully finessed, it reveals all elements used in the object’s makeup and manufacture.}

\spellentry{Armor Charm}{11}{1 Round}{None}{Touch}{4 + 1/Finesse Minutes}{1 suit of armor}{Enchant armor}{Causes any armor to function as if once enchanted for the duration of the spell. Using armor thus enchanted gives its wearer a \result{+1} on their \CDV and \MDV.}

\spellentry{Fasten}{12}{1 Round}{None}{20 + 5/Finesse Feet}{Instantaneous}{10 + 2/Finesse Foot Radius}{Work 1 + 1/Finesse fastenings}{Fastens or unfastens ropes, buckles, buttons, latches, etc. at a distance. By finessing, up to five fastenings can be worked in the area of effect.}

\spellentry{Shelter}{13}{1 Hour}{None}{40 Feet}{Instantaneous}{20 + 10/Finesse Foot Radius}{Create building}{The caster visualizes a structure within the given size limits as they begin to cast this spell. The magic then takes any needed trees within \measure{120 feet} of the target and processes them into lumber. The lumber, and any other building supplies left at the site by the caster, are used to construct a building as near to the visualization as the raw materials allow.}

\spellentry{Servants}{14}{1 Hour}{None}{0}{1 week}{200 + 100/Finesse Foot Radius}{1 + 1/Finesse servants}{The magician can automate certain aspects of their surroundings by creating invisible servants to handle such mundane chores as cooking, cleaning, laundering, trash removal, etc. Each servant has a specific task, so if more than five are required the spell must be cast again. The servants must stay within the spell’s area of effect.}

\spellentry{Know Magic}{15}{1 Hour}{None}{Touch}{Instantaneous}{1 magical object}{2 + 1/Finesse spells}{Reveals which spells were used to create a magical item. Up to six spells from the caster’s background can be discovered. Spells outside the caster’s background will not be revealed unless the magician first casts Know Element to discover their presence.}

\spellentry{Sunder}{16}{1 Round}{None}{Touch}{Instantaneous}{1000 pound object}{Break item}{Breaks any non-magical, non-living item, regardless of its composition.}

\subsection{Movement}
\spellentry{Tireless Walking}{1}{1 Round}{None}{Caster}{12 + 3/Finesse Hours}{Caster}{Walk without fatigue}{Increases the endurance of the caster’s body while walking (i.e., moving at half movement rate or less). While this spell is in effect, the caster will not tire or weaken.}

\spellentry{Cushion}{2}{1 Round}{None}{Caster}{2 + 1/Finesse Minutes}{Caster}{Non-damaging fall}{While under the protection of Cushion, the caster may fall from any height and suffer no impact damage. The speed of descent is unaffected by this spell; the impact occurs as if the laws of momentum have been temporarily suspended. This impact must occur before the spell duration expires, otherwise damage is calculated normally.}

\spellentry{Stickum}{3}{1 Round}{None}{Caster}{2 + 1/Finesse Minutes}{Caster}{Climbing aid}{Enables the caster to climb vertical surfaces, or hang from horizontal surfaces, with the surefootedness of a fly or spider. While sticking to a surface the caster can move at up to \measure{1/4} their normal land movement rate.}

\spellentry{Catspaw}{4}{1 Round}{None}{Caster}{6 + 2/Finesse Minutes}{Caster}{Silent movement}{Empowers the caster to move across potentially noisy areas (dry leaves, loose gravel, etc.) without a sound. Note that this spell will not silence speech, rustling clothing, etc.}

\spellentry{Quicken}{5}{1 Round}{None}{Caster}{8 + 2/Finesse Rounds}{Caster}{Increase speed}{Increase the caster's movement rate by \measure{10 feet per round} and allow them one additional attack each round for the duration of the spell. This does not affect spell casting in any way. Additional attacks are not cumulative with any other skill or spell.}

\spellentry{Leap}{6}{1 Round}{None}{Caster}{instantaneous}{Caster}{40 + 10/Finesse Foot jump}{Leap multiplies the force of the caster’s legs to propel them in a jump. The total length of the leap is up to \measure{40 feet} + \result{10 feet per finesse}, and may be any combination of vertical and horizontal distance that does not exceed the total.}

\spellentry{Tireless Running}{7}{1 Round}{None}{Caster}{12 + 3/Finesse Hours}{Caster}{Run without fatigue}{Tireless Running allows the caster to run (move at full movement rate) for the duration of the spell without fatigue.}

\spellentry{Slow}{8}{2 Rounds}{\tcdieroll{4d6} vs \PWR negates}{20 + 5/Finesse Feet}{8 + 2/Finesse Rounds}{10 + 5/Finesse Foot Radius}{Slow beings}{Slow reduces the movement and attack rates of all within the area of effect by \result{1/2}.}

\spellentry{Grasshopper}{9}{2 Rounds}{\tcdieroll{4d6} vs \WIL negates}{Multitouch}{4 + 2/Finesse Rounds}{1 + 1/Finesse living creatures}{Leap 40 + 10/Finesse Feet}{Grasshopper allows the caster to empower up to five beings they touch to perform one jump (as in \indy{Leap}) each round of the duration.}

\spellentry{Speed}{10}{2 Rounds}{\tcdieroll{4d6} vs \WIL negates}{Multitouch}{8 + 2/Finesse Rounds}{Living creatures}{Speed movement}{Speed increases the targets’ movement by \measure{20 feet per round} and allows them \result{2 additional attacks per round} for the duration of the spell. As many targets can be affected as the caster can touch within the casting time. Spell casting is \ul{not} affected in any way. Additional attacks are \ul{not} cumulative with any other skills or spells.}

\spellentry{Mire}{11}{3 Rounds}{\tcdieroll{4d6} vs \PWR negates}{20 + 5/Finesse feet}{8 + 2/Finesse Rounds}{10 + 5/Finesse feet Radius}{Slow 3 + 1/Finesse beings}{Mire reduces the movement and attack rates of all within the area of effect to \measure{1/3}.}

\spellentry{Puppet}{12}{1 Round}{\tcdieroll{4d6} vs \WIL negates}{0}{Special}{10 + 5/Finesse Foot Radius}{Endow with movement}{Puppet allows the caster to place all creatures in the area of effect in a trance-like state. They then casts a movement spell \tcdefine{Rank 7 or below} on themselves. All those under the effect of Puppet may now move as the caster moves. The effect ends when the chosen movement spell's duration ends.}

\spellentry{Hasten}{13}{4 Rounds}{\tcdieroll{4d6} vs \WIL negates}{Multitouch}{12 + 4/Finesse Rounds}{Living creatures}{Speed enhancement}{Hasten increases the targets’ movement by \measure{40 feet per round} and allows them an additional \result{4 attacks per round} for the duration of the spell. As many targets can be affected as the caster can touch within the casting time. Spell casting is \ul{not} affected in any way. Additional attacks are \ul{not} cumulative with any other skill or spell.}

\spellentry{Lightfoot}{14}{3 Rounds}{\tcdieroll{4d6} vs \WIL negates}{Touch}{60 + 10/Finesse Minutes}{1 Living creature}{Unnatural speed}{Endows target with the ability to run at a movement rate of \measure{500 feet per round}. This is approximately \mets{225}[hour].}

\spellentry{Snatch}{15}{1 Round}{\tcdieroll{4d6} vs \WIL negates}{Touch}{special}{1 Living being}{Move with objects}{Target can grab any moving object and move with it. The object must have enough momentum to lift the target, else nothing happens (GMs discretion). The target is also protected from any damage that might normally occur. \example{Thus an adventurer can snatch hold of a boulder} \example{flung by a siege machine,} \example{fly over a wall, and land unharmed.}}

\spellentry{Emulate}{16}{1 Round}{None}{Caster}{30 + 10/Finesse Minutes}{Caster}{Emulate movement}{Assume the movement ability of any creature the caster can see. The magician must still take precautions if movement is planned in a hostile environment (under water, through fire, etc.).}

\subsection{Tongues \& Scripts}
\spellentry{Lexesoterica}{1}{1 Round}{\tcdieroll{4d6} vs \WIL negates}{Touch}{\tcdieroll{1d4} + 1/Finesse Minutes}{1 Being}{Decipher magical text}{Enables target to read any script written in the special cipher of elemental magicians. Scrolls and other written material prepared with the \indy{Cipher} spell need this spell to be understood or used. A code or key phrase may need to be spoken while casting this spell if the Cipher was prepared in conjunction with the \indy{Encrypt} spell.}

\spellentry{Message}{2}{1 Round}{None}{LOS 500 + 250/Finesse Feet}{\tcdieroll{1d4} + 1/Finesse Minutes}{1 Creature}{Long range speech}{Communicate with another person or creature within line of sight by casting the spell, then whispering the message. Total privacy is not assured, as the whispered speech can be heard emanating from the target’s ear.}

\spellentry{Cipher}{3}{1 Round}{None}{Caster}{\tcdieroll{1d4} + 1/Finesse Minutes}{1 Scroll}{Magical writing}{Write something in elemental cipher. These ciphers can only be understood with the use of the Lexesoterica spell.}

\spellentry{Name Tongue}{4}{1 Round}{None}{LOS 5 Feet/hearing}{6 + 1/Finesse Rounds}{Special}{Identify language}{Identify an unknown language or script. Script identification has no duration. For spoken languages, the speaker must remain within hearing of the caster for the duration of the spell.}

\spellentry{Encrypt}{5}{2 Rounds}{None}{Touch}{Special}{1 Scroll}{Encrypt magical cipher}{Used in conjunction with \indy{Cipher}, this spell allows the caster to use a word or a short phrase to code the produced cipher. To use the cipher, the code must be spoken when Lexesoterica is cast.}

\spellentry{Quiet}{6}{1 Round}{None}{20 + 5/Finesse Feet}{30 + 10/Finesse Minutes}{10 + 5/Finesse Foot Radius}{Eliminate sound}{Creates an immobile area absolutely devoid of noise of any kind. It makes verbal communication impossible, and masks the natural
noises of anyone or anything in the affected area. Those within the area may still cast spells by mouthing the focus words, even if they cannot be heard.}

\spellentry{Speed Read}{7}{2 Minutes}{None}{Caster}{1 + 0.25/Finesse Hours}{Caster}{Read 4 + 1/Finesse Times Faster}{Read, with perfect comprehension, anything written in a language the caster already knows. The basic rate is \measure{1 page per minute}; the hastened rate depends on the finesses applied.}

\spellentry{Fluency}{8}{4 Rounds}{None}{Caster}{\tcdieroll{1d4} + 1/Finesse Minutes}{Caster}{Enhance communication}{Understand any spoken or written communication. The caster can also speak in this tongue if it is humanly possible. This encompasses all languages, and even animal grunts and squeaks if such could be considered normal communication for the animals involved.}

\spellentry{Scriptknow}{9}{1 Minute}{None}{Caster}{Time to Read}{1 Document}{Handwriting analysis}{Know the personality of the writer of a document at the time it was written. The veracity of the words is not revealed, though some clue may be provided. Note that analysis of a dictated message reveals the personality of the scribe, not the speaker.}

\spellentry{Lockjaw}{10}{1 Round}{\tcdieroll{4d6} vs \PWR negates}{LOS 20 + 5/Finesse Feet}{15 + 4/Finesse Rounds}{1 Being}{Prevent speech}{Prevents an individual from speaking. \example{Thus orders can be disrupted and spell casting prevented.}}

\spellentry{Interpolate}{11}{2 Minutes}{Special}{Touch}{\tcdieroll{1d8} + 1/Finesse Minutes}{1 Document}{Determine missing words}{Interpolate supplies five words for each illegible word written on any surface, one of which is correct. Each finesse reduces the number of suggestions by one. If Interpolate is combined with \indy{Fluency}, the writing can be in any language. Combined with \indy{Lexesoterica}, Interpolate can be used to complete fragmented magical text.}

\spellentry{Scribe}{12}{2 Rounds}{None}{Caster}{10 + 5/Finesse Minutes}{Caster}{Record speech}{Accurately record, and later recall or write down, any normal speech within the caster's hearing. It can be combined with \indy{Fluency} if the language is unknown to the caster.}

\spellentry{Babble}{13}{1 Round}{\tcdieroll{5d6} vs \PWR negates}{LOS 40 + 10/Finesse Feet}{15 + 4/Finesse Rounds}{Single Target}{Confuse speech}{Sabotages any attempt
to speak or form coherent communication. The target's voice comes out as gibberish, their writing as meaningless scrawls. Attempts to use hand signal turns into a hilarious stomping and waving gesture. Any attempt to cast magical spells fail since forming the focus for the cast is impossible.}

\spellentry{Need}{14}{1 Round}{None}{Touch}{Instantaneous}{1 Willing Creature}{Communicate need}{The caster casts the spell upon the target to attune the target to the magic. Thereafter, any time the caster casts Need and names the recipient, the caster’s desire is immediately communicated to the target, which must make every effort to comply.}

\spellentry{Translator}{15}{4 Rounds}{\tcdieroll{4d6} vs \WIL negates}{Multitouch}{20 + 10/Finesse Minutes}{Humanoids}{Enhance communication}{Allows others to also understand otherwise incomprehensible speech or writing. Willing targets need not make an RC.}

\spellentry{Mindspeak}{16}{4 Minutes}{\tcdieroll{4d6} vs \WIL negates}{Infinite}{4 + 2/Finesse Minutes}{1 Being}{Voiceless speech}{Open two-way mental communications with anyone the caster has actually seen, regardless of where they are, for the duration of the spell.}

\subsection{Shadow Magic}
\spellentry{Shadow Spawn}{1}{1 Round}{None}{LOS 40 + 10/Finesse Ft}{\tcdieroll{1d4} + 1/Finesse Minutes}{10 + 3/Finesse Foot Radius}{make shadows}{}

\spellentry{Shadow Carry}{2}{1 Round}{None}{Caster}{\tcdieroll{1d4} + 1/Finesse Minutes}{caster's shadow}{give shadow strength}{}

\spellentry{Contrast}{3}{1 Round}{None}{LOS 40 + 10/Finesse Feet}{\tcdieroll{1d4} + 1/Finesse Minutes}{10 + 3/Finesse' Radius}{reveal the hidden}{}

\spellentry{Shadow Object}{4}{2 Rounds}{None}{Touch}{\tcdieroll{1d4} + 1/Finesse Minutes}{20' radius shadow}{create objects}{}

\spellentry{Shadow Servant}{5}{2 Rounds}{None}{Caster}{2 + 1/Finesse Hours}{caster's shadow}{make a servant}{}

\spellentry{Shadow Hide}{6}{1 Round}{None}{Caster}{2 + 1/Finesse Hours}{caster's shadow}{see text}{}

\spellentry{Shadow Horse}{7}{4 Rounds}{None}{Touch}{8 + 2/Finesse Hours}{Caster}{create equine}{}

\spellentry{Light Bane}{8}{1 Round}{None}{20 + 10/Finesse Ft}{\tcdieroll{1d8} + 2/Finesse Minutes}{20 + 10/Finesse' Radius}{banish illumination}{}

\spellentry{Shadow See}{9}{2 Rounds}{None}{Caster}{\tcdieroll{1d4} + 1/Finesse Minutes}{caster's shadow}{use shadow to see}{}

\spellentry{Multishadow}{10}{1 Round}{None}{Caster}{\tcdieroll{1d4} + 1/Finesse Minutes}{caster's shadow}{2 + 1/Finesse shadows}{}

\spellentry{Shadow Leap}{11}{1 Round}{None}{Caster}{\tcdieroll{1d4} + 1/Finesse Rounds}{caster's shadow}{see text}{}

\spellentry{Shadow Mount}{12}{4 Rounds}{None}{LOS 10 Ft}{\tcdieroll{1d8} + 2/Finesse Hours}{single shadow mount}{create mount}{}

\spellentry{Shadow Life}{13}{2 Rounds}{None}{Caster}{\tcdieroll{1d4} + 1/Finesse Minutes}{caster's shadow}{vivify shadow}{}

\spellentry{Shadow Target}{14}{1 Round}{see text}{Touch}{Instantaneous}{1 shadow}{attack shadow}{}

\spellentry{Shadow Guardian}{15}{1 Minutes}{None}{LOS 10 Ft}{see text}{10' Radius}{create a trap}{}

\spellentry{Shadow 'Port}{16}{1 Round}{None}{Touch}{Instantaneous}{Caster}{go to source}{}
