\chapter{Elemental Core Magic}
\label{ch:elemental-core-magic}
Non-divine magic is derived from the natural forces present in the environment: Air, Water, Earth, and Fire. Magicians must learn how to control and shape these forces to create magical manifestations. Each type of force, or element, is controlled in very different and incompatible ways. When a magician begins to learn their art, they must choose an element with which to work. Once declared, the magician may only buy spell groups from that element, or the element it dominates (see \chpage{acquiring-spells-other-elements}.

There are seven spell groups common to all magicians, and four unique spell groups related to each element. The base cost of the Necromancy/Time, Mind Twisters, Charms and the Dimensions groups is 600 EP. The base cost of the rest of the spell groups is 300 EP.

All magicians have access to certain "core" spells. These spells are created by the power of the caster's element, but manifest themselves similarly regardless of the element. Thus, some form of the caster's element must be present so that they may draw elemental power from it.


Defer spells are the means by which a magician manufactures magical items and creates permanent or delayed spell effects.

Common Magic is a group that makes a magician's everyday and professional life easier.

Movement spells enable the magician to move quickly and efficiently from place to place.

Revocation spells are used to cancel another magician's magic. It has no effect on divine magic.

Tongues \& Scripts provides the means to communicate by spoken, written, or cyphered word. All elemental mages understand and use the same basic concepts and ideas when casting, teaching and learning about their magics. These concepts transcend any normal human tongue, being the same for all elemental casters regardless of race, sex, or creed. This group of spells includes the ability to create, interpret and use
this elemental cypher.

Shadow Magic enables the magician to manipulate shadow for useful and wondrous effects. They involve the use of both light and darkness. Light is used to delineate darkness, and thus these spells cease to function in the presence of extreme
light (greater than daylight) or absolute darkness. The shadows created have no physical substance, and cannot be harmed by any physical or magical attacks. They cannot, on the other hand, have any effect upon things in the physical world except as explicitly stated within each spell.

\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{tabular}{@{} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth}}
\begin{enumerate}
	\item Defer
	\begin{enumerate}
		\item Defer 1
		\item Defer 2
		\item Defer 3
		\item Defer 4
		\item Defer 5
		\item Defer 6
		\item Defer 7
		\item Defer 8
		\item Defer 9
		\item Defer 10
		\item Defer 11
		\item Defer 12
		\item Defer 13
		\item Defer 14
		\item Defer 15
		\item Defer 16
		\item Defer 17
		\item Defer 18
		\item Defer 19
		\item Defer 20
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{1}
	\item Revocation
	\begin{enumerate}
		\item Revocation 1
		\item Revocation 2
		\item Revocation 3
		\item Revocation 4
		\item Revocation 5
		\item Revocation 6
		\item Revocation 7
		\item Revocation 8
		\item Revocation 9
		\item Revocation 10
		\item Revocation 11
		\item Revocation 12
		\item Revocation 13
		\item Revocation 14
		\item Revocation 15
		\item Revocation 16
		\item Revocation 17
		\item Revocation 18
		\item Revocation 19
		\item Revocation 20
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{2}
	\item Common Magic
	\begin{enumerate}
		\item Lamp
		\item Change Temperature
		\item Renew Object
		\item Glue
		\item Mending
		\item Fetch
		\item Find
		\item Weapon Charm
		\item Reassemble
		\item Know Element
		\item Armor Charm
		\item Fasten
		\item Shelter
		\item Servants
		\item Know Magic
		\item Sunder
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{3}
	\item Movement
	\begin{enumerate}
		\item Tireless Walking
		\item Cushion
		\item Stickum
		\item Catspaw
		\item Quicken
		\item Leap
		\item Tireless Running
		\item Slow
		\item Grasshopper
		\item Speed
		\item Mire
		\item Puppet
		\item Hasten
		\item Lightfoot
		\item Snatch
		\item Emulate
	\end{enumerate}
\end{enumerate} \\
\begin{enumerate}
	\setcounter{enumi}{4}
	\item Tongues \& Scripts
	\begin{enumerate}
		\item Lexesoterica
		\item Message
		\item Cipher
		\item Name Tongue
		\item Encrypt
		\item Quiet
		\item Speed Read
		\item Fluency
		\item Scriptknow
		\item Lockjaw
		\item Interpolate
		\item Scribe
		\item Babble
		\item Need
		\item Translator
		\item Mindspeak
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{5}
	\item Shadow Magic
	\begin{enumerate}
		\item Shadow Spawn
		\item Shadow Carry
		\item Contrast
		\item Shadow Object
		\item Shadow Servant
		\item Shadow Hide
		\item Shadow Horse
		\item Light Bane
		\item Shadow See
		\item Multishadow
		\item Shadow Leap
		\item Shadow Mount
		\item Shadow Life
		\item Shadow Target
		\item Shadow Guardian
		\item Shadow 'Port
	\end{enumerate}
\end{enumerate}
\end{tabular}
\pagebreak

\section{Defer}

\spellentry{Defer}{1-20}{1 Minute}{None}{Touch}{2 Hours}{1 Spell}{Delay spell effects}{}

Casters use the Defer spell to delay the effects of another spell. The caster picks some non-living, non-magical object that they can hold in one hand and lift and cast the Defer spell on the object. This places a magical field about the object that can hold one other spell of a rank equal to or less then the rank of the Defer spell used. Immediately after this, the caster casts the other spell into the same item. The magical field created by the defer ensnares this second spell, hold within the field. 

As long as the defer spell lasts, the caster can touch the item and direct the ensnared spell to discharge, having the same effect as if cast normally. The spell’s release from the defer takes a single round, despite the casting time of the spell. If the defer spell expires, without triggering its contained spell, all magic energies involved dissipate harmlessly.

If the caster wants someone other than themselves to be able to release and direct the spell, they may finesse the original defer to include a trigger that others can use. The number of finesses determines what kind of trigger is used, as follows:

\begin{tabular}{@{} l l}
0 & Thought Activated for the original caster\\
1 & Mechanical\\
2 & Spoken Word or Phrase\\
3 & Thought Activated by anyone\\
4 & Activated by a condition
\end{tabular}

Spells cast with an unfinessed defer can only be thought activated by their original caster, when they are touching the item. One finesse allows the spell to be triggered by any moving part of the item. Two finesses allow the spell to be triggered by a keyword or phrase. Three finesses allow the spell to be thought activated by anyone holding the item. The fourth finesse allows the caster to state the single condition that will activate the item. This condition can only describe a state of being, and can only be based on the position, condition or existence of physical objects or energies within 10 Feet of the item.

The caster may choose to target the spell either at the time of casting the defer spell, or when it is triggered. If it is targeted when triggered, the person triggering the spell can direct it mentally, just as if they were the caster. If it is targeted when the defer is originally cast, the caster gives targeting information relative to the position and orientation of the object used for the defer spell.

Defer spells used on items with already existing defer spells cast by a different caster don’t work. Recasting a new defer spell on your deferred spell extends the duration of the original defer spell by an additional 2 hours.

The Elemental Defer spell can only be used to contain elemental spells.
\pagebreak
\section{Revocation}

\spellentry{Revocation}{1-20}{3 Rounds}{None}{LOS 80 + 20 / Finesse Feet}{Instantaneous}{1 Magical Effect}{Nullify elemential power}{}

\subsection{Revoking Spells}
\label{magic-elemental-core-revoking-spells}

The Revocation group is different from others in that the spell remains the same throughout, except that each successive spell is more powerful than the previous. It can affect both spells and items; potions and other single use items are considered as spells for revocation purposes.

The mechanics for determining if a revocation is successful are simple. The player rolls a number of d6 equal to the rank of the revocation their character is casting and totals them. The GM rolls a number of d6 equal to the rank of the effect the caster is attempting to revoke and totals them. The higher total wins, i.e., if the player's total is higher the targeted effect is negated; if the GM's total is higher the revocation fails.

\subsection{Affecting Items}

Each magical item has at least two components: a spell efffect and a Defer. These are specified in the item's description. To temporarily suspend the effect of an item (or draw charges from a charged item), the caster attempts to revoke the spell effect. Dice are rolled as described under \secref{magic-elemental-core-revoking-spells}. If the player’s total is higher the Revocation works and the difference between their and the GM's totals is the number of rounds the effect is negated, or the number of charges drawn.

To permanently cancel an item the caster must revoke the Defer spell. Dice are rolled, and if the player’s total is higher than the GM's the item is made permanently non-magical.

The Elemental Revocation spell can only be used to revoke elemental spells.
\pagebreak
\section{Common Magic}

\spellentry{Lamp}{1}{1 Round}{None}{Touch}{1d6 + 1/Finesse Hours}{1 non-living object}{Illuminate}{Causes any non-living object to shed light equivalent to a normal torch.}

\spellentry{Change Temperature}{2}{1 Round}{None}{10 + 5/Finesse Feet}{1d8 + 2/Finesse Hours}{15 Foot Radius}{Set air temperature}{Sets and maintains any desired air temperature between 0 and 100 degrees Fahrenheit (-18 to 38 degrees Celsius) within an enclosed and moderately well insulated space without need for normal heating or cooling devices.}

\spellentry{Renew Object}{3}{1 Round}{None}{Touch}{Instantaneous}{30 + 15/Finesse pound object}{Restore appearance}{Restores one non-living object, within the given weight, to its new/clean appearance. Clothing, paintings, books, etc. are likely targets.}

\spellentry{Glue}{4}{1 Round}{None}{Touch}{Special}{1 surface/2 objects}{Glue together}{Glues two surfaces of any two items together. The items can not be separated by any physical means while this spell is in effect. Normally it last 4 minutes. One finesse allows it to last for 1 hour, two for 1 day, three for 1 week and four for 1 month. Needless to say, the caster's finger is not stuck to the items while they target this spell.}

\spellentry{Mending}{5}{4 Rounds}{None}{Touch}{Instantaneous}{50 pound object}{Fix it}{Restores any broken, non-living item to its original condition if the object weighs no more than 50 pounds, the caster has handled the object before it was broken, and all the pieces are at hand.}

\spellentry{Fetch}{6}{1 Round}{None}{LOS 200 + 50/Finesse Feet}{Instantaneous}{20 + 5/Finesse pound object}{Retrieve item}{Causes any non-living, free standing item in the caster's line of sight to be immediately brought to them.}

\spellentry{Find}{7}{1 Hour}{None}{Caster}{Instantaneous}{300 + 150/Finesse met Radius}{Locate material}{The caster concentrates on a particular material and casts this spell. If that material is within the area of the spell, an image of the material will appear in the general direction of its location. This spell locates the material to the nearest met and is fairly useless at close range. The caster cannot use this spell to locate a particular object.}

\spellentry{Weapon Charm}{8}{1 Round}{None}{Touch}{4 + 1/Finesse Minutes}{1 weapon}{Enchant weapon}{Causes any weapon to function as if once enchanted for the duration of the spell. Using a weapon thus enchanted gives its bearer a +1 on any roll "to Strike" or "to Hit" (CM and MM) and +1 DP on any damage roll.}

\spellentry{Reassemble}{9}{1 Round}{None}{Touch}{Instantaneous}{100 pound object}{Make whole again}{Restores any broken, non-living object to its original condition. The caster must have the majority of the original pieces, but need not know the object's original shape.}

\spellentry{Know Element}{10}{1 Minutes}{None}{Touch}{Instantaneous}{10 Foot radius object}{Reveal composition}{Reveals the elemental makeup of any object to the caster. Unfinnessed, the spell reveals the dominant element only; fully finessed, it reveals all elements used in the object's makeup and manufacture.}

\spellentry{Armor Charm}{11}{1 Round}{None}{Touch}{4 + 1/Finesse Minutes}{1 suit of armor}{Enchant armor}{Causes any armor to function as if once enchanted for the duration of the spell. Using armor thus enchanted gives its wearer a +1 on their CDV and MDV.}

\spellentry{Fasten}{12}{1 Round}{None}{20 + 5/Finesse Feet}{Instantaneous}{10 + 2/Finesse Foot Radius}{Work 1 + 1/Finesse fastenings}{Fastens or unfastens ropes, buckles, buttons, latches, etc. at a distance. By finessing, up to five fastenings can be worked in the area of effect.}

\spellentry{Shelter}{13}{1 Hour}{None}{40 Feet}{Instantaneous}{20 + 10/Finesse Foot Radius}{Create building}{The caster visualizes a structure within the given size limits as they begin to cast this spell. The magic then takes any needed trees within 120 feet of the target and processes them into lumber. The lumber, and any other building supplies left at the site by the caster, are used to construct a building as near to the visualization as the raw materials allow.}

\spellentry{Servants}{14}{1 Hour}{None}{0}{1 week}{200 + 100/Finesse Foot Radius}{1 + 1/Finesse servants}{The magician can automate certain aspects of their surroundings by creating invisible servants to handle such mundane chores as cooking, cleaning, laundering, trash removal, etc. Each servant has a specific task, so if more than five are required the spell must be cast again. The servants must stay within the spell's area of effect.}

\spellentry{Know Magic}{15}{1 Hour}{None}{Touch}{Instantaneous}{1 magical object}{2 + 1/Finesse spells}{Reveals which spells were used to create a magical item. Up to six spells from the caster's background can be discovered. Spells outside the caster's background will not be revealed unless the magician first casts Know Element to discover their presence.}

\spellentry{Sunder}{16}{1 Round}{None}{Touch}{Instantaneous}{1000 pound object}{Break item}{Breaks any non-magical, non-living item, regardless of its composition.}

\section{Movement}

\spellentry{Tireless Walking}{1}{1 Round}{None}{Caster}{12 + 3/Finesse Hours}{Caster}{Walk without fatigue}{Increases the endurance of the caster's body while walking (i.e., moving at half movement rate or less). While this spell is in effect, the caster will not tire or weaken.}

\spellentry{Cushion}{2}{1 Round}{None}{Caster}{2 + 1/Finesse Minutes}{Caster}{Non-damaging fall}{Fall from any height and suffer no impact damage. The speed of descent is unaffected by this spell; the impact occurs as if the laws of momentum have been temporarily suspended. This impact must occur before the spell duration expires, otherwise damage is calculated normally.}

\spellentry{Stickum}{3}{1 Round}{None}{Caster}{2 + 1/Finesse Minutes}{Caster}{Climbing aid}{Enables the caster to climb vertical surfaces, or hang from horizontal surfaces, with the surefootedness of a fly or spider. While sticking to a surface the caster can move at up to 1/4 their normal land movement rate.}

\spellentry{Catspaw}{4}{1 Round}{None}{Caster}{6 + 2/Finesse Minutes}{Caster}{Silent movement}{Empowers the caster to move across potentially noisy areas (dry leaves, loose gravel, etc.) without a sound. Note that this spell will not silence speech, rustling clothing, etc.}

\spellentry{Quicken}{5}{1 Round}{None}{Caster}{8 + 2/Finesse Rounds}{Caster}{Increase speed}{Increase the caster's movement rate by 10 feet per round and allow them one additional attack each round for the duration of the spell. This does not affect spell casting in any way. Additional attacks are not cumulative with any other skill or spell.}

\spellentry{Leap}{6}{1 Round}{None}{Caster}{Instantaneous}{Caster}{40 + 10/Finesse Foot jump}{Allows the to jump much further. The total length of the leap is up to 40 feet + 10 feet per finesse, and may be any combination of vertical and horizontal distance that does not exceed the total.}

\spellentry{Tireless Running}{7}{1 Round}{None}{Caster}{12 + 3/Finesse Hours}{Caster}{Run without fatigue}{Allows the caster to run (move at full movement rate) for the duration of the spell without fatigue.}

\spellentry{Slow}{8}{2 Rounds}{4d6 vs PWR negates}{20 + 5/Finesse Feet}{8 + 2/Finesse Rounds}{10 + 5/Finesse Foot Radius}{Slow beings}{Reduces the movement and attack rates of all within the area of effect to 1/2x.}

\spellentry{Grasshopper}{9}{2 Rounds}{4d6 vs WIL negates}{Multitouch}{4 + 2/Finesse Rounds}{1 + 1/Finesse living creatures}{Leap 40 + 10/Finesse Feet}{Allows the caster to empower up to five beings they touch to perform one jump (as in Leap) each round of the duration.}

\spellentry{Speed}{10}{2 Rounds}{4d6 vs WIL negates}{Multitouch}{8 + 2/Finesse Rounds}{Living creatures}{Speed movement}{Increases the targets' movement by 20 feet per round and allows them 2 additional attacks per round for the duration of the spell. As many targets can be affected as the caster can touch within the casting time. Spell casting is not affected in any way. Additional attacks are not cumulative with any other skills or spells.}

\spellentry{Mire}{11}{3 Rounds}{4d6 vs PWR negates}{20 + 5/Finesse feet}{8 + 2/Finesse Rounds}{10 + 5/Finesse feet Radius}{Slow 3 + 1/Finesse beings}{Reduces the movement and attack rates of all within the area of effect to 1/3x.}

\spellentry{Puppet}{12}{1 Round}{4d6 vs WIL negates}{0}{Special}{10 + 5/Finesse Foot Radius}{Endow with movement}{All creatures in the area of effect enter into a trance-like state. The caster then casts a movement spell Rank 7 or below on themselves. All those under the effect of Puppet may now move as the caster moves. The effect ends when the chosen movement spell's duration ends.}

\spellentry{Hasten}{13}{4 Rounds}{4d6 vs WIL negates}{Multitouch}{12 + 4/Finesse Rounds}{Living creatures}{Speed enhancement}{Increases the targets' movement by 40 feet per round and allows them an additional 4 attacks per round for the duration of the spell. As many targets can be affected as the caster can touch within the casting time. Spell casting is not affected in any way. Additional attacks are not cumulative with any other skill or spell.}

\spellentry{Lightfoot}{14}{3 Rounds}{4d6 vs WIL negates}{Touch}{60 + 10/Finesse Minutes}{1 Living creature}{Unnatural speed}{Endows target with the ability to run at a movement rate of 500 feet per round. This is approximately \mets{225}[hour].}

\spellentry{Snatch}{15}{1 Round}{4d6 vs WIL negates}{Touch}{Special}{1 Living being}{Move with objects}{Target can grab any moving object and move with it. The object must have enough momentum to lift the target, else nothing happens (GMs discretion). The target is also protected from any damage that might normally occur. Thus an adventurer can snatch hold of a boulder flung by a siege machine, Fly over a wall, and land unharmed.}

\spellentry{Emulate}{16}{1 Round}{None}{Caster}{30 + 10/Finesse Minutes}{Caster}{Emulate movement}{Assume the movement ability of any creature the caster can see. The magician must still take precautions if movement is planned in a hostile environment (under water, through fire, etc).}

\section{Tongues \& Scripts}
\spellentry{Lexesoterica}{1}{1 Round}{4d6 vs WIL negates}{Touch}{1d4 + 1/Finesse Minutes}{1 Being}{Decipher magical text}{Enables target to read any script written in the special cipher of elemental magicians. Scrolls and other written material prepared with the Cipher spell need this spell to be understood or used. A code or key phrase may need to be spoken while casting this spell if the Cipher was prepared in conjunction with the Encrypt spell.}

\spellentry{Message}{2}{1 Round}{None}{LOS 500 + 250/Finesse Feet}{1d4 + 1/Finesse Minutes}{1 Creature}{Long range speech}{Communicate with another person or creature within line of sight by casting the spell, then whispering the message. Total privacy is not assured, as the whispered speech can be heard emanating from the target's ear.}

\spellentry{Cipher}{3}{1 Round}{None}{Caster}{1d4 + 1/Finesse Minutes}{1 Scroll}{Magical writing}{Write something in elemental cipher. These ciphers can only be understood with the use of the Lexesoterica spell.}

\spellentry{Name Tongue}{4}{1 Round}{None}{LOS 5 Feet/hearing}{6 + 1/Finesse Rounds}{Special}{Identify language}{Identify an unknown language or script. Script identification has no duration. For spoken languages, the speaker must remain within hearing of the caster for the duration of the spell.}

\spellentry{Encrypt}{5}{2 Rounds}{None}{Touch}{Special}{1 Scroll}{Encrypt magical cipher}{Used in conjunction with Cipher, this spell allows the caster to use a word or a short phrase to code the produced cipher. To use the cipher, the code must be spoken when Lexesoterica is cast.}

\spellentry{Quiet}{6}{1 Round}{None}{20 + 5/Finesse Feet}{30 + 10/Finesse Minutes}{10 + 5/Finesse Foot Radius}{Eliminate sound}{Creates an immobile area absolutely devoid of noise of any kind. It makes verbal communication impossible, and masks the natural
noises of anyone or anything in the affected area. Those within the area may still cast spells by mouthing the focus words, even if they cannot be heard.}

\spellentry{Speed Read}{7}{2 Minutes}{None}{Caster}{1 + 0.25/Finesse Hours}{Caster}{Read 4 + 1/Finesse Times Faster}{Read, with perfect comprehension, anything written in a language the caster already knows. The basic rate is 1 page per minute; the hastened rate depends on the finesses applied.}

\spellentry{Fluency}{8}{4 Rounds}{None}{Caster}{1d4 + 1/Finesse Minutes}{Caster}{Enhance communication}{Understand any spoken or written communication. The caster can also speak in this tongue if it is humanly possible. This encompasses all languages, and even animal grunts and squeaks if such could be considered normal communication for the animals involved.}

\spellentry{Scriptknow}{9}{1 Minute}{None}{Caster}{Time to Read}{1 Document}{Handwriting analysis}{Know the personality of the writer of a document at the time it was written. The veracity of the words is not revealed, though some clue may be provided. Note that analysis of a dictated message reveals the personality of the scribe, not the speaker.}

\spellentry{Lockjaw}{10}{1 Round}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{15 + 4/Finesse Rounds}{1 Being}{Prevent speech}{Prevents an individual from speaking. Thus orders can be disrupted and spell casting prevented.}

\spellentry{Interpolate}{11}{2 Minutes}{Special}{Touch}{1d8 + 1/Finesse Minutes}{1 Document}{Determine missing words}{Interpolate supplies five words for each illegible word written on any surface, one of which is correct. Each finesse reduces the number of suggestions by one. If Interpolate is combined with Fluency, the writing can be in any language. Combined with Lexesoterica, Interpolate can be used to complete fragmented magical text.}

\spellentry{Scribe}{12}{2 Rounds}{None}{Caster}{10 + 5/Finesse Minutes}{Caster}{Record speech}{Accurately record, and later recall or write down, any normal speech within the caster's hearing. It can be combined with Fluency if the language is unknown to the caster.}

\spellentry{Babble}{13}{1 Round}{5d6 vs PWR negates}{LOS 40 + 10/Finesse Feet}{15 + 4/Finesse Rounds}{Single Target}{Confuse speech}{Sabotages any attempt
to speak or form coherent communication. The target's voice comes out as gibberish, their writing as meaningless scrawls. Attempts to use hand signal turns into a hilarious stomping and waving gesture. Any attempt to cast magical spells fail since forming the focus for the cast is impossible.}

\spellentry{Need}{14}{1 Round}{None}{Touch}{Instantaneous}{1 Willing Creature}{Communicate need}{The caster casts the spell upon the target to attune the target to the magic. Thereafter, any time the caster casts Need and names the recipient, the caster's desire is immediately communicated to the target, which must make every effort to comply.}

\spellentry{Translator}{15}{4 Rounds}{4d6 vs WIL negates}{Multitouch}{20 + 10/Finesse Minutes}{Humanoids}{Enhance communication}{Allows others to also understand otherwise incomprehensible speech or writing. Willing targets need not make an RC.}

\spellentry{Mindspeak}{16}{4 Minutes}{4d6 vs WIL negates}{Infinite}{4 + 2/Finesse Minutes}{1 Being}{Voiceless speech}{Open two-way mental communications with anyone the caster has actually seen, regardless of where they are, for the duration of the spell.}

\section{Shadow Magic}

\spellentry{Shadow Spawn}{1}{1 Round}{None}{LOS 40 + 10/Finesse Ft}{1d4 + 1/Finesse Minutes}{10 + 3/Finesse Foot Radius}{Make shadows}{Make pockets of shadow as if the natural lighting direction had shifted to the caster's desire.}

\spellentry{Shadow Carry}{2}{1 Round}{None}{Caster}{1d4 + 1/Finesse Minutes}{Caster's shadow}{Give shadow strength}{Allows the caster's natural shadow to carry any items the caster could normally carry. These items are visible while being carried.}

\spellentry{Contrast}{3}{1 Round}{None}{LOS 40 + 10/Finesse Feet}{1d4 + 1/Finesse Minutes}{10 + 3/Finesse' Radius}{Reveal the hidden}{Deepens the tone of any shadow, thus causing anything hidden or concealed therein to stand out sharply against a pure black background.}

\spellentry{Shadow Object}{4}{2 Rounds}{None}{Touch}{1d4 + 1/Finesse Minutes}{20' radius shadow}{Create objects}{Make a non-living object from its shadow. The shadow must be motionless, and the caster must touch it to call the object's form from its shadow. The created object is the size of the shadow.}

\spellentry{Shadow Servant}{5}{2 Rounds}{None}{Caster}{2 + 1/Finesse Hours}{Caster's shadow}{Make a servant}{Create a servant from the caster's own shadow. The magician gives the shadowy butler a task, which it immediately carries out. This cycle can be repeated until the duration expires. The tasks must be those that the caster could actually perform and must not involve combat. The shadow servant has none of the magical abilities of the caster. If struck by a weapon, the shadow disperses.}

\spellentry{Shadow Hide}{6}{1 Round}{None}{Caster}{2 + 1/Finesse Hours}{Caster's shadow}{See text}{An enhancement of Shadow Carry. Items given to the
shadow are not visible while being carried.}

\spellentry{Shadow Horse}{7}{4 Rounds}{None}{Touch}{8 + 2/Finesse Hours}{Caster}{Create equine}{Call forth an equine creature from any natural shadow.
This can be a horse to be ridden, a mule for carrying heavy items, etc. The creature functions in all respects like its normal counterpart, except it does not need food or water.}

\spellentry{Light Bane}{8}{1 Round}{None}{20 + 10/Finesse Ft}{1d8 + 2/Finesse Minutes}{20 + 10/Finesse' Radius}{Banish illumination}{Cause any non-magical illumination device to go dark (though its fuel is still consumed).}

\spellentry{Shadow See}{9}{2 Rounds}{None}{Caster}{1d4 + 1/Finesse Minutes}{Caster's shadow}{Use shadow to see}{The caster is able to see whatever is in the line of sight of the head of their own shadow. Multishadow can be used in conjunction with this spell.}

\spellentry{Multishadow}{10}{1 Round}{None}{Caster}{1d4 + 1/Finesse Minutes}{Caster's shadow}{2 + 1/Finesse shadows}{Create duplicates of the caster's own shadow as if more than one directional light sources were striking them from different directions. This spell is most useful in conjunction with other shadow spells.}

\spellentry{Shadow Leap}{11}{1 Round}{None}{Caster}{1d4 + 1/Finesse Rounds}{Caster's shadow}{See text}{Instantaneously move to any point the caster's own shadow touches. One such move can be made each round for the duration of the spell.}

\spellentry{Shadow Mount}{12}{4 Rounds}{None}{LOS 10 Ft}{1d8 + 2/Finesse Hours}{Single shadow mount}{Create mount}{Create any type of creature for use as a mount by throwing its shadow on a surface and calling the creature forth. The shadow can be created with hand gestures, a paper cutout, etc.}

\spellentry{Shadow Life}{13}{2 Rounds}{None}{Caster}{1d4 + 1/Finesse Minutes}{Caster's shadow}{Vivify shadow}{Endows the caster's shadow with independent action. The shadow takes on the physical attributes of the caster, and for the duration of the spell can do any physical actions the caster could, including combat, but not spell casting. The shadow must remain within the caster's line of sight for the duration, or else the spell is cancelled.}

\spellentry{Shadow Target}{14}{1 Round}{See text}{Touch}{Instantaneous}{1 shadow}{Attack shadow}{Affect a target by performing an action or attack on the target's shadow. The effect or attack is adjudicated normally, with the effect targeted at the shadow but affecting the shadow's source.}

\spellentry{Shadow Guardian}{15}{1 Minutes}{None}{LOS 10 Ft}{See text}{10' Radius}{Create a trap}{The magician first makes the shadow of a creature on a surface, through hand manipulation, a paper cutout. etc. The caster then states (in 25 words or less) the condition that will trigger the spell. When triggered, the Shadow Guardian steps out of its shadow and attacks with the normal physical (not magical) abilities of the creature it mimics. The shadow disperses after it or the intruder is slain.}

\spellentry{Shadow 'Port}{16}{1 Round}{None}{Touch}{Instantaneous}{Caster}{Go to source}{When standing on a shadow, travel instantly to the exterior or interior of the shadow's source. The caster may choose where to appear in relation to the source, though they must know the layout of interior destinations to be completely successful. If they don't, they must roll their AGI or less on 4d6 to avoid a stumbling re-entry and loss of initiative for 1d4 rounds.}