\chapter{Divine Core Magic}
\label{ch:divine-core-magic}
\section{Organization}
Divine Core Magic
All priests, except those who worship T’or, have access to certain basic magics in addition to the spell groups specific to their sects.

Ceremonies are the magical component of the standard ceremonies performed by all sects. While these magics are shared, the actual form of the ceremony always differs from sect to sect. The spells have few noticeable effects, but they are very valuable from a roleplaying point of view.

Revocation spells are used to cancel another priest’s magic. They have no effect on elemental magic.

Blessings allow the priest to lay his hope for good fortune on deserving followers of his flock and favored friends. All blessings last no longer than 24 hours, or until the time they take effect. The priest can only cast blessings on others, not himself. When the priest casts a blessing on one of his own faith, it works automatically. When cast on one outside the faith, the GM will ask the target to succeed a check against the target’s PWR of a number of dice reflecting the GM’s opinion of the target’s piety and similarity of creed. Only one blessing can be on a target at one time. Blessings may not be used in conjunction with the Defer spell.

Fabrication spells allow the caster to create and manipulate various objects and holy writs.

Detections are used to inform the caster of different things around him. These spells also allow the priest to reach into a man’s mind and learn his inner self.

Influence spells are used by priests to show others the way of his faith, and guide them along the true path.

Bind spells are the means by which a priest manufactures magical items and creates permanent or delayed spell effects.

Defer spells are the means by which a priest manufactures divine items or delayed spell effects.

\section{Divine Core Spells}
\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{tabular}{@{} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth}}
\begin{enumerate}
	\item Ceremonies
	\begin{enumerate}
		\item Worship
		\item Consecrate Item
		\item Dedication
		\item Unification
		\item Last Rites
		\item Ordination
		\item Excommunication
		\item Atonement
		\item Mortify
		\item Sanctify
		\item Forbiddance
		\item Miracle
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{1}
	\item Revocation
	\begin{enumerate}
		\item Revocation 1
		\item Revocation 2
		\item Revocation 3
		\item Revocation 4
		\item Revocation 5
		\item Revocation 6
		\item Revocation 7
		\item Revocation 8
		\item Revocation 9
		\item Revocation 10
		\item Revocation 11
		\item Revocation 12
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{2}
	\item Defer
	\begin{enumerate}
		\item Defer 1
		\item Defer 2
		\item Defer 3
		\item Defer 4
		\item Defer 5
		\item Defer 6
		\item Defer 7
		\item Defer 8
		\item Defer 9
		\item Defer 10
		\item Defer 11
		\item Defer 12
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{3}
	\item Blessings
	\begin{enumerate}
		\item Divine Grace
		\item Deter Fate
		\item Abate Fatigue
		\item Optimize Onslaught
		\item Enhance Potential
		\item Defy Injury
		\item Augment Task
		\item Attract Fate
		\item Forestall Conflict
		\item Second Chance
		\item Abate Outcome
		\item Share Grace
	\end{enumerate}
\end{enumerate} \\
\begin{enumerate}
	\setcounter{enumi}{4}
	\item Fabrications
	\begin{enumerate}
		\item Create Water
		\item Speak The Word
		\item Create Bread
		\item Capture The Word
		\item Create Fish
		\item Create Meat
		\item Transfer The Word
		\item Create Holy Water
		\item Create Holy Symbol
		\item Create Fervor
		\item Produce Vestments
		\item Produce Truth
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{5}
	\item Detections
	\begin{enumerate}
		\item Detect Divinity
		\item Uncover Affection
		\item Detect Life
		\item Unveil Fear
		\item Detect Captivation
		\item Share Joy
		\item Discern Motivation
		\item Disclose Sin
		\item Reveal the Past
		\item Divulge Virtue
		\item Commune
		\item Manifest Destiny
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{6}
	\item Influences
	\begin{enumerate}
		\item Revoke Fear
		\item Instill Fear
		\item Share Vision
		\item Attention
		\item Paralyze
		\item Curse
		\item Revoke Curse
		\item Enthrall
		\item Devotion
		\item Disenchant
		\item Quest
		\item Divine Word
	\end{enumerate}
\end{enumerate}
\end{tabular}