\chapter{Nomadic Mysticism}
\label{ch:nomad}
\setlength{\columnsep}{\defcolwidth}\begin{multicols*}{2}
\begin{center}
\textbf{Nomadic Culture and External Interaction}
A Study of a Misunderstood People\\
Prepared by\\
Miche Aaper, FA EL CU/Archive at Karfelon\\
Torn 12, 9997 SF\\
IDX \#KARF-3025-003486\\
\end{center}
Seen only when they visit cities, towns, or villages to peddle their skills and entertainments, nomads are one of the most elusive and misunderstood cultures on Jaern today. Learning their beliefs and traditions seemed a hopeless task from the outside, so I undertook to research them by joining one of their \fixlater{rondos} to gain firsthand knowledge of their culture.

\textbf{Topic One – Wherein I gain Entrance to Rondo Ezol}

During a visit of the Rondo Ezol to Karfelon, I contacted Car-Hargon, a young man whom had consulted the
archive on a number of occasions. Speaking with him about my wish to learn more about nomadic culture and to allow
others to better understand them, he arranged a meeting with Edo-Marek, patro of the Rondo Ezol. The meeting was for
that evening, over dinner, on board the rondo.

Wanting to properly demonstrate my respect of the patro, and realizing the importance of first impressions, I prepared for the meeting by returning to the archive to procure a friendship gift, a small artifact recovered by the Elotions from the wreck of an ancient rondo. Donning my finest ceremony-robe, I made my way to the docks.

Arriving precisely at the agreed upon hour, I was intrigued to hear the strains of music coming from the deck of the ship. It was just past sundown, and the brightly colored banners flying from the rigging of the ship were illuminated by many Akvolumo, the self-contained oil lamps that the nomads favor for their safety when dropped or broken. The two large, triangular sails were furled, but it was clear that they were dyed a deep red. I was later to learn that each rondo has a distinctive color and symbol on their sails to identify them to other rondos. While the vessel was similar in many respects to merchant ships of the same class, it was obvious that the sails and the rigging were unique to this particular rondo.

I was met at the gangplank by Car-Hargon, whereupon he cautioned me to hide my ear scroll, as such holy symbols would be an insult to the elder nomads. Quickly placing it in a pouch, we boarded the rondo. I reflected that the younger members of this rondo were apparently more tolerant of the beliefs of outsiders, which was confirmed when I discovered that many of the elder nomads never left the ship, and rarely had to deal with others, while the younger ones were responsible for these outside dealings.

On deck, a great many men and women were seated in a large circle, each on a red cushion that was affixed to the decking. There were about 20 people, each dressed in different colored robes, scarves and silks. Many of them were adorned with flashy jewelry: rings, bracelets, earings, anklets, necklaces, and the like. In spite of this diversity, I did notice that all had a small symbol painted on their left cheekbone: a curved horizontal red line with an overlaid blue slash from the upper right to the lower left. Car-Hargon later explained this was a symbolic representation of an Ezol, a small flying lizard, after which this rondo was named.

One of the men was obviously more elderly then the rest, and many eyes looked to him as he carefully took my measure. I realized that this must be Edo-Marek, the patro, or leader, of this rondo. He spoke with an accent, heavily articulating his t's, k's, and other guttural consonants.

"Many have said the ways of the Brethren are for their eyes alone. That the path to enlightenment is not open to those who lean on the help of non-existent deities. But to those, I must say No! Non-believers will never follow the path if they cannot know the truth. My grandson has sought you the audience of Rondo Ezol, but only I can offer you Amikodevo."

Only later was I to learn that \fixlater{Amikodevo} was a nomad word that roughly indicates a relationship of mutual respect and obligation. It is rarely offered to outsiders, but I had previously been instrumental in clearing up several mysteries for Rondo Ezol in my position as a researcher at the Archive. Often I would be able to find corroborating evidence or information which the nomads had learned from their guardian spirits, and this information had allowed Rondo Ezol to extract itself from some difficult situations.

Edo-Marek turned to one of the older nomads and made a hand gesture. The man, Ald-Kestral, was the rondo's Sondivenki, or sound master. He took up the lute lying in his lap and began to play. Three others near him prepared their instruments and begin to weave a tapestry of sound around his emotional melody. The Patro looked at me and motioned to a cushion in the circle opposite his. I lowered myself to the deck and watched as four young men stepped into the circle and began to dance.

It was obvious from the first that their dance was symbolic in nature. Their practiced movements and careful cadence was remarkable. Each movement of the feet or hands held meaning, and each sound interacted with the dance. Not even needing to see the other dancers, each danced with confidence, as alone, but together they were one. As sculpture and painting are works of art, so truly was this dance.

As they ended, they stood in an arc in front of me, hands outstretched in greeting. The eyes of the nomads rested on me and I realized some response was needed. This was the point of failure for many Elotians in the past , but I had carefully researched what little information which had been gathered about similar rituals. I stepped forward to the Patro, and presented him with the name ring of the lost Rondo Festria, the wreck which the archive had excavated but last year. His eyes lit up as he reverently took the ring from me and showed it to his comrades. He placed it on his finger as the others cried out in joy.

It was obvious to me that I would be welcome by these people. Car-Hargon later explained that I had made a very fortuitous choice. The ring of the patro bearing the symbol of the rondo is considered the icon of the rondo, and its loss always heralded bad luck and evil. Rondo Ezol was descendant from Rondo Festria and the recovery of the ring heralded good fortune.

The musicians began to play and the rondo in general broke out in song and dance. The Patro approached me, and invited me to accompany Rondo Ezol on its trip to the nearby cities of Red Harbor and Gaila. After graciously accepting, I was accosted by two woman of the rondo, who insisted that I dance with them. All accounted for, I spent a number of hours dancing, eating, and drinking with these people. The sweet, syrupy wines of the nomads proved too intoxicating for my frame and I eventually fell into a deep sleep.

\textbf{Topic Two – Where I learn the significance of Names}

The next morning was long delayed, as the sun was already at its zenith when I at last regained consciousness. Understanding the necessity of gaining the nomad's trust, I still regretted the rather painful result of the evening's excesses. I was still collecting my wits when Car-Hargon burst into my sleeping quarters.

This was a small chamber but twelve feet long and eight wide, which appeared to be bachelor quarters for the rondo's unpaired young men. Eight pallets were arranged four to a side, half just above floor level and the others stacked above them at chest height. A single porthole at one end of the chamber lit the room with the noonday sun. The other end of the chamber housed the door, which was now open. Apparently my bag and I had been carried here after I passed out the previous evening.

"Debark Time," Car-Hargon shouted heartily. "The Rondo prepares to leave dock. You should join us." He headed up the stairs to the main deck as I forced myself
awake. About the last thing I wanted to do was to leave this rather uncomfortable, but warm, bunk, but I knew my reason for being here. Making sure my journal was safely nestled in its place in my robes, I ascended the stairs to the main deck.

On deck , it seemed that every person there had a duty to fulfill. Loading and stowing of supplies occupied many of the young men. A group of elders carefully rigged the sails to catch the midday winds. Many of the women were also busily preparing the rondo for travel, stowing deck equipment, organizing duty shifts, etc. After taking some notes I quickly realized that my continued presence would only disturb the carefully orchestrated debarkment, so I sought a quiet place to avoid obstructing the busy nomads.

Below decks, I was attracted to a very aromatic smell issuing from an open door. Stepping inside, I found myself in what only cou ld be a kitchen. Pots, pans, and utensils were everywhere, and a huge crock of stew slowly simmered in the brick fireplace. The only incongruity was the adolescent bending over the pot, carefully adding spices. About 15 years of age, he was dressed in a green one-piece shift, belted at the waist, and wearing leather thongs.

"Greetings young man," I spoke. "I was seeking shelter from the crowded deck. May I rest here awhile?"

"Of course, Sir. You must then be the archivist that came on board yesterday?" he asked.

"Yes, I am Archivist Aapar, from the archive at Karfelon. I was on deck watching the preparations to debark, but after a few near-collisions, I thought it might be best to find someplace out of the way until we are at sea."

"I can understand that, Sir." he replied. "You can see I found duties below deck as well. I am called Ejeck, I greet you."

Saying this he extended his two open palms to me, fingers stretched outward. Knowing the proper nomadic response, I held my palms to his and spoke "I greet you and
do honor to your ancestors."

"Thank you sir," he responded, "but I am not yet of age. Only those whom have joined with a guardian, and have earned their epokonom should be honored thusly."

"Forgive my error. You seemed old enough, and doing an important task. At the archive where I grew up, overseeing the preparation of food was left to adults trained in the task. May I ask you a question?"

Little was I to believe where his answer would lead me...

\hspace*{\fill}Respectfully Submitted,\\
\hspace*{\fill}Miche Aapar

Nomads, daughters and sons of Jaern, live a harsh existence, devoid of the more civilized comforts that modern society can provide. Traveling from island to island, never calling one place home, they eke out a meager living peddling their skills and entertainments to the inhabitants of the places they visit.

\section{History}
When the rogue planet \fixlater{Jaern} was first detected approaching \fixlater{Torandor}, the peoples there realized that their very world was threatened. Masses of people flocked to the city of Rigel, where the \fixlater{Kaaren of Destruction} was being prepared for its journey into space in an attempt to destroy the rogue. Among these people was a small group of men and women from the offshore island of Alborn. They believed in peace, love, and harmony, and tried to avoid the many evils that plagued mankind.

When they first heard of the coming destruction of the world they did not believe the doomsayers, as this had been prophesied often in the past. But when they consulted the spirits of the land and of their departed, they learned that the end was indeed at hand. Traveling to nearby Rigel, they purchased passage on the Kaaren at the cos of all their possessions and wealth. As the time to debark drew near, they were betrayed by Jhireen, the priest of Neptune who had made the arrangements. They watched in horror as the Kaaren rose skyward and the intruder, Jaern, loomed larger and larger in the night sky.

Jaern's proximity to Torandor triggered severe storms and earthquakes. The quakes started massive fires in the cities, and as the flames spread, it became apparent that the Kaaren had failed in its task. The end of their home world was at hand. In their despair, they called upon the gods of Torandor to hear their plea for rescue. In vain they called, forat that very time, many of the deities were busy rescuing themselves from the impending destruction of their world. With the city around them in ruins, great gaping rips in the land spewing forth molten rock, and the air itself being ripped from the land, the Albornians angrily spurned the gods. They then and there vowed never to rely on any one other then themselves for their direction and destiny.

With the final pass of Jaern, Torandor burst into countless fragments, each going off into space. The Albornians perished, having been betrayed, abandoned and left helpless. Having denounced their deities, these peoples' spirits went neither to the gods, or to the depths of Infero, but resided without resolution in the trackless emptiness of the Kurago.

The refugees from the Kaaren landed on the rogue planet, which assumed a regular orbit around Onra. Years passed, the atmosphere thawed, and the icy seas melted. From the store of Torandorian seeds stowed aboard the Kaaren, life sprang forth again. The refugees started a new civilization on the planet they once sought to destroy.

Within the Kurago, the spirits of the Albornians refused to accept their fate. Searching for a way to return to the material world, they eventually encountered a spirit called \fixlater{Llan of the Five}. He had been a powerful magician and hero in life, and was awaiting the time of need to come forth again. Llan showed the Albornians how to contact the spirits of those still alive. The Albornians found those people most able to communicate with them, becoming their Guardian spirits.

These people merged with the Albornian guardians to become the first generation of nomads. They sought each other out and banded together, calling themselves \fixlater{The Brethren}. They learned to live on the seas of Jaern, and withdrew from the society of those who had crossed from Torandor on the Kaaren. The Brethren blamed the refugees for the destruction of Torandor and vowed to never place themselves at the mercy of the unbelievers and their deities.

\section{Social Structure}

Nomads grow up and live together in groups called Rondos of about 12 to 24. Most of the members of a Rondo are related by blood or marriage. The rondo is lead by the senior-most male elder, called the Patro, who decides all matters of importance. Most rondo owns a two masted ship that sports triangular sails painted in the distinctive colors and symbol of the rondo.

\section{Training}

Children are reared by all the adult members of the rondo. While they often know their biological parents, all members of the rondo are responsible for the children's physical and emotional well-being. They are taught the  traditions and ways of the nomads at any early age. At the age of 16 they are encouraged to take the Memvoj, the journey of self discovery, to go out on their own for at least a year to experience independence, self-reliance, and how to deal with unbelievers.

\section{Induction}

When the young ones return from their travels, they are welcomed back, and a ceremony of seeking, The Seraei is performed. They are given a certain trance-inducing mixture that connect them for the first time with the Kurago, the spirit plane. They wander about until they meet a spirit with whom they will form a life-long bond. This guardian spirit will be their guide to the forces of Kurago and will help them learn to harness them. In return, the inductee allows their guardian spirit the use of their body to access the normal world.

If a full day passes without the inductee bonding with a guardian spirit, they is brought out of the trance and allowed to rest. They will never gain a guardian spirit, and so can create no incantations, mixtures, perform any rituals, or hold any authority within the rondo. These unfortunates are not looked down upon. They may stay with the rondo, for they may aspire to become spirits of the Kurago when they pass on.

\fixlater{All adventurer nomads are considrered having succeeded in establishing a bond with a guardian spirit.}

\section{Nomadic Missions}

Throughout a nomad's childhood, they see and learn from the adult nomads around them. Depending on their particular talents, they will end up specializing in one of four missions. Each mission dictates a particular lifestyle and will determine which style of interaction with the Kurago which the nomad uses most easily.

\fixlater{The nomad's player should choose which mission to play from the four below.}

\subsection{Preservers}

Keeping of the old ways of The Brethren are very important to nomads of this mission. They form the bulk of those whom spend their lives aboard the rondo. They are responsible for the health of The Brethren. They maintain the ships, raise the young, manage their mercantile trade, and lead their rondos. The Patro and those with skills related to sailing are generally preservers.

\subsection{Troubadors}

Responsible for keeping the memory of the past alive, these nomads use their skills to remember and retell the histories of all the peoples they meet. At each port of call, performances are arranged where the nomads play music, sing, and dance to convey stories of ancient heroes, great tragedies, and lost loves. Townspeople and others from all around come to these festivals to listen to the old stories and celebrate life. The other nomads sell wares and crafts from far away, helping support the rondo. Aboard the rondo, the troubadors remember and repeat all the intricate dances that tell the stories of the nomads' past and the Ways of The Brethren. These stories record their history and teach their young. These dances are usually kept private to the rondo, except at the time of The Gathering.

\subsection{Seekers}

A nomad's guardian spirit will sometimes have passed on to the Kurago with their life task incomplete. In its bonding with the nomad, it will strike a bargain to correct this. These nomads will often leave their rondo to travel the world, at the direction of their spirit, in an attempt to complete an unfinished quest. These quests often involve the righting of past wrongs, or repayment for a mistake in the guardian spirit's past. Sometimes these tasks will span several nomad lifetimes, and many hundreds of years. The skills and incants these nomads develop assist them in their quest.

\subsection{Second Lifers}

A particularly forceful spirit can completely submerge the life force of the nomad, taking the body as their own. Such a spirit and its captive nomad is called a Second Lifer, for the spirit is attempting to live a second life through the nomad. These spirits are so strong that when the nomad does die their own spirit is destroyed, with no chance to go to the Kurago. This final destruction is particularly hated and feared by other nomads, and, curiously, by the Onivero as well, for the spirits' powerful life force violates their ethics.

Nomads and Onivero alike hunt down these unfortunates, and attempt to part them from their guardians spirit. Unfortunately, all the easiest ways of accomplishing this involve the death of the nomad, but even that is preferred to the final destruction of their soul, which would happen should the second lifer keep control.

\fixlater{If your GM allows this mission, and you elect to be a second lifer, your adventurer will end up being the spirit controlling the nomad, rather than the nomad himself. You need to work with your GM in creating a second adventurer card with the stats needed for this spirit. They will assist you in creating a past life and picking a particular motive and quest. During play of a second lifer, the GM will occasionally ask you to make checks to stay in control of the nomad, and if they fail, They will roleplay the nomad until you regain control.}

\subsection{Guardian Spirits}

All guardian spirits are souls of deceased nomads seeking access to the real world. When the nomad binds to this spirit, they share the nomad's life force. The nomad can sense the presence of their guardian, and can occasionally silently converse with them. As the nomad grows older, this ability increases, allowing the nomad to experience some of their guardian's emotions, and then their memories. The division between nomad and guardian becomes more and more indistinguishable as the nomad grows old. When a nomad passes on, their own spirit melds with their guardian's, and from the Kurago seeks a new nomad with which to bind. Nomad name prefixes, or epokonoms, are based on the number of lives that the nomad's guardian has experienced.

\begin{normbox}[Nomad Epokonoms]
\small
\begin{tabular}{@{}l l l l}
Raz &1&Edo &7\\
Car &2&Ijo &8\\
Oka &3&Bez &9\\
Vem &4&Sag & 10+\\
Lar &5&Tor & Original Albornian\\
Ald &6&&\\
\end{tabular}
\end{normbox}

The standing of a nomad within the rondo depends greatly on the number and standing of their bonded spirit as
well as their  own skills and accomplishments.

\fixlater{After the player chooses their adventurer's mission, The GM may, at their discretion, give the adventurer the name of their guardian spirit. As the adventurer plays, the GM may slowly give out more information about the guardian, and can, at their option, converse and generally roleplay the guardian. The guardian can never give the nomad any skills or abilities that the nomad has not earned and bought with experience, but it may give information.}

\section{Beyond Death}

When a nomad dies, their spirit merges with that of their Guardian Spirit and enters the Kurago. Time passes as the newly merged spirit recuperates from the traumatic death and joining and its thoughts, ideas, motives and knowledge stabilizes. These unbound spirits then seek out living nomads awaiting their initial bonding.

\fixlater{Your GM may allow the deceased nomad's player to take on the role of a guardian spirit to another player's newly created nomad. If this occurs, both players play the adventurer's body, the guardian only providing advice, ideas and conversing freely with the new nomad. A player may not play both adventurers as this would not allow them to separate the thoughts, knowledge and emotions of each fairly. Your GM might allow your dead nomad's merged spirit to play as a Second Lifer if the personality of the spirit is willful and a bit ruthless. But be warned that this is a hard course, and other nomads and Onivero alike will seek their destruction.}

\section{At Sea}

After induction, most nomads stay within their rondos for the rest of their lives. They travel with the rondo from port to port, trading goods and entertaining the people they meet to earn the silver and goods necessary to support their needs. The visit of a rondo to a small town is often cause for a holiday, as the people turn out to hear the music and stories of the nomads, buy the exotic clothes and trade goods from afar, seek answers about the future and the spirit world, and buy elixirs of health and tokens of good fortune.

When children grow and the number of nomads within the rondo becomes more than can be supported, the nomads visit one of their base islands, where the few that stay on land will assist them in building a new ship. When complete, some members form a new rondo, and go their own way. The excess members of several rondos sometimes combine in a new rondo.

\section{On Land}

Some nomad rondos have chosen to abandon life at sea and have established villages on land. Usually these will be on islands, remote from other peoples. These nomads will have grown very skillful at farming and the domestication of animals, and are a valued source for well trained horses. Many of these rondos will serve as a "home base" for a number of mobile rondos, providing a place to communicate, to trade and to seek mates for their young.

\section{Clothing}

Clothing styles among the nomads depend on the places they have visited. Attempting to predict what style a nomad would wear, or telling a nomad apart from others based on clothing, is fairly pointless. Being at sea, they usually wear appropriate clothing when onboard, including clothes unlikely to get in the way while sailing, light footwear, and a strip of cloth called a tamenwrap which they wrap about their forehead to hold back their hair. They do have a tendency toward large and flashy adornments, and often paint their rondo's symbol somewhere on their bodies.

\section{Music and Dance}

No one within a rondo is more loved and respected then the Sondivenki, or sound master. They are the senior-most musician, and besides a required skill at musical instruments of all kinds, they know and play all the songs of the past, a sort of audio history of The Brethren. During their lifetime, each Sondivenki is responsible for adding another song to their rondo's songs, and teaching them all, note and word perfect, to a successor. No rondo is thought to be destroyed if anyone knows all its songs.

The most agile and strong of the young men often learn nomad dances to accompany their rondo's music. Dances are used for celebrating, preparing for confrontations, espousing affection for a young woman, and just about any other reason. Movement sequences are tied to meanings, making dance almost as expressive as speech and much more entertaining. The communication between the dancer and the musician is almost telepathic, the music melding itself into an extension of the dance, and the dance a part of the music.

\section{The Gathering}

Once every five years, all the rondos in a given region will gather in one place to learn what others have done, swap goods, songs, stories and often young people
(through marriage). As the appointed time approaches, the nomads sail toward the appointed place, decided by various omens and signs that all rondos can see and decipher. Once there, the Rondos are lashed together, and for the next four weeks a vast celebration is held. This renewal of their brotherhood keeps them a united people.

\section{Life Force}

Nomads use their \fixlater{Life Force} to open a conduit to the Kurago when creating their incants. A nomad's life force starts as the sum of their \HEA and \PER stats, but can be increased separately. Creating an incant depletes this life energy. One night (8 hours) of rest restores one fifth of this starting life force.

If preparing an incant completely depletes a nomad's life force, their body dies, and their spirit retreats to the Kurago, where it mergess with their Guardian Spirit. The resulting spirit rests for time there, and then seeks to merge with another new nomad.

\section{Preparing Incants}

Nomads perform Rituals that channel energy from the Kurago to create incants. There are five types of incants. Mixtures are powders, liquids, and sprays that manifest various effects. Talisman are objects with bound magic. Songs effect the listeners in many ways. Imprints leave some spiritual connection on targeted people. Invocations provide a path for the spirits of the Kurago to interact with the real world. Each type of incant is created differently.

Incants come in different strengths and are arranged in ranks, from first, the weakest, to sixth, the strongest. To have the ability create an incant, the nomad must buy the incant up to the desired rank. The base cost for each incant is specific to the incant, and is modified by the mission of the nomad. Incants within the nomad's \fixlater{Mission} are purchased at \tcdefine{base cost}, while incants outside their mission are purchased at \result{2x} the base cost.
Adventurers whom are inducted as nomads from outside the rondo pay \result{3x} the base cost for incants in their declared mission, and \result{4x} for incants
of other missions. Priests, or anyone having any divine magic whatsoever may not buy incants as they are barred from opening a conduit to the Kurago.

Imprints and talismans marked "permanent", having permanent effects, have to be repurchased from zero after they are created. Other incants are limited only by the current life force of the nomad.\\

\begin{normboxlist}[Preparing Incants]
\begin{enumerate}
\item A nomad first must learn the desired incant.
\item Once learned, it is assumed that they are able to acquire most of the materials for the cost listed under "Silver."
\begin{itemize}
\item[] Any ingredients listed under "Ingredients" must be acquired by the nomad. This can be hand-waved as part of the Silver cost, done as a flat fee, a die roll, or acted out as part of the adventure for especially rare ingredients (GM's discretion). 
\end{itemize}
\item The nomad spends the required time preparing the incant, expending a portion of their Life Force matching the rank of the incant they wish to prepare.
\begin{itemize}
\item[] If interrupted, the materials are not wasted and the Life Force is not expended, but the ritual must be restarted from the beginning.
\end{itemize}
\end{enumerate}
\end{normboxlist}

\subsection{Mixtures}

Powders, liquids and sprays that are ingested or spread on the skin of the target have a myriad of different effects and potencies.  To create a mixture, the nomad must posess any required rare ingredients. For targeted incants he must have something from the target, a lock of hair, a scrap of worn clothing, a favorite tool or weapon. The nomad then starts the Ritual, using their Life Force to open a conduit to the Kurago while speaking or singing the words and performing the proper mixing actions. Mixtures are divided by how they are activated.

Delivery Methods:
\begin{itemize}
\item Generics: These mixtures can be used by anyone after they are created, and are activated by ingesting , spreading or spraying on the target.

\item Triggered: These prepared mixtures must be released or administered by the nomad by whom they were created. The nomad administers the mixture, reaching into their mind to open the connection for the spiritual energy to act.

\item Targeted: These are similar to Triggered mixtures, but when they are created, something of the eventual target must be used. These most powerfull mixtures only effect the one person targeted by the nomad.
\end{itemize}

\subsection{Talismans}

A physical focus is needed for these incants. The material of the focus is specific to each incant, but its form is chosen by the incanter. The nomad must form the object into a shape that has meaning to them, and then go into a trance, chanting the incant, communicating with the Kurago until the talisman is complete. Their life force opens the conduit to the Kurago, and if the incant has a permament effect, they must spend the experience up to the rank of the incant.

\subsection{Songs}

With the assistance of any required instrumentalists or singers, the incanter performs the incant. Those hearing the incant are subject to its spiritual effects.

\subsection{Imprints}

Chanting the incant while creating the tattoo somewhere on the target's body, the incanter pull's the  appropriate spirit from the Kurago, binding it to the target. If the incanter is imprinting themselves, another nomad tattooist may assist in creating the tattoo. These incants require a specific ink, but the art work is always unique to the tattooist, and identifiable to others of their rondo. Since all imprints have a permanent effect, the nomad must expend the experience to buy this up from rank zero each time it is created.

\subsection{Invocations}

Invocations allow the spirits from the Kurago access to the real world. The incanter chants the incant, going into a trance opening the connection. The selected spirit then manifests itself through the nomad or physically depending on the incant. The nomad must expend the needed life force to create this connection. Unless otherwise stated in the specific incant, the spirit manifesting itself must be willing for these incants to work.



\end{multicols*}
\setlength{\columnsep}{\defcolwidth}\begin{multicols*}{4}
\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{itemize}[leftmargin=0pt]
	\item[] \textbf{\ul{Core Incants}}
	\begin{itemize}[leftmargin=0pt]
		\item[] Abeyance Ritual
		\item[] AkvoSeraei
		\item[] Aversion Amulet
		\item[] Armoring Powder
		\item[] Assume
		\item[] Bond Summon
		\item[] Dark Transcript
		\item[] Elixir of Wakefulness
		\item[] Endow Thought
		\item[] Eviction Ritual
		\item[] Green Touch
		\item[] Life Light
		\item[] Living Staff
		\item[] Memory Elixir
		\item[] Memory Stone
		\item[] Object of Spirit Snare
		\item[] Object of Spirit Ward
		\item[] Philter of Regression
		\item[] Portal Disruption Powder
		\item[] Revealing Mist
		\item[] Sleep Draught
		\item[] Sleep Mist
		\item[] Smoke Bombard
		\item[] Spirit Barricade
		\item[] Spirit Bombard
		\item[] Spirit Call
		\item[] Spirit Cage
		\item[] Spirit Guard
		\item[] Spirit Sentinel
		\item[] Spirit Sound
		\item[] Stone Guides
		\item[] Third Eye
		\item[] Wind Call
	\end{itemize}
	\item[] 
	\item[] \textbf{\ul{Preserver Incants}}
	\begin{itemize}[leftmargin=0pt]
		\item[] Aphrodisiac Oil
		\item[] Arithmetic Snuff
		\item[] Brew
		\item[] Courage Draught
		\item[] Corpse Ward
		\item[] Deep Sleep
		\item[] Descry Ships
		\item[] Dram of Energy
		\item[] Elixir of Health
		\item[] Evaluator
		\item[] Find Weakness
		\item[] Future Scry
		\item[] Hangover Liquor
		\item[] Healing Sleep
		\item[] Infertility Lotion
		\item[] Love Potion
		\item[] Luck Ritual
		\item[] Mask of Disclosure
		\item[] Powder of Preservation
		\item[] Poultice
		\item[] Power Patch
		\item[] Remedy
		\item[] Sago Weed
		\item[] Ship's Veil
		\item[] Shore Object
		\item[] Stimulant Chew
		\item[] Strength Patch
		\item[] Suppress Pain
		\item[] Tonic
		\item[] Will Patch
	\end{itemize}
	\item[] 
	\item[] \textbf{\ul{Troubador Incants}}
	\begin{itemize}[leftmargin=0pt]
		\item[] Animal Dance
		\item[] Baton of Dancing
		\item[] Calling of the One
		\item[] Co-Ordination Jig
		\item[] Dance of Distraction
		\item[] Dance of the Lights
		\item[] Erasure Song
		\item[] Heart of Courage
		\item[] Lute Lamp
		\item[] Message Song
		\item[] Mood Ring
		\item[] Pacifier Pipe
		\item[] Robes of Style
		\item[] Rolling Hills
		\item[] Shadow Cream
		\item[] Sleep Awaits
		\item[] Spirit Dirge
		\item[] Still Waters
		\item[] Stone Song
		\item[] Synchro Dance
		\item[] The Howling
		\item[] The Mocking
		\item[] The Praising
		\item[] The Telling
		\item[] Wary of the Night
		\item[] Yes Song
		\item[] Zephyr Polka
	\end{itemize}
	\item[] 
	\item[] \textbf{\ul{Seeker Incants}}
	\begin{itemize}[leftmargin=0pt]
		\item[] Acid Draught
		\item[] Animal Bind
		\item[] Anklet of Dexterity
		\item[] Armband of Fortitude
		\item[] Assimilation
		\item[] Cinnamon Bane
		\item[] Ethereal Bane
		\item[] Crystal Gate
		\item[] Ferret Object
		\item[] Fertility Seed
		\item[] Fire Shield
		\item[] Gargoyle Mark
		\item[] Immaterial Bane
		\item[] Insight Gem
		\item[] Iron Tonic
		\item[] Moon Boots
		\item[] Moon Crystal
		\item[] Moon Shroud
		\item[] Moon Sleep
		\item[] Shock Buffer
		\item[] Silver Bane
		\item[] Spirit Bane
		\item[] Soul Sight
		\item[] Spirit Exchange
		\item[] Spirit Guise
		\item[] Spirit Scout
		\item[] Spirit Trance
		\item[] Spirit View
		\item[] Tamener
		\item[] Thermal Gel
		\item[] Time Shifter
		\item[] Tongue Ring
		\item[] Underwater Gel
	\end{itemize}
	\item[] 
	\item[] \textbf{\ul{Second Lifer Incants}}
	\begin{itemize}[leftmargin=0pt]
		\item[] Addiction Potion
		\item[] Amulet of Deception
		\item[] Arid Seed
		\item[] Arthritis Potion
		\item[] Bond Pain
		\item[] Brain Fevers Draught
		\item[] Dose of the Itch
		\item[] Grippe Dose
		\item[] Fear Draught
		\item[] Leprosy Tonic
		\item[] Mark of the Lamprey
		\item[] Mark of Homing
		\item[] Ointment of Corruption
		\item[] Pain Smoke
		\item[] Plague Dram
		\item[] Powder of Phantasm
		\item[] Ring of Likeness
		\item[] Ring of Subjugation
		\item[] Scourge
		\item[] Sleeping Sickness Solution
		\item[] Spirits of Consumption
		\item[] Vial of Ill Omen 
	\end{itemize}
\end{itemize}
\end{multicols*}
\pagebreak
\section{Core Incants}

\incantentry{Abeyance Ritual}{Invocation}{10001}{2000 / Rank}{4 Hours / Rank}{}{A pact with the spirits allow the nomad to hold back the effects of aging. The incanter appears not to age for \tcdefine{10 years} for each rank of this incant. At the end of this time the nomad's appearance reverts back to what would be appropriate for their true age.}

\incantentry{AkvoSeraei}{Mixture/Triggered}{100}{50 / Rank}{2 Hours / Rank}{}{This potion, used in the \fixlater{Seraie}, opens a mind to communication with the Kurago for the first time so a potential nomad can find and bond with his guardian spirit. The imbiber must \ul{FAIL} a check of the ritual's rank in dice against \HEA. Failing this check puts the target into the proper trance. If the potion fails, repeated attempts with another \nspell{AkvoSeraei} are allowable. \fixlater{Other nomad rituals use the potion made by the nomad attempting the ritual.}}

\incantentry{Aversion Amulet}{Talisman}{1200}{100 / Rank}{4 Hours / Rank}{0.5 ounces Terisium / Rank}{When activated by the incanter this talisman reduces the effect of magic on its wearer. The wearer takes a number of dice off the check of a spell equal to the ranks of the talisman and further reduces one die of damage on any damaging spell. The talisman is effective for \measure{1 Week} after it is activated. }

\incantentry{Armoring Powder}{Mixture/Triggered}{600}{25 / Rank}{3 Hours / Rank}{Groken shell}{When the incanter activates this powder, spreading it on the skin of a target, the skin hardens making them harder to hit. For every rank of this incant the target's \CDV and \MDV is increased by \result{+1}. This effect lasts \measure{4 Hours} from the time it is activated.}

\incantentry{Assume}{Invocation}{1000}{30 / Rank}{1 Hour / Rank}{}{Allow the incanter's guardian spirit to temporarily gains access to the material world. The nomad's spirit is set aside and the guardian spirit takes possession of the nomad's body. The nomad's appearance changes to resemble the guardian's semblance in life. All magical means of detecting, scanning and identification will sense the spirit instead of the incanter. However, a mirror will reflect hte image of the nomad, not the guardian. This possession is a large strain on the nomad's body, and after one hour the spirit must leave, or make a \tcdefine{\tcdieroll{4d6} vs \HEA every 10 Minutes}. Failure on the check results in the death of the body, trapping the noad in the Kurago and destroying the guarding spirit. Since each spirit is usually a gestalt of a number of spirits, each rank allows a spirit one further back in the chain sole access during the invocation.}

\incantentry{Bond Summon}{Invocation}{500}{20 / Rank}{0.5 Hours / Rank}{Target sample}{The nomad requires something from the body of the target, like a piece of hair or a fingernail cutting. The nomad creates an effigy of the target, insering the sample, and calls the target to them. The target must make a check of \tcdefine{\tcdieroll{1d6} per rank vs \WIL or travel as fast as possible to the incanter.}}

\incantentry{Dark Transcript}{Imprinting}{1000}{75 / Rank}{3 Hours / Rank}{}{Creates a tattoo of an eye on the incanter's body, and lures an animal spirit to the tattoo. If the tattoo is exposed while the nomad sleeps, images seen by the spirit during their sleep appear in their mind when they awaken. The spirit can perceive up to \tcdefine{10 Feet per Rank} of the incant. Creating a new \nspell{Dark Transcript} releases any previous animal spirits bound to this incant.}

\incantentry{Elixir of Wakefulness}{Mixture/Triggered}{600}{25 / Rank}{2 Hours / Rank}{}{This bitter potion keeps the target awake and alert. The target cannot fall asleep by tiredness, physical exhaustion, or spells, but can be knocked unconsious by damage. The target stays awake for a number of hours equal to the rank of the mixture. After this time the target will immediately fall into a deep sleep for the same amount of time, during which the target does not regenerate \DP, \DU, or \EU. }

\incantentry{Endow Thought}{Invocation}{200}{5 / Rank}{1 Minute / Rank}{}{The nomad touches an object, or part of a large object and concentrates on one sight, and one impression or thought, of \tcdefine{eight words or less}. The vision and impression are then felt by anyone touching the object within the next \tcdefine{2 Days per rank} of the invocation. }

\incantentry{Eviction Ritual}{Invocation}{300}{60 / Rank}{2 Hours / Rank}{}{The dreaded Second Lifer becomes the responsibility of every nomad who encounters one. This invocation gives a nomad a chance to remove this parasitic spirit from the its victim. The nomad begins by chanting for at least a round for every rank of the ritual. At the end of the chant, the possessing spirit must make a \WIL check of a number of dice equal to the rank of the invocation. Failing this roll forces the spirit from body and back to the Kurago, banishing it for at least as many years as the rank of the ritual. The formerly possessed person retains the memory of all the events that occurred during their possession.

Being trapped in their own body, unable to control any of their actions can unhinge even the most sane of minds. The freed person gets a \tcdieroll{4d6} check against \WIL to retain sanity. If failed, they will be unable to perform any takss and will slowly die within a few weeks. The GM may modify this to suit events of the possession or the support of the possessed.

If they stay sane, they no longer possess a guardian spirit, and have no nomadic abilities. They may attempt another inductuon again, should they wish.

This is a difficult ritual to perform alone, as the second lifer will always react violently to prevent the completion of the chant. Many nomads are forced to simply kill the body, which will at least allow both spirits to go to the Kurago.}

\incantentry{Green Touch}{Invocation}{200}{5 / Rank}{1 Minute / Rank}{Powdered Oorn root}{The nomad touches a damaged plant, and it is restored to health. The \textbf{1st Rank} of this invocation could restore a wilted flower, the \textbf{2nd Rank} would restore a small bush, the \textbf{3rd Rank} a single tree, the \textbf{4th Rank} would restore a \measure{10 Foot Radius} circle of plants and each higher rank would \result{double} the area. This can only help a still living plant; dead plants are unaffected.}

\incantentry{Life Light}{Talisman/Permanent}{500}{20 / Rank}{2 Hours / Rank}{Target's blood\\
Silver dust}{A small vial is filled with a mixture of silver dust and the blood of a targeted individual. The nomad holds the vial and touches the target during the creation of this talisman. The vial begins to glow softly with a silver glow, as long as the target is still alive and within \result{50 x Rank} mets of the vial.}

\incantentry{Living Staff}{Talismen/Permanent}{500}{20 / Rank}{2 Hours / Rank}{Wooden staff}{This talisman is produced by the nomad infusing a wooden staff with the power to grow again. When activated by the incanter the staff sprouts branches that surround the staff and the nomad. This growth hides the nomad if a searcher fails a \PER check of \measure{\tcdieroll{1d6} per rank} of the ritual. The growth remains for \measure{1 Hour} or until the incanter commands it to reutrn to normal. The staff must be planted in the ground for \measure{8 Hours} before it can be used again.}

\incantentry{Memory Elixir}{Mixture/targeted}{1200}{250 / Rank}{4 Hours / Rank}{Target's blood\\
Honey\\
Ulihar leaves}{This elixir reverses the effect of memory loss through old age, dementia, amnesia, and spirit magic effects. The blood of the recipient is mixed with honey and the leaves of the Ulihar tree. Upon drinking this mixture the recipient goes into a state of trance and remembers what was forgotten. They remain in this state for \tcdefine{1 Hour per Rank} of the mixture and recover memories of the last \tcdefine{5 Years per Rank}. THis elixr only works once on a given indivual.}

\incantentry{Memory Stone}{Talisman/premament}{300}{20 / Rank}{2 Hours / Rank}{Small stone}{Holding a small stone in his hand the nomad concentrates on one sight, and the emotions associated with the sight. This could be the face of a loved one, a pleasant secret alcove, or a stark scene of devastation. Placing themselves into a trance, this sight and emotion, as they experienced them, will be conveyed back to the incanter when they touch the stone. At \textbf{Rank 2}, others may feel the memory. At \textbf{Rank 3}, the memory is detailed enough to use as a target of a spell requiring a memorized target, \ul{but not a memorized location.} At \textbf{Rank 4}, the sight envelopes the one holding the stone, as a visible image. At \textbf{Rank 5}, the image overwhelms reality, causing the viewer to only see this sight. At \textbf{Rank 6}, all others within \measure{10 Feet} also are enveloped in this image.}

\incantentry{Object of Spirit Snare}{Talisman/Permanent}{500}{30 / Rank}{2 Hours / Rank}{Quartz crystal}{The nomad prepares a large quartz crystal to trap spirits that it comes into contact with. Any spirit other than the guardian spirit of the creator who comes into contact with the crystal must make a check against \WIL of \tcdefine{\tcdieroll{1d6} per Rank} or be trapped. A spirit may only leave the crystal if it is broken. An elemental \espell{Revocation} or divine \dspell{Revocation} treats this as a spell of \result{2x the Incant Rank}, shattering the crystal on a successful revocation.}

\incantentry{Object of Spirit Ward}{Talisman/Permanent}{300}{60 / Rank}{2 Hours / Rank}{Symbolic object}{The nomad designates a symbol expressing a symbolic meaning to them particular to one spirit that can be also brandished in one hand. Drawing power from the Kurago, the object becomes anathema to that spirit. If it attempts to approach the person brandishing this object, it must make a check against \WIL of \tcdefine{\tcdieroll{1d6} per Rank}. This object breaks, crumbles, bursts into flames, etc. the first time a spirit succeeds against it.}

\incantentry{Philter of Regression}{Mixture/targeted}{800}{100 / Rank}{3 Hours / Rank}{Target's blood\\
Ulihar sap\\
Exotic spices}{This philter lets the nomad help a willing imbiber to voluntarily regress in experience, knowledge, and spirit. Preparing a mixture, the nomad chants and contacts the spirits, requesting they take away the knowledge to be regressed. When imbibed, the imbiber watches the events of their life pass before their eyes backwards, starting with the most recent events. As each event is viewed, the memory of it is erased from their mind. Each rank can regress up to \measure{5 Years}. This mixture is only effective on any one indivual once in their life.}

\incantentry{Portal Disruption Powder}{Mixture/targeted}{600}{20 / Rank}{2 Hours / Rank}{0.5 Oz of Terisium / Rank\\
Target Sample}{This powder disrupts teleportation. When prepared to target a person, a sample from that person is needed. After being applied to that person, if they attempt to teleport or use a portal of any kind they must make a check of the incant rank vs \PWR to succeed. 

When thrown through an established portal, the portal ceases to function as it previously did for a number of hours equal to the rank of the mixture. 

When spread through an area no greater than \measure{10 Feet by 10 Feet}, it causes anyone attempting to teleport in or out to make a check of the rank of the mixture vs \PWR to succeed in the transport.}

\incantentry{Revealing Mist}{Mixture/Triggered}{300}{25 / Rank}{2 Hours / Rank}{}{This mixture is held in an easily broken glass container. When broken, the mixture mixes with air to generate a cloud of orange smoke. The cloud is \result{10 Feet per Rank} of the mixture in radius and is subject to normal wind effects. Any invisible objects within are clearly outlined within it.}

\incantentry{Sleep Draught}{Mixture/Triggered}{500}{30 / Rank}{1 Hour / Rank}{}{This draught causes the imbiber to fall into a deep restful sleep. If unwilling, the imbiber must succeed a check of against \WIL of \result{\tcdieroll{1d6} per Rank} to resist its effect. They will then sleep for at least as many hours as the rank of the mixture.}

\incantentry{Sleep Mist}{Mixture/Triggered}{600}{50 / Rank}{2 Hours / Rank}{}{This mixture is held in an easily broken glass container. When broken the mixture reacts with the air to form a cloud a blue, pleasant smelling smoke. The smoke is subject to wind effects and has a radius of \measure{10 Feet}. Humanoids caught in this smoke must make an \RC against \WIL of \result{\tcdieroll{1d6} per Rank} or fall into a magic sleep for \tcdieroll{1d6} Hours. }

\incantentry{Smoke Bombard}{Mixture/Triggered}{400}{20 / Rank}{1 Hour / Rank}{Torgon mist organ}{This mixture is held in an easily breakable vial. When activated the powder ignites with the air, forming a dense cloud of black, impenetrable smoke that blocks all vision. The cloud is subject to wind effects normally and covers a circular area of 10 fee}
%Core
\incantentry{Spirit Barricade}{Talisman/Permanent}{1200}{150 / Rank}{4 Hours / Rank}{}{The nomad infuses an article of jewelry to create this talisman that hampers spirits to which he does not have a bond. Any disenbodied spirit attempting to interact with the incanter must make a check against WIL of one die per rank of the talisman. If fa}
%Core
\incantentry{Spirit Bombard}{Mixture/Triggered}{400}{40 / Rank}{2 Hours / Rank}{}{This mixture consists of a powder in a easily broken glass container. On contact with the air the powder produces a cloud of red scintillating smoke. This smoke is affect normally by wind effects and has a radius of 10 feet per rank of the mixture. Any sp}
%Core
\incantentry{Spirit Call}{Invocation}{500}{25 / Rank}{5 Minutes / Rank}{}{The nomad uses this ritual to call forth a spirit from the Kurago to manifest itself in the material plane. Those present must concentrate on him and not disturb the area with negative thoughts. The nomad spends the time building up the ritual, and then w}
%Core
\incantentry{Spirit Cage}{Talisman}{1200}{150 / Rank}{2 Hours / Rank}{}{The nomad constructs an enclosed cage or container with no holes or gaps greater than two inches. The rank of the ritual times 3' is the maximum of any dimension of this container. Once complete the cage isolates its interior from the Kurago and any spiri}
%Core
\incantentry{Spirit Guard}{Imprinting}{2000}{500 / Rank}{4 Hours / Rank}{Creature blood
Rank  Creature
1  Osser
2  Torgon
3  Vatrole
4  Jacer
5  Gargoyle
6  Fajrodrako}{The nomad calls a spirit to defend him in the case of misfortune. He must have a small amount of blood from a creature of the same species. While being tattooed the nomad seeks out an animal spirit to aid him. The tattoo must resemble that animal. Wheneve}
%Core
\incantentry{Spirit Sentinel}{Imprinting}{1000}{300 / Rank}{2 Hours / Rank}{
Rank Creature
1  Marokrifo  stinger
2  Batisunui  fangs
3  Hejtifea   feather
4  Venevermo  digesting organ
5  Flugofiso  tenticle
6  Quezl      poison sac}{The tattooist begins by forming the image of a venomous creature upon the body of the target. The incanter calls for a spirit of the tattooed animal to reside in the tattoo. Once the animal spirit enters the tattoo, if the target fulfills the condition of}
%Core
\incantentry{Spirit Sound}{Talisman/Permanent}{100}{400 / Rank}{1 Hour / Rank}{Small conch shell}{Through this ritual a nomad infuses an earring to allow him to hear spirits. When worn the nomad makes a check of 6 minus the rank of the ritual in dice against PER. If the check is successful the nomad can faintly hear the nearby spirits.}
%Core
\incantentry{Stone Guides}{Talisman/Permanent}{500}{20 / Rank}{0.5H / Rank}{Identical pearls}{The nomad gathers a number of nearly identical small pearls. He may use as many pearls as the ritual rank plus one. After the ritual is complete, each pearl will glow if it is within 1 met times the rank of the ritual from any of the other pearls. The bri}
%Core
\incantentry{Third Eye}{Imprinting}{1000}{200 / Rank}{3 Hours / Rank}{}{Needing the aid of a tattooist, the nomad has a tattoo of an eye placed somewhere on his body. While this is occurring, he goes into a deep trance and lures an animal spirit to the tattoo. The nomad is afterward able to see from this eye as well as his no}
%Core
\incantentry{Wind Call}{Invocation}{400}{10 / Rank}{5 Minutes / Rank}{Parchment}{Preparing a message on a piece of parchment, the nomad folds the parchment into an aerodynamic shape, and begins to chant while concentrating on the receiver of the parchment. His guardian spirit infuses the physical manifestation of the parchment as the }

\section{Preserver Incants}

%Preserver
\incantentry{Aphrodisiac Oil}{Mixture/generic}{400}{20/rank}{1 Hour / Rank}{Genitalia of a tiger}{This oil is applied to the target's body causing the target to become more excited by physical pleasure. The target is able to enjoy the experience of intimacy better and longer. This experience lasts one house for each rank of the incant.}
%Preserver
\incantentry{Arithmetic Snuff}{Mixture/generic}{400}{15/rank}{1 Hour / Rank}{}{This powder when sniffed through the nose awakens latent areas in the user's brain. Tasks involving mathematical abilities are made at a die less per rank of the mixture. Also missiles are thrown/shot with +1 added to the Missile Modifier.}
%Preserver
\incantentry{Brew}{Mixture/generic}{1200}{100/rank}{4 Hours / Rank}{}{This collection of various brews alleviate and fortify the body against a number of different effects. The incanter mixes one ounce of the proper ingredient with the fluid mixed from some common herbs. The target drinks the brew and is protected from some}
%Preserver
\incantentry{Courage Draught}{Mixture/Triggered}{600}{30/rank}{1 Hour / Rank}{Liquor aged at
least 100 years}{This mixture endows the user with greater courage to face any task. Fear spells and other emotion changing magic have no effect on the target. Also all INT checks are made at one die higher than normal. The effect lasts for one house per rank of the mixtu}
%Preserver
\incantentry{Corpse Ward}{Talisman}{500}{40/rank}{3 Hours / Rank}{Symbolic object}{Through this ritual the nomad focuses spiritual power into a symbolic object. This object becomes an anathema to the undead. When brandished at an undead, it must make a WIL check equal to the rank of the ritual or flee from the object. If the check fails}
%Preserver
\incantentry{Deep Sleep}{Invocation}{1000}{200/rank}{3 Hours / Rank}{
Rank Time
1 6 months
2 2 years
3 8 years
4 24 years
5 100 years
6 500 years}{The nomad uses this to place himself into a deep, coma-like sleep from which he cannot be awakened. While beginning the ritual, he decides how long he should sleep, and then slowly enters a trance. In this state he will need neither food nor drink, and ve}
%Preserver
\incantentry{Descry Ships}{Talisman}{500}{50/rank}{2 Hours / Rank}{}{The nomad first inscribes a map of the sea and the islands about the ship on parchment. The map must be fairly accurate for this to work. Once prepared, the nomad can use the map any time his actual location is still within the confines of the map. Touchi}
%Preserver
\incantentry{Dram of Energy}{Mixture/Triggered}{600}{50/rank}{2 Hours / Rank}{}{This mixture alleviates the fatigue of the user and increases his PER by 4 for a house per rank of the mixture. He also becomes agitated and more argumentative. All CSE checks are at one die higher than normal. Repeated use of this mixture may cause depen}
%Preserver
\incantentry{Elixir of Health}{Mixture/Triggered}{600}{45/rank}{2 Hours / Rank}{}{This general restorative can soften the blow of many scourges. The nomad administers the potion, and the target rolls a check of 6 minus the mixture rank against health. If successful, the target's condition improves. If administered within 3 bells of con}
%Preserver
\incantentry{Evaluator}{Talisman/Permanent}{1000}{100/rank}{4 Hours / Rank}{Monocle}{Once created, looking through this monocle reveals imperfections, spoilage, and damage in vegetable and other food materials as a glowing red area. Each rank above one magnifies the image by two. Making smaller and smaller imperfections visible, as well a}
%Preserver
\incantentry{Find Weakness}{Invocation}{200}{10/rank}{1 Minute / Rank}{}{The nomad places his hands on an object, or area, made of wood or stone. He may then find the weakest spot (a patch, material imperfection, etc.) in a square area 10 feet per rank of the ritual on a side. If no such weakness exists, the ritual will reveal}
%Preserver
\incantentry{Future Scry}{Invocation}{200}{10/rank}{1 Minute / Rank}{@i(special)}{When a nomad uses the ritual for the first time, he must declare the method of scrying that he will always perform. Some methods are looking into a crystal ball, read tea leaf patterns, interpreting forest sounds, reading cards, reading the entrails of ch}
%Preserver
\incantentry{Hangover Liquor}{Mixture/generic}{200}{20/rank}{1/2 Hours / Rank}{}{This powerful, but deceptive, liquid causes effects similar to alcohol. The nomad administers the potion, and the target makes a check of the mixture rank against HEA. If he fails, he is inflicted (several hours later) with a severe headache, dizziness, a}
%Preserver
\incantentry{Healing Sleep}{Invocation}{600}{25/rank}{2 Hours / Rank}{Vinegar}{This ritual lulls a willing target into a deep sleep where they feel no pain. The production time of the ritual begins as the target falls asleep and must continue to the needed amount of time for the rank of the ritual. The healing rate of the target is }
%Preserver
\incantentry{Infertility Lotion}{Mixture/generic}{400}{40/rank}{2  Hours / Rank}{4 drops virgin's
blood given freely}{This lotion is applied to a woman's abdomen to prevent her from becoming pregnant by normal means. This mixture continues its effect for one month per rank of the mixture.}
%Preserver
\incantentry{Love Potion}{Mixture/targeted}{800}{40/rank}{2 Hours / Rank}{}{This brew directly effects the drinker's emotions. After the nomad administers the potion, the imbiber must make a check against WIL of one die per rank of the mixture. If the target fails, he will fall madly in love with the next person he sees. This sta}
%Preserver
\incantentry{Luck Ritual}{Invocation}{600}{20/rank}{1 Hour / Rank}{}{Through contact with the spirits a nomad may improve the flow of subtle forces that surround him and the nearby world. The nomad may choose to alter a die roll by a number of pips equal to the rank of the ritual used. A nomad may only have one active ritu}
%Preserver
\incantentry{Mask of Disclosure}{Talisman}{600}{60/rank}{3 Hours / Rank}{}{Rather than hide its wearer, this mask reveals them. The nomad places this leather mask on the target and fastens the straps to hold it in place. If that target is possessed, or has multiple personalities, a check against WIL of the ritual rank versus the}
%Preserver
\incantentry{Powder of Preservation}{Mixture/generic}{100}{10/rank}{1/2 Hours / Rank}{}{This powder treats 100 pounds of food to keep it fresh and tasty for two weeks times the rank of the mixture. At the end of that time the food immediately goes bad. Multiple applications have no effect.}
%Preserver
\incantentry{Poultice}{Mixture/generic}{400}{25/rank}{2 Hours / Rank}{}{This bandage wrap is specially made to treat wounds quickly. When applied this bandage stops bleeding from the wound and 10 minutes after application restores 1 DP per rank of the bandage. Also for every 2 ranks of the bandage 1 die is taken off a roll to}
%Preserver
\incantentry{Power Patch}{Mixture/generic}{400}{40/rank}{1 Hour / Rank}{}{A mixture of various components is applied to a cloth patch and applied to the target. As applied it increases the rank of the target's Exceptional PWR by 1 (to 1 if none previous). This effect lasts a house per rank of the mixture.}
%Preserver
\incantentry{Remedy}{Mixture/targeted}{800}{75/rank}{3 Hours / Rank}{
1 Putrihaut          powdered terik root
2 Maldormi           soki herb
3 Granjuke           fresh charn meat
4 Ste}{When used this mixture relieves the symptoms of the diseases it affects within 1d6 Bells and removes the disease entirely within 1d4 days.}
%Preserver
\incantentry{Sago Weed}{Mixture/Triggered}{600}{50/rank}{2 Hours / Rank}{}{This is a special blend of rare herbs, a spice called Grofelt, and fine tobacco. Once prepared, the nomad can invite up to 3 of his friends to share in smoking the mixture. This will take up at least 3 houses during which time the smoke penetrates the smo}
%Preserver
\incantentry{Ship's Veil}{Invocation}{1500}{100/rank}{3 Hours / Rank}{Gharton silk}{While aboard the target ship the nomad weaves a cloth (a 1d6 check versus the Weaving Skill to succeed) and uses this ritual to create a veil. A first rank veil renders the ship invisible for 10 minutes. Each additional rank doubles the amount of time the}
%Preserver
\incantentry{Shore Object}{Invocation}{300}{15/rank}{1/2 Hours / Rank}{}{Any wooden object with a patch or weakened spot is made whole again through this ritual. The area of an object affected is up to a cube 10 feet per ritual rank on a side.}
%Preserver
\incantentry{Stimulant Chew}{Mixture/generic}{400}{20/rank}{1 Hour / Rank}{}{This minty flavored chewing substance makes the target less fatigued and less drowsy than before. For every rank of the mixture the target is able to stay awake for another 3 houses. After the substance's effect is expired the target must make a check equ}
%Preserver
\incantentry{Strength Patch}{Mixture/generic}{400}{40/rank}{1 Hour / Rank}{}{A mixture of various components is applied to a cloth patch and applied to the target. As applied it increases the ranks of the target's Exceptional STR by 1 (to 1 if none previous). This effect lasts a house per rank of the mixture.}
%Preserver
\incantentry{Suppress Pain}{Invocation}{200}{10/rank}{5 Minutes / Rank}{Thin iron needles}{The nomad inserts thin iron needles in the body of the sufferer (himself possibly), into major nerves leading from an injured body area. He then uses the needles to channel his ritual to the nerve, ordering it to stop transmitting pain to the sufferer's b}
%Preserver
\incantentry{Tonic}{Mixture/generic}{400}{30/rank}{2 Hours / Rank}{}{This collection of various tonics alleviate and fortify the body against a number of different effects. The incanter mixes 1/2 of an ounce of the proper ingredient with the fluid mixed from some common herbs. The target drinks the tonic and is protected f}
%Preserver
\incantentry{Will Patch}{Mixture/generic}{400}{40/rank}{1 Hour / Rank}{}{A mixture of various components is applied to a cloth patch and applied to the target. As applied it increases the ranks of the target's Exceptional WIL by 1 (to 1 if none previous). This effect lasts a house per rank of the mixture.}

\section{Troubador Incants}

%Troubador
\incantentry{Animal Dance}{Song/play+dance}{200}{0}{1 Minute / Rank}{}{When the nomad begins to play an instrument and dance, all harmless animals within hearing must seek out the singer, and then form a line behind him and dance as he does. The creatures can make a check against WIL at one die per rank of the song.}
%Troubador
\incantentry{Baton of Dancing}{Talisman}{2000}{50/house}{2 Hours / Rank}{Saltpeter
oil
wooden staff}{A nomad spreads a saltpeter infused oil on any strait wooden stick of at least 2 feet in length. This ritual changes the shaft of wood into a baton. With the baton in hand, the nomad brandishes it before his target(s) and begins to dance. If the target(s)}
%Troubador
\incantentry{Calling of the One}{Song/sing}{300}{0}{2 Hours / Rank}{}{The nomad begins his song, naming one person he knows reasonably well (GM discretion). If during the course of the song the named person hears the music from whatever distance, and fails a check against WIL of one die per rank of the song, that person mus}
%Troubador
\incantentry{Co-Ordination Jig}{Song/sing or play}{400}{0}{2 Minutes / Rank}{}{While the nomad sings or plays this song, all his allies within earshot temporarily decrease all AGI checks by one die per rank of the dance.}
%Troubador
\incantentry{Dance of Distraction}{Song/dance}{600}{0}{2 Minutes / Rank}{}{While this dance is being observed, those not warned beforehand lose 1d6 per rank of the song on any check against PER. The nomad will often use this to distract opponents from noticing what his allies are up to.}
%Troubador
\incantentry{Dance of the Lights}{Song/dance}{200}{0}{2 Hours / Rank}{}{The nomad uses this song to attract the attention of fire spirits. Shortly after he starts dancing, they swarm about him, casting a flickering light into even the darkest places. They cannot, however overcome magical darkness. At rank one they are equival}
%Troubador
\incantentry{Erasure Song}{Song/sing}{200}{0}{2 Minutes / Rank}{}{The nomad starts singing this song, singing its lyrics about some event or piece of knowledge. Everyone within hearing whomis not his ally must make a check against WIL of one die per rank of the song or forget the subject of the song for at least 24 hour}
%Troubador
\incantentry{Heart of Courage}{Song/sing}{300}{0}{10 Minutes / Rank}{}{While singing, all resistance checks made by the singer's allies against WIL are made at one die less for each rank of the song.}
%Troubador
\incantentry{Lute Lamp}{Talisman/premament}{1000}{30/rank}{2 Hours / Rank}{}{A lute enchanted with this spiritual magic creates an area of 10 feet per rank of the talisman in faint shimmering and sparkling light about the nomad when in darkness or moonlight. This light is only equal to one half the light of a torch, but its shimme}
%Troubador
\incantentry{Message Song}{Song/sing}{100}{0}{1 Minute / Rank}{}{This song targets one person per rank of the song. Each of these people must be within earshot and well known to the nomad. The nomad while singing the nomad sends a message subtly using the song as the carrier. The message is of the nomad's choosing and }
%Troubador
\incantentry{Mood Ring}{Talisman/Permanent}{600}{30/rank}{2 Hours / Rank}{Silver ring}{A nomad creates this item by procuring a silver ring and assembling himself and any musician friends in a quiet place, where they will not be disturbed. The group begins to play, as the nomad begins to concentrate on the ring. Any instrumentation will wor}
%Troubador
\incantentry{Pacifier Pipe}{Talisman/Permanent}{2000}{200/rank}{3 Hours / Rank}{Pipe
grofelt weed}{A pipe or recorder made into this talisman is very soothing to hear. Those within hearing of it being played must make a WIL check at one die per rank of the talisman each round they attempt to be violent, if they fail, they do nothing. This effects all w}
%Troubador
\incantentry{Robes of Style}{Talisman/Permanent}{5000}{300/rank}{3 Hours / Rank}{Robes
colored dyes
buttons
cloth ties}{Taking a newly spun and sewn robe, the nomad endows it with the ability to change shape, size, and color to match his tastes and desires. This robe can change appearance once per day for each rank of the talisman.}
%Troubador
\incantentry{Rolling Hills}{Sont/sing or play}{300}{0}{2 Hours / Rank}{}{The sound of this song lulls earth spirits into making the nomad's and his allies' passage easier. Adventurers may run on solid ground for three houses per rank of this song without tiring. Those affected by this song cannot be effected again until a full}
%Troubador
\incantentry{Shadow Cream}{Mixture/targeted}{300}{40/rank}{1 Hour / Rank}{Target sample}{The nomad spreads this cream over the target's face to change the target's appearance to what the nomad desires. Others must make a PER check of the mixture rank plus one to see through this disguise. The effect lasts until the target sleeps, washes, gets}
%Troubador
\incantentry{Sleep Awaits}{Song/sing or play}{300}{0}{1/2 Hours / Rank}{}{The nomad's opponents and willing allies hearing the song must, sometime during its duration, make a check against WIL for a number of dice equal to the rank of the song. If they fail, they fail into a deep, but natural sleep. The music is soft, sublimina}
%Troubador
\incantentry{Spirit Dirge}{Song/play}{400}{0}{1/2 Hours / Rank}{}{The nomad plays this song for a dying companion. The song assures the spirit's path to the Kurago is clear, and makes the Kurago ready to receive the spirit. Pain and agony are partially alleviated, and the passing becomes as peaceful as possible.}
%Troubador
\incantentry{Still Waters}{Song/sing}{300}{0}{2 Hours / Rank}{}{The singer lulls water spirits into calming the waters about his boat for a radius of about 100 feet. Each rank of the song calms the waters surface by one step in this sequence: smooth water gentle waves (under 1') choppy water (1-2') heavy waves (3-6') }
%Troubador
\incantentry{Stone Song}{Song/sing}{300}{0}{5 Minutes / Rank}{}{The nomad weaves the name of one person into this song. If the target is within earshot, all checks against his STR are at 1d6 less per rank of the song.}
%Troubador
\incantentry{Synchro Dance}{Song/play+dance}{100}{0}{1R/rank}{}{The nomad plays a four count, and then all willing people within earshot dance as he does, move for move, for up to one song of up to about 10 minutes. Each rank above the first doubles the perfection of the dance, and adds 1d6 to any reaction checks to t}
%Troubador
\incantentry{The Howling}{Invocation}{600}{0}{4 Hours / Rank}{}{Drawing in a large breath and thinking about a single word or scream, the nomad uses energy from Kurago to multiply the effect of his shout. The word or animal scream he utters can be heard for a distance of up to 1 met per rank of the invocation.}
%Troubador
\incantentry{The Mocking}{Song/sing}{300}{0}{10 Minutes / Rank}{}{The nomad sings about a particular person, place, or object. All the listeners' reactions to the target that would be rolled against COM are done so at 1d6 more per rank of the song. The GM will establish an appropriate COM for places of objects.}
%Troubador
\incantentry{The Praising}{Song/sing}{300}{0}{10 Minutes / Rank}{}{The nomad sings about a particular person, place, or object. All the listeners' reactions to the target that would be rolled against its COM are done so at 1d6 less per rank of this song. The GM will establish an appropriate COM for places or objects.}
%Troubador
\incantentry{The Telling}{Song/sing+play}{10000}{0}{1 Day / Rank}{}{This song is very precious to the nomad. While playing his instrument and singing the nomad allows the listeners to experience what he sings about. Through the music and vocalization the story is crafted. At rank one, the listeners can hear any sound, no }
%Troubador
\incantentry{Wary of the Night}{Song/sing}{200}{0}{10M}{}{When the nomad's allies are camped in the dark (at night, below ground, etc.), this song focuses their perceptions, decreesing all PER checks bu one die. This effect lasts for 2 houses for each rank of the incant once the song has finished. A group can on}
%Troubador
\incantentry{Yes Song}{Song/sing}{400}{0}{2 Minutes / Rank}{}{The nomad starts to sing this song, singing its lyrics about some non-life threatening action he wishes the audience to perform. Everyone within hearing whom is not his ally must make a check against WIL of a number of dice equal to the rank of the song o}
%Troubador
\incantentry{Zephyr Polka}{Song/play}{300}{0}{2 Hours / Rank}{}{While this song is being played, air spirits lulled by its tune cause a breeze to blow that can increase a sailing ship's speed by five mets per rank of the song.}

\section{Seeker Incants}

%Seeker
\incantentry{Acid Draught}{Mixture/Triggered}{600}{50/rank}{4H/rank}{Akvovervo water}{This preventative absorbs damage done by acid. The potion absorbs 5 points of damage per rank of the potion. This potion remains active in the system of the imbiber for 1 days per rank.}
%Seeker
\incantentry{Animal Bind}{Mixture/targeted}{800}{50/rank}{2 Hours / Rank}{Animal sample}{The nomad needs something from an animal to prepare this mixture, some hair, saliva, nail clippings or discarded skin. The animal must not be injured in the proces. The incanter mixes this with the other ingredients while performing the incant ritual. Thi}
%Seeker
\incantentry{Anklet of Dexterity}{Talisman/Permanent}{3000}{200/rank}{4 Hours / Rank}{}{To produce this talisman the nomad infuses an anklet to provide him with better coordination and reflexes. The anklet gives the nomad a bonus to his AGI equal to the rank of the ritual used to create it. Only one of these anklets may affect one nomad at a}
%Seeker
\incantentry{Armband of Fortitude}{Talisman/Permanent}{3000}{300/rank}{4 Hours / Rank}{Copper armband
powdered fungigo}{To produce this talisman the nomad infuses an armband to provide him with greater stamina and hardiness. The armband gives the nomad a bonus to his HEA equal to the rank of the ritual used to create it. Only one of these armbands may affect one nomad at a}
%Seeker
\incantentry{Assimilation}{Invocation}{200}{10/rank}{2 Hours / Rank}{}{A nomad need not feel out of place after using the invocation, because he is able to read subtle physical and mental clues from the strangers about him and adapt to their culture. At rank one the nomad only is affected, though at higher ranks the number o}
%Seeker
\incantentry{Cinnamon Bane}{Mixture/Triggered}{1200}{25/rank}{2 Hours / Rank}{1lb cinnamon}{This bane deters attacks by hirudo against the target. The incanter spreads the mixture on a person or object, and for 1 house all resistance checks against any effects created by a hirudo are at one die less.}
%Seeker
\incantentry{Ethereal Bane}{Mixture/Triggered}{1600}{80/rank}{2 Hours / Rank}{}{This bane deters attacks by ethereal creatures against the target. The incanter spreads the mixture on a person or object, and for 1 house all resistance checks against any effects created by ethereal creatures are at one die less and all attacks have a p}
%Seeker
\incantentry{Crystal Gate}{Talisman/Permanent}{10000}{1000/rank}{12 Hours / Rank}{Ground quartz
1/2 ounce of
Terisium/rank}{The nomad, with the assistance of a glassblower if he lacks the skill, creates two glass spheres from the same batch of molten glass. Performing a ritual over them while they cool, links the two spheres to work as a two-way audio-visual gate that can be u}
%Seeker
\incantentry{Ferret Object}{Invocation}{200}{5/rank}{10 Minutes / Rank}{}{The nomad touches a person, who then concentrates on an item he has lost within the last 6 days. If the visualization is clear and the item is within 100 feet per rank of the invocation, the nomad hears a sound which will show him the direction and distan}
%Seeker
\incantentry{Fertility Seed}{Mixture/generic}{300}{5/rank}{1 Hour / Rank}{[add ing.]}{This powder, when spread over one acre of farm land, increases crop yield by 5\% for each incant rank.}
%Seeker
\incantentry{Fire Shield}{Mixture/Triggered}{600}{50/rank}{4 Hours / Rank}{Gharton egg shell}{This potion absorbs damage done by fire that the imbiber takes after effects are resolved. The potion absorbs 5 points of damage per rank of the potion. This potion remains in the system of the imbiber for 2 days per rank. The effects of this potion begin}
%Seeker
\incantentry{Gargoyle Mark}{Imprinting}{2000}{200/rank}{4 Hours / Rank}{Powdered/ashed
garggoyle}{The tattoo created with this ritual improves the incanter's muscular strength. He is treated as having a STR at the number of ranks of the ritual higher. Only one of these tattoos may affect one nomad at a time.}
%Seeker
\incantentry{Immaterial Bane}{Mixture/Triggered}{1000}{22/rank}{2 Hours / Rank}{Powdered bones
of a dead person}{This bane deters attacks by spirits from the Kurago against the target. The incanter spreads the mixture on a person or object, and for 1 house all resistance checks against any effects created by a spirits are at one die less and all attacks have a penal}
%Seeker
\incantentry{Insight Gem}{Talisman/Permanent}{3000}{200/rank}{4 Hours / Rank}{2 ounces silver
small gem}{This gem, mounted in a gold earing, focuses the sensory abilities of the wearer. This gives the wearer a bonus to his PER equal to the rank of the ritual used to create it. Only one of these earrings may affect one person at a time.}
%Seeker
\incantentry{Iron Tonic}{Mixture/trigger4ed}{600}{50/rank}{4 Hours / Rank}{Powdered iron
sea scorpion or
draco scales}{This tonic absorbs damage done by physical attacks to the imbiber. The potion absorbs 5 times the incant rank of points. This potion remains in the system of the imbiber for 2 days per rank.}
%Seeker
\incantentry{Moon Boots}{Talisman/permement}{10000}{100/rank}{2 Hours / Rank}{Boots
1 ounce Terisium}{Starting with a well fitting pair of boots, the nomad displays these under the light of one of the moons and creates this talisman. When done, while he wears the boots, and is in the light of the moon they were created under, he is 10\% lighter for each ra}
%Seeker
\incantentry{Moon Crystal}{Talisman/permement}{1000}{70/rank}{1 Hour / Rank}{Quartz}{Using a perfect quartz crystal of at least 2 inches in size, the incanter prepares this talisman by the light of one of the moons. When complete, he can then use the crystal to mezmorize / hypnotize willing subjects. He cannot force information from them,}
%Seeker
\incantentry{Moon Shroud}{Talisman/permement}{1000}{80/rank}{1 Hour / Rank}{Silver necklass}{The incanter uses a well crafter silver crescent, at least 2 inches in size, hung on a silver necklace. Taking this out into the light of on of the moons, he falls into a trance concentrating on the moon above and weaving power from the Kurago into the ta}
%Seeker
\incantentry{Moon Sleep}{Talisman/Permanent}{800}{50/rank}{2 Hours / Rank}{Silver
diamond dust
glass vial}{This vial is worn on a silver chain around the neck of the incanter. When it is lit by the light of its moon under which it was created, the incanter may roll a WIL check of 7 minus the rank of the talisman in dice. If successful the incanter goes into a }
%Seeker
\incantentry{Shock Buffer}{Mixture/Triggered}{600}{50/rank}{4 Hours / Rank}{Gargoyle hide}{This potion absorbs damage done by electricity that the imbiber takes after effects are resolved. The potion absorbs 5 points of damage per rank of the potion. This potion remains in the system of the imbiber for 2 days per rank. The effects of this potio}
%Seeker
\incantentry{Silver Bane}{Mixture/Triggered}{600}{15/rank}{2 Hours / Rank}{2lb silver dust}{This bane deters attacks by lycanthrops against the target. The incanter spreads the mixture on a person or object, and for 1 house all resistance checks against any effects created by a lycanthrope are at one die less. Any attacks on the target by a lyca}
%Seeker
\incantentry{Spirit Bane}{Mixture/Triggered}{800}{20/rank}{2 Hours / Rank}{Earth from a grave}{This bane deters attacks by ghosts and immaterial creatures against the target. The incanter spreads the mixture on a person or object, and for 1 house all resistance checks against any effects created by a ghost or immaterial creature is at one die less }
%Seeker
\incantentry{Soul Sight}{Imprinting}{15000}{400/rank}{1 Day / Rank}{}{This imprinting is tattooed on the eyelids of the incanter. When he closes his eyes he can see any immaterial spirits that are within 10 feet times the rank of the imprinting. Ghosts, spirits of the Kurago, and elemental spirits can be seen, but not invis}
%Seeker
\incantentry{Spirit Exchange}{Mixture/targeted}{1000}{100/rank}{2 Hours / Rank}{}{The propriety of this mixture had been debated among nomadic rondos for some time. It allows the nomad to exchange his spirit with the spirit of another living creature. Effectively, the nomad and the other exchange bodies. All mental attributes and memor}
%Seeker
\incantentry{Spirit Guise}{Talisman/Permanent}{500}{40/rank}{3 Hours / Rank}{Sample of dead
spirit}{Acquiring a piece of personal jewelry from one whom is dead, the incanter must permanently affix the jewelry to his own body. Concentrating on the visage of the dead target the nomad enters a trance and attempts to draw the dead spirit from the Kurago. If}
%Seeker
\incantentry{Spirit Scout}{Imprinting}{1000}{100/rank}{4 Hours / Rank}{}{The incanter prepares a mixture to put himself into a trance, focusing his mind on attracting the spirit of a lower animal. The tattooist then designs a representation of that animal onto the chosen section of the nomad's skin. The tattoo must be at least}
%Seeker
\incantentry{Spirit Trance}{Invocation}{100}{20/rank}{2 Hours / Rank}{}{This invocation allows the nomad's spirit to enter the Kurago and seek direct contact with its spirits. The nomad uses a number of potions, ointments, and a trance-inducing ceremony. Upon entering the Kurago, the nomad mentally pictures the spirit he seek}
%Seeker
\incantentry{Spirit View}{Mixture/targeted}{250}{20/rank}{2 Hours / Rank}{}{This mixture allows the inbiber to view the world through another living creature's senses. The incanter prepares the mixture with a sample of the creatures hair, feathers, blood, etc. After anointing himself with it he places himself in a trance, and may}
%Seeker
\incantentry{Tamener}{Talasiman/Permanent}{3000}{200/rank}{4 Hours / Rank}{}{A tamenwrap is a strip of cloth used by a nomad to keep his hair out of his eyes. With this talisman, the incanter gives a tamenwrap the ability to enhance his own INT by one point per rank of the ritual when the tamenwrap is worn.}
%Seeker
\incantentry{Thermal Gel}{Mixture/Triggered}{600}{50/rank}{4 Hours / Rank}{Jacer ooze}{The incanter spreads this cream on the target to activate it. It then absorbs 5 points of cold damage per rank of the incant. This remains active for 2 days per rank or intil it is washed off.}
%Seeker
\incantentry{Time Shifter}{Talisman/Permanent}{500}{300/silver}{3 Hours / Rank}{}{The incanter creates a connection to the Kurago to a small silver or glass mirror. The spirits there watch events visible from the mirror. Once per day the incanter can then touch the mirror and concentrate on any event occuring within 1 bell for each ran}
%Seeker
\incantentry{Tongue Ring}{Talisman/Permanent}{500}{100/rank}{1 Hour / Rank}{Target jewelry}{Possessing a small ring or earring from a person who is no longer living, the incanter prepares this talisman by piercing his tongue and permanently inserting the jewelry. Incanting the ritual, he creates a conduit to the spirit in whatever afterlife it s}
%Seeker
\incantentry{Underwater Gel}{Mixture/targeted}{800}{100/rank}{3 Hours / Rank}{Octopod blood}{This substance when applied to the nose and mouth allows the target to breathe underwater for half a house per rank of the gel.}

\section{Second Lifer Incants}

%Second Lifer
\incantentry{Addiction Potion}{Mixture/targeted}{1000}{40/rank}{3 Hours / Rank}{}{When someone drinks this potion mixed with another fluid, he becomes addicted to the other fluid. At the end of each full day he is without the fluid he must attempt a check against HEA of the number of dice of the mixture rank. If successful the addictio}
%Second Lifer
\incantentry{Amulet of Deception}{Talisman}{250}{50/rank}{1 Hour / Rank}{Locket
target's hair
drawing of target's
loved one}{With a small, openable amulet, the incanter uses a lock of hair from the intended target, and a drawing of someone to which the victim is emotionally attached to create this talisman. When the target sees the incanter wearing the unopened amulet, and he f}
%Second Lifer
\incantentry{Arid Seed}{Mixture/Triggered}{600}{40/rank}{1 Hour / Rank}{[add ing.]}{This powder is spread over one acre of farm land. The crop yield is decreased by 5\% per rank of the mixture.}
%Second Lifer
\incantentry{Arthritis Potion}{Mixture/Triggered}{800}{50/rank}{1 Hour / Rank}{}{This potion induces an illness of the joints making movement painful. If the drinker fails a check against HEA equal to the rank of the mixture, he loses 1d4 AGI points permanently.}
%Second Lifer
\incantentry{Bond Pain}{Mixture/Triggered}{800}{50/rank}{1 Hour / Rank}{Target sample}{The incanter must possess a small bit of his intended target's body, (a lock of hair, a nail clipping, a scrap of skin, etc.). He incorporates this object into an effigy of the target while infusing it with energy. At the appropriate time, he holds the ef}
%Second Lifer
\incantentry{Brain Fevers Draught}{Mixture/Triggered}{1000}{50/rank}{1 Hour / Rank}{}{This induces a fever that damages the brain. If the target fails a check against HEA equal to the rank of the mixture, he loses 1d4 INT points permanently. The incapacitating fever lasts from three to six days.}
%Second Lifer
\incantentry{Dose of the Itch}{Mixture/Triggered}{700}{50/rank}{1 Hour / Rank}{}{The only symptom of The Itch is a very aggravating full body rash. If the target fails a check against HEA equal to the rank of the mixture, he breaks out in an irritating, painful rash. The target's AGI and PER are lowered by 1d6 points for 1d6 days.}
%Second Lifer
\incantentry{Grippe Dose}{Mixture/Triggered}{700}{50/rank}{1 Hour / Rank}{}{This potion induces an illness which is temporarily incapacitating. The incanter administers the potion, and the target must make a check against HEA equal to the rank of the mixture. If he fails, he is incapable of performing any actions, except the easi}
%Second Lifer
\incantentry{Fear Draught}{Mixture/generic}{400}{30/rank}{1 Hour / Rank}{}{This liquid attacks the brain and makes its imbiber fear all unknown people and objects. Every time the imbiber meets a person, or is introduced to a new item, he must attempt a check against HEA equal to the rank of the mixture. If he fails he is uncontr}
%Second Lifer
\incantentry{Leprosy Tonic}{Mixture/Triggered}{800}{50/rank}{3 Hours / Rank}{Corpse mold}{This tonic induces a degenerative disease which will alienate its victim from society. After the incanter administers this tonic, the imbiber must make a check against HEA equal to the rank of the mixture. If he fails, his body begins to slowly deteriorat}
%Second Lifer
\incantentry{Mark of the Lamprey}{Imprinting}{3000}{300/rank}{4 Hours / Rank}{Target's blood}{With the aid of a skilled tattooist a incanter can create this imprinting. He prepares a special ink for the tattoo that includes the blood on another living person. The tattooist creates a tattoo in the shape of a lamprey, leech, or other parasitic creat}
%Second Lifer
\incantentry{Mark of Homing}{Imprinting}{1500}{75/rank}{3 Hours / Rank}{}{The incanter paints a symbol onto the target of this imprinting. The mark quickly fades from view of all, except the nomad who can still see it clearly. At anytime the nomad may know the direction and distance to the person he marked, if the target fails }
%Second Lifer
\incantentry{Ointment of Corruption}{Mixture/Triggered}{600}{40/rank}{3 Hours / Rank}{[add igr.]}{Crated in a crystal vial, this fluid can eat through lead, bronze, or rock up to one foot thick per rank of the mixture. On contact with the air (the first time the vial is opened) it corrupts any material in contact with it except steel and adamantite. I}
%Second Lifer
\incantentry{Pain Smoke}{Mixture/Triggered}{600}{50/rank}{2 Hours / Rank}{}{This mixture cause those who inhale this smoke to make a check against HEA equal to 2 plus half the mixture rank rounded up. Those who fail this check are afflicting with a wracking pain in the lower abdomen and a fierce headache. This causes the afflicte}
%Second Lifer
\incantentry{Plague Dram}{Mixture/Triggered}{1800}{70/rank}{4 Hours / Rank}{}{When the incanter administers this potion, the target must make a check against HEA equal to the rank of the mixture. If he fails, he contracts the deadly black plague. Once infected, the target loses 1d6 HEA permanently each day for the next three days. }
%Second Lifer
\incantentry{Powder of Phantasm}{Mixture/Triggered}{300}{20/rank}{1 Hour / Rank}{}{This powder is sprinkled over a circular area of a diameter 10 feet times the rank of the mixture. At the time the powder is used, the incanter focuses on an alternative appearance for the area. His static vision becomes visible to all looking at the area}
%Second Lifer
\incantentry{Ring of Likeness}{Talisman}{500}{60/rank}{2 Hours / Rank}{}{The incanter visualizes a particular visage and dress as he prepares this talisman. The ring when worn makes the wearer appear as what the incanter envisioned. All viewing the wearer may make a check against PER equal to the rank of the talisman to break }
%Second Lifer
\incantentry{Ring of Subjugation}{Talisman/Permanent}{2000}{200/rank}{4 Hours / Rank}{Gold or silver ring}{This talisman is created around a piece of gold or silver jewelry. When the jewelry is pierced through the target's skin, it fuses solid and can not be harmed or removed. Anytime the nomad directs a verbal command at the target, he must perform it, unless}
%Second Lifer
\incantentry{Scourge}{Invocation}{1400}{70/rank}{4 Hours / Rank}{Target's item}{After preparing a boiling mixture, the incanter adds a personal item from his intended target (some hair, a piece of jewelry, etc.) and repeats the target's name three times. Wherever he is, the victim experiences extreme, crippling pain for one minute. H}
%Second Lifer
\incantentry{Sleeping Sickness Solution}{Mixture/Triggered}{600}{50/rank}{2 Hours / Rank}{}{This sickness makes the imbiber sleep most of the time. The nomad administers the potion, and the victim makes a check against HEA equal to the rank of the mixture. If he fails, he is inflicted with sleeping sickness. The target always feels drained of en}
%Second Lifer
\incantentry{Spirits of Consumption}{Mixture/Triggered}{1600}{80/rank}{4 Hours / Rank}{}{This illness of the lungs causes chronic coughing and a general degradation of health. The nomad administers the potion and the victim makes a check against HEA equal to the rank of the mixture. If he fails he acquires this eventually fatal disease. He pe}
%Second Lifer
\incantentry{Vial of Ill Omen}{Mixture/targeted}{1000}{60/rank}{2 Hours / Rank}{Victim's blood
corpse's skin
basil}{The incanter combines a small amount of blood of his intended target with skin from a corpse and ground basil in a small vial. The nomad then infuses the mixture with the powers of bad luck. After the ritual, the target is struck with ill omen. In any sit}