\chapter{Air Magic}
\label{ch:elemental-air}

Magicians devoted to the Air element spend much of their time out of doors where they may feel the power of the wind and the subtleties of light breezes. Some dismiss Air magicians as flighty and fickle, but they are the most cerebral of the elemental magicians.

Air Magic typifies the power of the element, in spells like Lightning Bolt and Anaerobia, and the ability to manipulate the element.

Vision spells enhance the seeing ability of the magician.

Scrying spells are information-gathering magics.

Dimensions involve summoning or sending items to or away from the caster. Because of the great power that must be channeled to use these spells, they are more difficult to learn, and draining to the magician. The base costs for these spells is 600 EP.

\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{tabular}{@{} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth}}
\begin{enumerate}
	\item Air Magic
	\begin{enumerate}
		\item Minibolt
		\item Gust
		\item Arise
		\item Fresh Air
		\item Ironlungs
		\item Lightning Bolt
		\item Fly
		\item Wind
		\item Air Temperature
		\item Anaerobia
		\item Hawksoar
		\item Ekblovego
		\item Vacuum
		\item Blitzen
		\item Aerovoki
		\item Wrath
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{1}
	\item Vision
	\begin{enumerate}
		\item Murk
		\item Distortion
		\item Crystal
		\item Glamor
		\item Intuit
		\item One Eye
		\item Discovery
		\item Parallax
		\item Long Eyes
		\item Heat Vision
		\item Night Sight
		\item Invisibility
		\item Sight License
		\item Blind
		\item Projection
		\item See All
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{2}
	\item Scrying
	\begin{enumerate}
		\item Detect Motivation
		\item Hear
		\item Read Emotions
		\item Mask Motivation
		\item See
		\item Witchsmeller
		\item Know True Motivation
		\item Locate
		\item Senses
		\item Truthtell
		\item Futuresee
		\item Legends
		\item Thingtell
		\item Backfire
		\item Scan
		\item Godspeak
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{3}
	\item Dimensions
	\begin{enumerate}
		\item Shimmer
		\item Thought Move
		\item Send Object
		\item Peer
		\item Magic Fence
		\item Call Object
		\item Hole
		\item Giant Stride
		\item Send Package
		\item Pocket
		\item Portable Hole
		\item Call Package
		\item Magic Door
		\item Vanish
		\item Otherworld
		\item Banish
	\end{enumerate}
\end{enumerate} \\
\end{tabular}
\pagebreak
\section{Air Magic}

\spellentry{Minibolt}{1}{1 Round}{None}{LOS 60 + 15/Finesse Feet}{Instantaneous}{1 object or creature}{1d6 damage}{Sends a small electrical charge from the air blasting towards the target. A Minibolt always hits the target, but magical resistance can negate or reduce the damage (GM's discretion).}

\spellentry{Gust}{2}{1 Round}{None}{60 + 10/Finesse Feet}{1 Minute}{5 x (20 + 10/Finesse)' line}{30 + 10/Finesse mph gust}{Produces a blast of air in the area of effect sufficient to extinguish unprotected flames, fan a large fire, etc.}

\spellentry{Arise}{3}{1 Round}{None}{Caster}{2d6 + 1/Finesse Rounds}{Caster}{Accelerate 4 + 1/Finesse Feet /Rounds}{Allows the caster to float into the air like a kite. While concentrating, the caster can accelerate up or down, but is at the mercy of winds and other forces that could cause lateral movement. The caster cannot carry more than they can normally lift above their head.}

\spellentry{Fresh Air}{4}{1 Round}{None}{0}{1d4 + 1/Finesse Rounds}{10 + 5/Finesse Feet Radius}{Freshen air}{Cleanse a volume of fetid, putrid, or poisonous air to make it breathable. The purified area can move with the caster if they so choose.}

\spellentry{Ironlungs}{5}{1 Round}{4d6 vs PWR negates}{Touch}{10 + 3/Finesse Minutes}{1 creature}{Hold breath}{The target can hold their breath for the spell's duration. Total concentration by the target is necessary, so Ironlungs prevents most forms of spell casting.}

\spellentry{Lightning Bolt}{6}{1 Round}{3d6 vs HEA for 1/2}{LOS 120 + 30/Finesse Feet}{Instantaneous}{4 + 1/Finesse Feet wide line}{6d6 damage}{Create opposed electrical charges between the caster's fingertip and their intended target, causing a bolt of electrical energy to travel from their finger to the target along the path of least resistance. Intervening objects and people along this path are affected by the bolt, as is the target.}

\spellentry{Fly}{7}{1 Round}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{4d6 + 6/Finesse Rounds}{1 humanoid}{Fly 120 + 30/Finesse Feet /Rounds}{The target is able to fly. Simple actions can be performed without penalty while in flight, but weapon attacks are made at a penalty (GM's discretion) and spell casting is impossible. The target cannot carry more than they could lift above their head.}

\spellentry{Wind}{8}{4 Rounds}{None}{0}{6 + 3/Finesse Hours}{60 + 20/Finesse Feet Radius}{Control wind}{Control the speed and direction of the wind. The wind is strong enough to propel a sailing vessel if desired, although a sail is an obvious necessity.}

\spellentry{Air Temperature}{9}{2 Rounds}{None}{90 + 20/Finesse Feet}{6 + 2/Finesse Hours}{60 + 20/Finesse Feet Radius}{Change temperature}{Change the temperature in the chosen area. Current weather effects are modified by the change, as dictated by common sense. Air Temperature can be targeted on the caster or another person, and the area of effect will move with that person.}

\spellentry{Anaerobia}{10}{1 Round}{4d6 vs PWR negates}{LOS 20 + 10/Finesse Feet}{4d20 + 8/Finesse Minutes}{1 creature}{Survive without air}{This spell satisfies the target's need for oxygen, making normal breathing unnecessary. All actions may be performed normally. Anaerobia does not protect the target from possible physical damage caused by hostile environments, such as fire, extreme cold, etc.}

\spellentry{Hawksoar}{11}{1 Round}{4d6 vs PWR negates}{Touch}{15 + 4/Finesse Rounds}{1 humanoid}{Fly and attack}{Hawksoar adds the abilities of combat and spell casting to the Fly spell. All actions are allowable and not subject to any penalties.}

\spellentry{Ekblovego}{12}{3 Rounds}{5d6 vs STR for half}{0}{1d6 + 1/Finesse Rounds}{30'x(150 + 30/Finesse)' line}{Massive wind blast}{Create a massive blast or gust of wind. All free standing objects and creatures are blown backwards 80 feet and suffer 1d6 DP each round they remain in the effect. Some structural damage to buildings may even result (GM's discretion).}

\spellentry{Vacuum}{13}{3 Rounds}{4d6 vs HEA}{100 + 30/Finesse Feet}{Instantaneous}{30 + 10/Finesse Feet Radius}{Remove air}{Creates an airless space in the area of effect. Aerobic creatures who fail the RC become comatose. Fires are immediately extinguished, sailing vessels becalmed, and other processes involving air are likewise halted. The GM will adjudicate any side-effects of the sudden decompression.}

\spellentry{Blitzen}{14}{2 Rounds}{4d6 vs AGI for half}{160 + 40/Finesse Feet}{8 + 2/Finesse Rounds}{15 + 5/Finesse Feet Radius}{1 lightning strike/Rounds}{Causes lightning to strike the indicated area once each round. All creatures in the area of effect are hit and suffer 6d6 DP for each attack. If the magician is casting this during a thunderstorm, all base parameters are doubled.}

\spellentry{Aerovoki}{15}{4 Rounds}{None}{40 + 10/Finesse Feet}{4 + 1/Finesse Minutes}{Special}{Summon air elementals}{Summon 1d6+1 air elementals to any point within the caster's range. Air must be available. The elementals will serve the caster, even in combat. They will not turn on them, but will make note of who summoned them. There is a 1d10 chance on each successive summoning that the elementials will attempt to take the caster with them at the end of the spell.}

\spellentry{Wrath}{16}{4 Rounds}{5d6 vs HEA for half}{160 + 40/Finesse Feet}{4 + 1/Finesse Minutes}{Up to 1/4 met Radius}{Summon wind storm}{Summon a wind storm of unbelievable strength. It can be in the form of a hurricane, tornado, or a straight line wind. Wind velocity is 300 mph or more, few structures will survive (the GM may give stone buildings an RC). Each round in the effect, creatures suffer 3d6 DP from being struck by flying debris. Spell casting is impossible. The RC must be made each round.}

\section{Vision}

\spellentry{Murk}{1}{1 Round}{3d6 vs PER negates}{20 + 5/Finesse Feet}{1d20 + 4/Finesse Rounds}{10 + 5/Finesse Feet Radius}{Reduce visibility}{Causes dust motes in the environment to coalesce onto water droplets to create a dirty, murky area. Vision is reduced by 40 + 10/Finesse Feet. Murk is effective in underwater as well as air.}

\spellentry{Distortion}{2}{1 Round}{None}{None}{1d20 + 4/Finesse Minutes}{5 + 2/Finesse Feet Radius}{Obscure vision}{Bends and shifts light rays erratically in the area of effect, causing forms within the area become blurry, indistinct, and are 90\% unrecognizable.}

\spellentry{Crystal}{3}{1 Round}{None}{20 + 5/Finesse Feet}{1d4 + 1/Finesse Minutes}{10 + 5/Finesse Feet Radius}{Increase visibility}{Cleansing the air of impurities clinging to water droplets. It is only effective in haze, fog, murky water, etc.}

\spellentry{Glamor}{4}{2 Rounds}{4d6 vs WIL negates}{Touch}{1d20 + 1/Finesse Minutes}{1 creature}{Boost COM 2 + 1/Finesse}{Alters the natural COM of the target, affecting reactions accordingly.}

\spellentry{Intuit}{5}{2 Rounds}{None}{Caster}{1d20 + 1/Finesse Minutes}{Caster}{Boost PER 2 + 1/Finesse}{Temporarily increase their PER for the duration of the spell.}

\spellentry{One Eye}{6}{1 Round}{3d6 vs PWR negates}{20 + 5/Finesse Feet}{1d8 + 2/Finesse Rounds}{1 creature}{Blind one eye}{Temporarily blinds target in one eye (caster's choice). The target's melee modifiers are reduced by 2 + 1/Finesse to reflect the loss of depth perception. If the target only had one eye to begin with, it will become totally blind.}

\spellentry{Discovery}{7}{2 Rounds}{None}{Caster}{1d20 + 4/Finesse Minutes}{Caster}{Magnify 8 + 2/Finesse times}{See minute things or details that escape normal vision. Rolls against PER are made with one less die than normal.}

\spellentry{Parallax}{8}{1 Round}{None}{0}{1d8 + 2/Finesse Rounds}{Caster}{Boost DV by 2 + 1/Finesse}{The caster's body seems to shift about suddenly. Attackers must try to correct for the sudden parallax error, in effect raising all the caster's DVs.}

\spellentry{Long Eyes}{9}{2 Rounds}{None}{0}{1d6 + 2/Finesse Minutes}{Caster}{Closer 8 + 2/Finesse times}{See distant objects as if they are much closer. They must have LOS, and certain atmospheric conditions (e.g. fog, haze, precipitation, etc.) can reduce the effectiveness by 1/2x. This also will increases the range of other LOS spells cast during Long Eyes duration. Finneses to increase the range of these spells are appled before Long Eyes multiples the range.}

\spellentry{Heat Vision}{10}{2 Rounds}{4d6 vs PWR negates}{Touch}{1d20 + 4/Finesse Minutes}{1 creature}{See heat @ 40 + 10/Finesse Feet}{See infrared instead of (NOT in addition to) light. Objects will vary in color from jet black (very cold) to bright red (very hot). A cold draft or emanations from ice interfere with heat vision by making forms opaque.}

\spellentry{Night Sight}{11}{2 Rounds}{None}{Caster}{1d4 + 1/Finesse Minutes}{Caster}{See in the dark}{See in natural darkness and shadows as well as in bright moonlight. If the effect is fully finessed they can even see in magical darkness or shadows.}

\spellentry{Invisibility}{12}{1 Round}{4d6 vs PWR negates}{Touch}{5 + 2/Finesse Minutes}{1 object or creature}{Make unseeable}{Surrounds the target with a magical field which shifts light out of phase. If the field is touched by an ungrounded physical object, the spell is broken. An invisible person can walk about or open doors, but cannot drop or pick up anything nor physically attack. Spells which have a physical manifestation break the enchantment. Invisible creatures are in the same phase and can see one another.}

\spellentry{Sight License}{13}{1 Round}{4d6 vs PWR negates}{Touch}{8 + 3/Finesse Hours}{1 creature}{Allows sight}{Temporarily grants the target the power of sight, even if it is blind or lacks sight organs.}

\spellentry{Blind}{14}{1 Round}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{Permanent}{1 creature}{Remove vision}{Makes the target blind until cancelled by the caster or Revoked. The target must be capable of vision for the spell to be effective, i.e., Blinding a worm is of little consequence.}

\spellentry{Projection}{15}{4 Rounds}{None}{0}{5 + 2/Finesse Minutes}{LOS 250 + 100/Finesse Feet}{Create mass vision}{All people and/or creatures in the area of this spell see just what the caster desires them to see. The vision appears quite real, but has no audio, temperature or touch components. The caster MUST maintain concentration for the duration of this spell, or the vision disappears immediately.}

\spellentry{See All}{16}{4 Rounds}{None}{Caster}{12 + 4/Finesse Rounds}{Caster}{Create LOS}{See through one solid object or creature. This can be as small as a rabbit or as large as a mountain, but can be only one thing. Thus the caster can cast a line of sight spell through any interposing creature or object subject to that spell's normal range restrictions.}

\section{Scrying}

\spellentry{Detect Motivation}{1}{1 Round}{3d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{1d6 + 1/Finesse Rounds}{1 humanoid}{Confirm motivation}{For each round of the spell's duration, the magician may think of a motivation. If the target has that motivation, the caster will know it positively. On the 3rd attempt, the target is allowed an RC. If the RC is successful, further attempts fail, but the results of the third guess are still revealed to the caster.}

\spellentry{Hear}{2}{2 Rounds}{None}{40 + 10/Finesse Feet}{1d4 + 1/Finesse Minutes}{10 + 5/Finesse Feet Radius}{Remote hearing}{Hear any noises or speech in the chosen area as if your ears are there. The area does not have to be in the caster's LOS.}

\spellentry{Read Emotions}{3}{1 Minute}{3d6 vs WIL negates}{LOS 40 + 10/Finesse Feet}{1d4 + 1/Finesse Minutes}{1 living creature}{Know thoughts}{Know the surface thoughts of the targeted creature. These are very general and primal in nature, such as hunger, anger, contentment, sadness, etc., regardless of the creature's INT.}

\spellentry{Mask Motivation}{4}{1 Round}{None}{Touch}{1d4 + 1/Finesse Hours}{1 creature}{Mask motivation}{Mask Motivation enables the caster to deny that information, or feed false information (caster's choice), to anyone using spells.}

\spellentry{See}{5}{2 Rounds}{None}{60 + 15/Finesse Feet}{1d4 + 1/Finesse Minutes}{10 + 5/Finesse Feet Radius}{Remote vision}{See everything in the chosen area as if your eyes are there. The area does not have to be in the caster's LOS, but after casting See, the caster must remain motionless and concentrate or the spell is broken.}

\spellentry{Witchsmeller}{6}{1 Round}{None}{0}{1d4 + 1/Finesse Rounds}{10 + 5/Finesse Feet Radius}{Uncover magic}{Any enchanted items in the area of effect glow with a silver aura in the eyes of the caster. Very magical items, such as relics, glow with a gold aura. The strength of the glow is reflective of the strength of the magic. Other spells must be used to gain more knowledge (e.g. Know Element, etc).}

\spellentry{Know True Motivation}{7}{1 Round}{4d6 vs WIL negates}{LOS 60 Feet}{Instantaneous}{1 target}{See true motivation}{Allows the caster to know the true motivation of any creature in the area with a signle casting. Mask Motivation, however, will deceive this spell. If the target succeeds on their RC, they will become aware of the scrying and the caster will not receive the information.}

\spellentry{Locate}{8}{2 Rounds}{4d6 vs WIL negates}{Caster}{1d4 + 1/Finesse Minutes}{200 + 20/Finesse Feet beam}{Discover location}{Locate any item or creature which the caster previously seen or touched. The caster first visualizes the target of their search. A search beam moves with the caster, and an image forms as soon as the subject comes within range of the beam. The vision can only be seen by the caster, and receds into the distance to where the item or creature is actually located.}

\spellentry{Senses}{9}{2 Rounds}{None}{60 + 15/Finesse Feet}{1d4 + 1/Finesse Minutes}{Special}{1 + 1/Finesse senses}{Allows one of more senses to operate remotely. The caster must remain motionless for the spell's duration. This spell poses risks, for damaging effects present in the remote area can attack the caster through the spell. For example, poison gas in the remote area may cause damage if the caster choses smell as a remote sense.}

\spellentry{Truthtell}{10}{1 Round}{4d6 vs WIL negates}{Touch}{1d4 + 1/Finesse Minutes}{1 creature}{Prevent lying}{Any creature failing the RC must answer all the caster's questions truthfully for the duration of the spell.}

\spellentry{Futuresee}{11}{2 Minutes}{None}{Caster}{3 + 1/Finesse questions}{Caster}{See the future}{Futuresee enables the caster to determine what will happen in the immediate future (1 + 1/Finesse days) by asking yes-or-no questions of an elemental spirit. The truthfulness of the answers is from 50 + 10/Finesse percent. If not truthful, responses can be totally false, or couches as half-truths (GM's discretion).}

\spellentry{Legends}{12}{See text}{None}{Special}{1 + 1/Finesse clues}{Special}{Know antiquity}{Gain knowledge of the items, places, and people of antiquity. If the caster is touching the item, standing on the site, or next to a grave, the casting time is one day; otherwise one week. The caster can learn from one to five pieces of information about the subject in the form of clues, or as a range of numbers. The information is always truthful, but the player must make correct interpretations.}

\spellentry{Thingtell}{13}{4 Rounds}{None}{Touch}{1d6 + 1/Finesse Minutes}{1 object}{Speak with objects}{Speak with inanimate objects like chairs, rocks, trees, bushes, etc. The object cannot volunteer information, but will answer the caster's questions truthfully, one question per round of the spell's duration.}

\spellentry{Backfire}{14}{1 Round}{None}{Caster}{3d6 + 3/Finesse Hours}{Caster}{Reverse scrying}{Protects the caster from scrying attempts, and also reverses the process so that they find out what information the opposing caster was seeking from them, ignoring any normal RC. They can also learn the direction and distance of the attempt.}

\spellentry{Scan}{15}{1 Minute}{None}{Caster}{1d8 + 1/Finesse Rounds}{800 + 300/Finesse Feet beam}{Scan area}{After deciding on what piece of information the caster wishes to know ("horses nearby?", "any carnivorous plants in the area?", etc.), a beam of violet light shoots from the magician's eyes and sweeps the area like a searchlight. The beam stops and points at matching objects if contacting it, giving the direction and distance. The beam moves with the caster, but is blocked by walls, trees, etc.}

\spellentry{Godspeak}{16}{1 D}{None}{Caster}{1 + 1/Finesse questions}{Caster}{Find unmitigated truth}{Ask any question of the caster's sworn diety and be assured of a 100\% truthful answer. The caster must be on good terms with their sworn deity, through regular obeisance, performance of deeds to further the deity's goals, etc. Optionally, the spell always functions, but there may be some penalty on EU recharging time if the caster has been lax, or the diety may only answer one question regardless of finesses (GM's discretion).}

\section{Dimensions}

\spellentry{Shimmer}{1}{1 Round}{3d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{1d4 + 1/Finesse Rounds}{1 creature}{Increase CDV 2 + 1/Finesse}{While under the influence of Shimmer, the target's physical form appears indistinct and wavering, effectively increasing their CDV. The caster must be touched by air while murmuring the spell.}

\spellentry{Thought Move}{2}{1 Round}{3d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{Instantaneous}{1 object}{Move 20 + 5/Finesse lb}{Move a non-living, free standing object within the weight limit and in the caster's LOS from one place to another within the allowable range. Since the motion is slow, no damage is inflicted from striking other objects, unless circumstances dictate otherwise (walking into a set spear, falling from a height, etc)}

\spellentry{Send Object}{3}{1 Round}{None}{Touch}{Instantaneous}{1 object}{Send small object}{Send an object enclosed in the caster's fist to any previously memorized location. The object can not exceed 10 pounds.}

\spellentry{Peer}{4}{2 Rounds}{None}{Memloc}{1d6 + 1/Finesse Minutes}{60' Radius}{See memorized place}{Monitor any memorized location for the spell's duration. Only sight is allowed; no sounds, smells, etc. can be discovered.}

\spellentry{Magic Fence}{5}{1 Round}{4d6 vs WIL negates}{Caster}{1d6 + 1/Finesse Minutes}{5 + 2/Finesse Feet Radius}{Exclude summonings}{Create a glowing boundary (shaped to the caster's choosing), centered on the caster. Other planer and immaterial creatures, and magically transported objects cannot pass this line unless the RC is made (objects save vs the WIL of the caster of the spell transporting it).}

\spellentry{Call Object}{6}{3 Rounds}{None}{Special}{Instantaneous}{1 + 1/Finesse objects}{Recall from storage}{Retrieve up to five objects from the place where they were previously stored by a Send Object spell. The five objects do not have to have been sent to the same location.}

\spellentry{Hole}{7}{3 Rounds}{None}{Touch}{3d6 + 2/Finesse Rounds}{6 + 6/Finesse Foot Radius}{Create hole}{Creates a hole (6 + 6/Finesse inches deep) in any non-living object. No light can pass through this hole, making it appear totally black, but objects and creatures can pass through with no resistance. Magical objects get a 4d6 RC versus PWR to negate the effect. For example, the caster could cast this on a door next to the lock, reach inside, and unlock the door.}

\spellentry{Giant Stride}{8}{2 Rounds}{None}{Multitouch}{1d6 + 1/Finesse Rounds}{Targets}{Port 160 + 40/Finesse Feet}{Open a temporary two-way door to a spot within their LOS. The door is invisible, except to the caster and anyone touched during the casting time. The starting and ending points must be touched by air. The effect for observers is that those affected disappear into thin air and reappear in the blink of an eye.}

\spellentry{Send Package}{9}{3 Rounds}{None}{Touch}{Instantaneous}{1 object}{Send package}{Send an inanimate object of up to 100 + 10/Finesse Pounds and no more than 2 + 1/Finesse Feet Radius can be sent to any previously memorized location.}

\spellentry{Pocket}{10}{3 Rounds}{None}{Multitouch}{2d6 + 2/Finesse Rounds}{Targets}{Open elemental pocket}{Create an opening to a pocket dimension. The opening appears suspended in air, visible only to the caster and anyone touched during the casting time. If they enter they appear to vanish into thin air. Normal human life is supported in the dimension, and at the end of the duration any living creatures still within the dimension are returned to their starting point. The caster must be touching air for this spell to work.}

\spellentry{Portable Hole}{11}{7 Rounds}{None}{Touch}{1d6 + 1/Finesse Hours}{6 + 6/Finesse Foot Radius}{Create portable hole}{This creates a hole (6 + 6/Finesse Inches Deep) in any non-living object. No light can pass through this hole, making it appear totally black, but objects and creatures can pass through. Grasping the hole by the outside edge, the Portable Hole can be peeled off and reused elsewhere. Magical objects get a 4d6 RC vs PWR to negate the effect.}

\spellentry{Call Package}{12}{5 Rounds}{None}{Special}{Instantaneous}{1 + 1/Finesse objects}{Recall package}{Retrieve up to five packages from the place where they were previously stored by a Send Package spell. The five packages do not have to have been sent to the same location.}

\spellentry{Magic Door}{13}{4 Rounds}{None}{Multitouch}{1d4 + 1/Finesse Hours}{Memloc}{Open door to memloc}{Open a door to any location the caster has memorized, visible only to themselves and those touched during the casting time. The caster must be touching air, and air must be available at the destination for the spell to work. The location can be any distance away, but must be on the same plane.}

\spellentry{Vanish}{14}{4 Rounds}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{3d6 + 3/Finesse Hours}{1 creature}{Send to pocket dim.}{The caster must be touching air for this spell to work. The target is immediately dispatched to a small, life supporting pocket dimension in the air for the duration of the spell.}

\spellentry{Otherworld}{15}{4 Rounds}{None}{0}{Permanent}{10 + 5/Finesse Feet Radius}{Open portal}{The caster must have at hand still air with a volume of at least the area of effect. When the spell is cast, the caster concentrates on a dimension and the place within that dimension where the portal will lead. A visible portal opens, providing a corridor to the target. Anyone can then step through this portal, and be there. The corridor cannot be opened onto the same dimension from which it is cast.}

\spellentry{Banish}{16}{4 Rounds}{5d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{Permanent}{1 creature}{Send creature}{Banish is the same as Vanish, except the RC is more difficult and the duration is permanent unless Revoked.}