\chapter{Fire Magic}
\label{ch:elemental-fire-magic}

Magicians of the Fire element spend endless hours of hot and sweaty concentration to unlock the secrets of flames and heat. Their studies make many of them nimble and quick, and most have passionate emotions. Their spell groups are described below.

Fire Magic typifies the power of the element, in spells like Fireball and Conflagration, and the ability to manipulate the element.

The Illumination group provides the caster with light and darkness spells of varying power.

Conjurations are used to create phantasms and shams to delude the foolish and gullible.

Mind Twisters are similar to the Charms group of the Water magicians, but are generally more powerful and, perhaps, even sinister. Because of the great power that must be channeled to use this spells, they are more difficult to learn, and drain their user. The base costs for these spells is 600 EP. 

\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{tabular}{@{} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth}}
\begin{enumerate}
	\item Fire Magic
	\begin{enumerate}
		\item Flint and Steel
		\item Flaming Arrows
		\item Flame
		\item Part Fire
		\item Immolate
		\item Quench
		\item Nimble
		\item Fireball
		\item Flame Dance
		\item Firebreath
		\item Flame Barrier
		\item Flameshape
		\item Conflagration
		\item Comet
		\item Fajrovoki
		\item Spirit Fire
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{1}
	\item Illumination
	\begin{enumerate}
		\item Glowworm
		\item Magic Torch
		\item Light Shield
		\item Magic Lantern
		\item Whip
		\item Limn
		\item Alarm
		\item Latent Vision
		\item Rainbow Bridge
		\item Dazzle
		\item Blade
		\item Mirror Wall
		\item Reflective Travel
		\item Sun Call
		\item Ruby Death
		\item Beamrider
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{2}
	\item Conjurations
	\begin{enumerate}
		\item Veil
		\item Emit Self
		\item Phantasmal Fire
		\item Emit Object
		\item Mask
		\item Emanate Self
		\item Phantasmal Air
		\item Emanate Object
		\item Shrouding
		\item Radiate Self
		\item Phantasmal Water
		\item Radiate Object
		\item Envelop
		\item Project Self
		\item Phantasmal Earth
		\item Project Object
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{3}
	\item Mind Twisters
	\begin{enumerate}
		\item Gutter
		\item Flame Dreams
		\item Dismay
		\item Unman
		\item Forget
		\item Stray Thought
		\item Chaos
		\item Befriend
		\item Neophyte
		\item Betask
		\item Charlatan
		\item Arson
		\item Amnesia
		\item Babe
		\item Fever
		\item Pith
	\end{enumerate}
\end{enumerate} \\
\end{tabular}
\pagebreak
\section{Fire Magic}

\spellentry{Flint and Steel}{1}{1 Round}{None}{Caster}{4 + 2/Finesse Rounds}{Caster's thumb}{Make a flame}{Snap your fingers and produce a one inch high flame from your thumb.}

\spellentry{Flaming Arrows}{2}{1 Round}{None}{LOS 20 + 5/Finesse Feet}{6 + 2/Finesse Rounds}{1 arrow}{Make burning arrows}{Cause one arrow each round which is ready to fire to burst into flames. The arrow will inflict an additional 1d8 points of damage if it hits. The caster must have uninterrupted concentraion for the duration of the spell.}

\spellentry{Flame}{3}{1 Round}{3d6 vs AGI for half}{LOS 30 Feet}{3 + 1/Finesse Rounds}{1 target}{1d6 fire damage}{A ball of flame appears harmlessly in the caster's hand. They may then hurl it up to 30 Feet, once per round, to automatically hit any target in direct LOS. Non-living, flammable targets can be ignited by the fire. Living things are not ignited, but do take damage.}

\spellentry{Part Fire}{4}{1 Round}{Special}{LOS 10 + 3/Finesse Feet}{8 + 2/Finesse Rounds}{3 Feet Wide x 20 + 5/Finesse Long}{Part fire}{Clear a narrow path through continuously burning natural or magical fires (magical fires get an initial RC of 9 or less on 3d6). The path must be straight, and creatures no larger than a normal human may pass through the flames unscathed. This spell is of no use against flame attacks.}

\spellentry{Immolate}{5}{1 Round}{4d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{4 + 1/Finesse Rounds}{1 target}{Flaming form}{Envelop the target's form in illusory magical flames. The target gains resistance to flame attacks (subtract 3d6 DP from each damage roll) and is also more difficult to see through the flames (increase CDV and MDV by +2) for the duration of the spell.}

\spellentry{Quench}{6}{1 Round}{None}{LOS 50 + 10/Finesse Feet}{Instantaneous}{5 + 3/Finesse torches}{Extinguish fire}{Snuffs out normal fires of the size indicated by the area of effect. A small campfire is equivalent to 5 torches, a large bonfire 10 torches, etc. This spell has no effect versus magical fires or sudden fire attacks.}

\spellentry{Nimble}{7}{1 Round}{4d6 vs AGI negates}{LOS 20 + 5/Finesse Feet}{8 + 2/Finesse Rounds}{1 living target}{Raise AGI and skills}{Increases the AGI of the target creature by 1d6 ranks, and all existing AGI based skills increase +2 ranks for the duration of the spell.}

\spellentry{Fireball}{8}{1 Round}{4d6 vs HEA for half}{LOS 120 + 30/Finesse Feet}{Instantaneous}{40 + 5/Finesse Feet Radius}{8d6 fire damage}{Draw a bit of elemental fire from the surrounding environment and direct it to explode in a designated area. All creatures failing their RC take the indicated damage. Dry, combustible materials can be easily set aflame with this spell.}

\spellentry{Flame Dance}{9}{2 Rounds}{None}{Caster}{8 + 2/Finesse Rounds}{1 being}{Move with flames}{Walk through normal and magical flames for the duration of the spell without damage. While within the flames, the caster's CDV and MDV are increased by +4.}

\spellentry{Firebreath}{10}{2 Rounds}{4d6 vs AGI for half}{0}{3 + 1/Finesse Rounds}{30' x 10' cone}{4d6 fire damage}{Spout flame from the caster's mouth once each round. All creatures within this area that fail their RC take the indicated damage. Non-living, flammable objects can be ignited with the spell.}

\spellentry{Flame Barrier}{11}{2 Rounds}{4d6 vs AGI for half}{40 Feet}{15 + 4/Finesse Rounds}{20 + 10/Finesse Feet Radius}{4d6 fire damage}{Call forth a roaring wall of fire, which may be in any shape the caster chooses as long as it does not exceed the area indicated. The flames can be extinguished by normal or magical rain in 5 rounds. Creatures attempting to cross the flames take 4d6 DP each round for as long as the crossing takes, subject to an RC each round.}


\spellentry{Flameshape}{12}{4 Rounds}{Special}{Touch}{1 + 0.25/Finesse Hours}{2 + 1/Finesse Feet Radius}{Create fiery object}{Transforms a normal flame (magical flame gets an initial RC of 9 or less on 3d6) into a solid and harmless (to the caster) clay-like material. The flame will hold its shape until its spell expires.}

\spellentry{Conflagration}{13}{1 Round}{None}{100 Feet}{15 + 4/Finesse Rounds}{40 + 20/Finesse Feet Radius}{4d6 fire damage}{An enhanced version of the Flame Barrier spell. The flames are more intense, and can set combustible materials afire even if they are wet. Normal rain cannot extinguish a Conflagration, though Water spells can.}

\spellentry{Comet}{14}{2 Rounds}{5d6 vs HEA for half}{200 + 50/Finesse Feet}{Instantaneous}{Special}{14d6}{A super potent Fireball. Creatures within 20 feet of the target take full damage, those within 60 feet take 1/2x damage, and those within 100 feet take 1/4x damage. A successful RC reduces damage by a further 1/2x.}

\spellentry{Fajrovoki}{15}{4 Rounds}{None}{LOS 40 + 10/Finesse Feet}{4 + 1/Finesse Minutes}{Special}{Summon fire elementals}{Summon 1d6+1 fire elementals to any point within the caster's range. A source of fire must be available (even a candle) within the target area. The elementals will serve the caster, even in combat. They will not turn on them, but will make note of who summoned them. There is a 1d10 chance on each successive summoning that the elementials will attempt to take the caster with them at the end of the spell.}

\spellentry{Spirit Fire}{16}{2 Rounds}{None}{Special}{Special}{Special}{Selective damage}{Allows the caster to duplicate Flame, Firebreath, Flame Barrier, Fireball, or Comet. The range, duration, area, resistance check and effect are the same as the duplicated spell. A Spirit Fire burns with a white flame which is harmless to creatures the caster states are friends. Only living creatures can be damaged. Combustable materials cannot be ignited with this spell, even if the effect it is duplicating is capable of such.}

\section{Illumination}

\spellentry{Glowworm}{1}{1 Round}{None}{Touch}{10 + 3/Finesse Minutes}{50 lb object}{Make item glow}{Causes an inanimate target which is liftable by the caster to give of a soft, candle-like glow. The amount of light produced is suitable for reading or other close work.}

\spellentry{Magic Torch}{2}{1 Round}{None}{LOS 10 + 5/Finesse Feet}{2 + 1/Finesse Hours}{1 object}{Create torchlight}{Causes a handheld object to shed light equivalent to a normal torch.}

\spellentry{Light Shield}{3}{1 Round}{None}{Caster}{1d6 + 1/Finesse Rounds}{Caster}{See text}{Gather existing light into a circular shield. The base spell yields a buckler, two finesses a shield, and four finesses a shield which subtracts -2 from any damage done on a successful attack. There must be at least one torch within 10 feet of the caster for this spell to function.}

\spellentry{Magic Lantern}{4}{1 Round}{None}{Touch}{2 + 1/Finesse Hours}{1 object in hand}{Create lantern}{Create a controllable light source. The magician can vary the amount of light produced, from full lantern light to completely dark, on command.}

\spellentry{Whip}{5}{1 Round}{None}{Caster}{2d6 + 2/Finesse Rounds}{Caster}{Create enchanted whip}{Gather existing light into a long, flexible beam and wield it like a once enchanted whip for the spell's duration. The Whip does 1d6+1 DP per successful attack.}

\spellentry{Limn}{6}{1 Round}{3d6 vs AGI negates}{LOS 20 + 5/Finesse Feet}{1d6 + 2/Finesse Rounds}{5 + 3/Finesse Feet Radius}{Outline with light}{Outlines creatures and objects in dim or dark surroundings with light from tendrils in the caster's fingertips, making them easier to see. Mobile creatures are allowed an RC to avoid the effect; immobile creatures and objects get no RC.}

\spellentry{Alarm}{7}{1 Minute}{4d6 vs PER (detect)}{LOS 100 Feet}{1d8 + 2/Finesse Hours}{See text}{2 + 1/Finesse beams}{Protect a room, building, campsite, etc. with a faint beam of light which, when broken, emits a shrill alarm. Each beam can extend up to 20 feet, if necessary.}

\spellentry{Latent Vision}{8}{3 Rounds}{None}{Caster}{1d6 + 2/Finesse Minutes}{LOS}{View recent past}{Allows the caster to view the events of the recent past. The vision starts at the begining of the duration and moves forward in real time until reaching the time of casting. The caster is blind to anything in the presnet, but is able to move around (though should be be wary of obstacles).}

\spellentry{Rainbow Bridge}{9}{4 Rounds}{4d6 vs PWR negates}{Multitouch}{1d8 + 1/Finesse Minutes}{120 + 30/Finesse Feet}{Create bridge}{Create a bridge of light. As many people or creatures the caster can touch during the casting time may use the bridge.}

\spellentry{Dazzle}{10}{1 Round}{4d6 vs AGI negates}{40 + 10/Finesse Feet}{2 Rounds}{10 + 5/Finesse Feet Radius}{Blind for 1d4 Rounds}{Causes a ball of light to appear in the caster's hand and may be hurled up to the target distance. The ball explodes after 2 rounds, producing a shower of brilliant light fragments that temporarily blind any vision-endowed creature in the area of effect.}

\spellentry{Blade}{11}{1 Round}{None}{Caster}{2d6 + 2/Finesse Rounds}{Caster}{See text}{Gathers existing light into a narrow, blade-shaped beam which the magician may wield as a twice enchanted, one-handed sword. Such a weapon does 1d10+2 DP per successful strike.}

\spellentry{Mirror Wall}{12}{4 Rounds}{4d6 vs PER negates}{Touch}{1d4 + 1/Finesse Minutes}{10 + 5/Finesse Feet Radius}{Creates one way mirror}{Turn any flat surface into a temporary one-way mirror. Anyone on the backside of the wall can see through it, but it looks like a normal wall to those who fail their RC.}

\spellentry{Reflective Travel}{13}{2 Rounds}{None}{Caster}{Instantaneous}{Caster}{Travel 1000 met LOS}{Instantly travel to the source of a reflection.}

\spellentry{Sun Call}{14}{2 Rounds}{None}{0}{1d4 + 1/Finesse Hours}{20 + 10/Finesse Feet Radius}{Make daylight}{Bring the brilliance of the noon sun to any area.}

\spellentry{Ruby Death}{15}{1 Round}{5d6 vs AGI avoids}{LOS 120 Feet}{Instantaneous}{1 target}{15d6 damage}{Fire a red ray of concentrated light energy from the caster's finger. Any target which fails its RC takes the indicated amount of damage.}

\spellentry{Beamrider}{16}{4 Rounds}{None}{Caster}{Instantaneous}{Caster}{Travel 1000 mets LOS}{Travel to the source of any light within LOS in the blink of an eye.}

\section{Conjurations}

\spellentry{Veil}{1}{1 Round}{4d6 vs PWR negates}{Caster}{15 + 5/Finesse Minutes}{Caster}{Veils caster}{Cast a distorting veil over the caster's face and clothes. Anyone viewing the caster who fails the RC will believe the caster to be a total stranger.}

\spellentry{Emit Self}{2}{2 Rounds}{None}{Caster}{4 + 2/Finesse Minutes}{Caster}{Make an image}{Create a transparent image of the caster. The image can be directed to appear in any direction, possibly through intervening obstacles. The image moves, sounds and smells like the caster. The image can range from one inch in size up to the caster's true size.}

\spellentry{Phantasmal Fire}{3}{1 Round}{None}{Special}{Special}{Special}{Duplicate Fire spell}{Duplicates the visual, sound and smell effects of any Fire spell, but completely harmless. This spell cannot duplicate any spell whose effect is visual only (such as Magic Torch) or any Conjuration spells.}

\spellentry{Emit Object}{4}{2 Rounds}{None}{LOS 10 Feet}{4 + 2/Finesse Minutes}{Single 1000 lb object}{Make an image}{Project an image of another person or object under the same constraints as the Emit Self spell. The object can be no larger than the caster.}

\spellentry{Mask}{5}{2 Rounds}{None}{Caster}{2d4 + 1/Finesse Hours}{Caster}{Disguises caster}{The caster first visualizes the memorized image of a person and concentrates. Mask then makes the caster appear as that person, including clothing. The caster cannot change race or general size with this magic. The adventurer must inform the GM when they are memorizing a person, which takes 5 minutes of game time.}

\spellentry{Emanate Self}{6}{4 Rounds}{None}{Caster}{6 + 3/Finesse Minutes}{Caster}{Make an image}{Project a non-transparent, real-seeming image of themselves. Other aspects of the image are as in the \u{Emit Self} spell.}

\spellentry{Phantasmal Air}{7}{1 Round}{None}{Special}{Special}{Special}{Duplicate Air spell}{Duplicates the visual, sound and smell effects of the any Air spell, but completely harmless. The spell cannot duplicate any spell whose effect is visual only.}

\spellentry{Emanate Object}{8}{4 Rounds}{None}{LOS 10 Feet}{6 + 3/Finesse Minutes}{Single 1000 lb object}{Make an image}{Project an image of another person or object under the same constraints as the Emanate Self spell. The object can be no larger than the caster.}

\spellentry{Shrouding}{9}{3 Rounds}{4d6 vs PWR negates}{Multitouch}{2d4 + 1/Finesse Minutes}{Targeted creatures}{Disguises group}{Disguises those touched to appear to be a those expected to be found in that environment to anyone failing the RC. For example, if in a hostile castle they might appear as a group of appropriately dressed guards.}

\spellentry{Radiate Self}{10}{6 Rounds}{None}{Caster}{8 + 4/Finesse Minutes}{Caster}{Make an image}{Creates an image which can move, sound and smell either independently or under control of the caster. All other aspects of this spell are the same as the Emanate Self spell.}

\spellentry{Phantasmal Water}{11}{1 Round}{None}{Special}{Special}{Special}{Duplicate Water spell}{Duplicates the visual, sound and smell effects of any Water spell, but completely harmless. It cannot duplicate any spell whose effect is visual only.}

\spellentry{Radiate Object}{12}{6 Rounds}{None}{LOS 10 Feet}{8 + 4/Finesse Minutes}{Caster}{Make an image}{Project an image of another person or object under the same constraints as the Radiate Self spell. The object can be no larger than the caster.}

\spellentry{Envelop}{13}{4 Rounds}{None}{Multitouch}{20 + 10/Finesse Minutes}{Targeted creatures}{Disguises group}{The targets appear to blend perfectly into the background. They must be touched to be found.}

\spellentry{Project Self}{14}{1 Minute}{None}{Caster}{10 + 5/Finesse Minutes}{Caster}{Program image}{Preprogram an image to appear at sometime up to one year into the future. It can be triggered by a specific time, the utterance of a particular word, or the occurrence of a particular event. All other aspects of the image are as in the Radiate Self spell.}

\spellentry{Phantasmal Earth}{15}{1 Round}{None}{Special}{Special}{Special}{Duplicate Earth spell}{Duplicate the visual, sound and smell effects of any Earth spell, but completely harmless. It cannot duplicate any spell whose effect is visual only.}

\spellentry{Project Object}{16}{1 Minute}{None}{LOS 10 Feet}{10 + 5/Finesse Minutes}{Object}{Make an image}{This spell allows the caster to project an image of another person or object under the same constraints as the Project Self spell. The object can be no larger than the caster.}

\section{Mind Twisters}

\spellentry{Gutter}{1}{1 Round}{None}{Caster}{1 Round}{5' Radius}{Cause doubt}{Causes beings who are within melee range and able to see a flame held in the caster's hand to suffer momentary doubt as to how to attack. They automatically lose initiative the next round.}

\spellentry{Flame Dreams}{2}{1 Minute}{None}{LOS 20 + 5/Finesse Feet}{1d3 + 1/Finesse Minutes}{1 humanoid}{Plan dream}{Allows the caster to dictate circumstances of the target's dream. The target must already be asleep and within 20 Feet of a fire (campfire, fireplace, etc.). Fire must be a component of the dream. If the target is present in the dream and killed, they will wake just before their dreamed death.}

\spellentry{Dismay}{3}{1 Round}{3d6 vs WIL negates}{Touch}{1d10 + 2/Finesse Rounds}{80 + 20/Finesse Feet Radius}{INT \verb+<+ 3 flee}{Touching a large fire, any non-intelligent creatures within sight of the fire flee from the caster at their maximum movement rate, for the spell's duration.}

\spellentry{Unman}{4}{1 Round}{3d6 vs WIL negates}{Touched}{1d4 + 1/Finesse Rounds}{60 + 15/Finesse Feet Radius}{INT \verb+>+ 3 flee}{Touching a large fire, any intelligent creatures within sight of the fire pause one round in fear before fleeing the caster at their maximum movement rate, for the spell's duration.}

\spellentry{Forget}{5}{1 Round}{4d6 vs WIL negates}{Touch}{Instantaneous}{1 creature}{Forget recent events}{Causes the target to forget the events from the present back four rounds. One finesse changes the time period to four minutes, two finesses to one hour, three finesses to one day, and four finesses to three days. Willing subjects need not make an RC. Revoking the spell doesn't restore the lost memories.}

\spellentry{Stray Thought}{6}{1 Round}{4d6 vs WIL negates}{LOS 40 + 10/Finesse Feet}{1d4 + 1/Finesse Rounds}{1 creature INT \verb+>+ 4}{Insinuate action}{Causes the target to undertake an action which is reasonable under the circumstances as though it was their own. An unreasonable (GM discresion) action will cause the spell to fail.}

\spellentry{Chaos}{7}{1 Round}{4d6 vs WIL negates}{LOS 40 + 10/Finesse Feet}{1d6 + 1/Finesse Rounds}{10 + 5/Finesse Feet Radius}{Create disorder}{While targeting a fire at least the size of a camp fire, creatures behave erratically. On a 1 on 1d10, creatures carry out planned actions. Otherwise, creatures with INT \verb+>+ 4 bicker among themselves. Creatures with INT \verb+<+ 4 who roll 1 on 1d4 attack each other, otherwise they simply mill about.}

\spellentry{Befriend}{8}{2 Rounds}{4d6 vs WIL}{60 + 15/Finesse Feet}{1d20 + 4/Finesse Rounds}{10 + 5/Finesse Feet Radius}{Enemies = friends}{Cancels any hatred towards the caster and their companions by creatures within the area of effec. This will last as long as the none of the adventurers take a hostile action (including casting spells) towards the Befrended targets. As with other spells in this group, targets must be within 80 Feet of a medium fire (campfire or larger) to be affected.}

\spellentry{Neophyte}{9}{2 Rounds}{4d6 vs WIL negates}{LOS 40 + 10/Finesse Feet}{1d8 + 2/Finesse Rounds}{1 humanoid INT \verb+>+ 4}{Forget background}{If the target fails the RC  they suddenly forgets all their background training for the spell's duration. Fighters don't have a clue as to the use of their weapons, magicians can no longer remember spells, priests cannot remember the name of their deity nor the dogma of their religion (thus becoming powerless), etc.}

\spellentry{Betask}{10}{1 Minute}{4d6 vs WIL negates}{Touch}{Permanent}{1 being INT \verb+>+ 7}{Set a task}{The caster touches the target with a magical flame held in the caster's hand and sets them a single task. The object of the task must be within the ability of the target to accomplish, and cannot take longer than a year, else the spell fails. The target feels duty-bound to perform the assigned mission, and will devote all their energies and resources to its accomplishment.}

\spellentry{Charlatan}{11}{2 Rounds}{4d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{2d6 + 2/Finesse Rounds}{1 humanoid}{"act" other background}{The target immediately assumes a different background of the caster's choosing for the spell's duration. Whatever the new background, the target feels they are giving a stellar performance, but blows never hit, spells never succeed, etc.}

\spellentry{Arson}{12}{2 Rounds}{5d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{1d8 + 2/Finesse Minutes}{1 being INT \verb+>+ 4}{Change surroundings}{The target must be capable of seeing a flame held in the caster's hand, and suddenly finds themselves in a vaguely familiar, but perplexing, network of flaming walls. The basic layout is any building the target is familiar with (preferably their home), and the target must act accordingly.}

\spellentry{Amnesia}{13}{1 Round}{5d6 vs WIL negates}{Touch}{Instantaneous}{1 creature}{Forget events}{Causes the target to forget events from the present back one week. One finesse changes the time period to four weeks, two finesses to four months, three finesses to four years, and four finesses cause total memory loss. Willing subjects need not make an RC. Revoking the spell doesn't restore the lost memories.}

\spellentry{Babe}{14}{1 Minute}{5d6 vs WIL negates}{Touch}{Permanent}{1 creature}{Turn into baby}{Immediately reduces the target's intellect and physical coordination to that of an infant, though their body remains full-sized. The effect is permanent unless cancelled by the caster or Revoked.}

\spellentry{Fever}{15}{1 Minute}{5d6 vs WIL negates}{Caster}{8 + 4/Finesse Hours}{Caster's scryer}{Obviate sendings}{Anyone attempting a scrying spell on a magician under a Fever spell must make an to avoid serious brain damage (INT drops to 1) as they suddenly feel overcome by a fever. If the RC is made, the scrying spell proceeds.}

\spellentry{Pith}{16}{4 Rounds}{5d6 vs WIL negates}{Touch}{Permanent}{1 creature}{Scramble brain}{Magically scrambles the target's brain. Basic brain function remains to keep the body alive, but thought and co-ordinated action are impossible until Pith is Revoked.}