\chapter{Water Magic}
\label{ch:elemental-water}

Water magicians are interested in the properties of all liquids, but are especially drawn to the rivers and oceans of Jaern.
Like the ocean, they seem timeless, with an inevitability to their actions.

Water Magic typifies the power of the element, in spells like Well and Ocean Cold, and the ability to manipulate the element.

Changings allow the magician or target to assume a different physical form.

Wardings provide protection to creatures and objects.

Charms are used to impose the caster's will on living creatures. Because of the great power that must be channeled to use these spells, they are more difficult to learn, and drain their user. Their base cost is 600 EP.

\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{tabular}{@{} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth} p{0.25\linewidth}}
\begin{enumerate}
	\item Water Magic
	\begin{enumerate}
		\item Spray
		\item Well
		\item Level
		\item Ice Ball
		\item Gills
		\item Hot or Cold
		\item Rain Dance
		\item Hail
		\item Dam
		\item Riverman
		\item Demolish
		\item Ocean Cold
		\item Torrent
		\item Divert
		\item Akvovoki
		\item Tidal Wave
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{1}
	\item Wardings
	\begin{enumerate}
		\item Forfend
		\item Fountain
		\item Poisonward
		\item Stay
		\item Weapon Ward
		\item Snugbug
		\item Monitor
		\item Debar
		\item Claw Ward
		\item Water Tongue
		\item Magic Ward
		\item Master Ward
		\item Spirit Ward
		\item Bliss
		\item Spongelungs
		\item Multicheck
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{2}
	\item Changings
	\begin{enumerate}
		\item Distill
		\item Drought
		\item Memorize Profile
		\item Condense
		\item Dissolve
		\item Dowse
		\item Evaporate
		\item Disguise
		\item Endwarf
		\item Transmute
		\item Enmass
		\item Elemental Mastery
		\item Age
		\item Transmogrify
		\item Mutate
		\item Rebirth
	\end{enumerate}
\end{enumerate} &
\begin{enumerate}
	\setcounter{enumi}{3}
	\item Charms
	\begin{enumerate}
		\item Rockabye
		\item Drowse
		\item Jackfrost
		\item Frostfeet
		\item Snare
		\item Shock
		\item Witchy Sleep
		\item Hibernate
		\item Dormant
		\item Siren
		\item Drover
		\item Greenthumb
		\item Nightmare
		\item Davy Jones
		\item Mesmerize
		\item Decant
	\end{enumerate}
\end{enumerate} \\
\end{tabular}
\pagebreak
\section{Water Magic}

\spellentry{Spray}{1}{1 Round}{3d6 vs AGI negates}{40 + 10/Finesse Feet}{4 + 1/Finesse Rounds}{10 + 5/Finesse Feet Radius}{Saturate}{A very fine elemental water mist spurts from the caster's fingertips, and everyone in the area of effect who fails the RC is drenched as if in a thunderstorm. Torch-sized fires are extinguished, and items susceptible to water damage might be harmed (GM's discretion.)}

\spellentry{Well}{2}{1 Minute}{None}{LOS 40 + 10/Finesse Feet}{30 + 8/Finesse Rounds}{6 + 2/Finesse Feet Radius}{Dig a well}{Find potable water in any natural environment by digging a well from which any creature may drink. The water is real and can be stored for later use.}

\spellentry{Level}{3}{1 Round}{None}{LOS 30 + 5/Finesse Feet}{15 + 4/Finesse Rounds}{5 + 2/Finesse Feet Radius}{Change 4 + 1/Finesse Feet}{Control the level of any aqueous liquid within the area indicated. The liquid is displaced, not created or destroyed.}

\spellentry{Ice Ball}{4}{1 Round}{None}{Touch}{5 + 1/Finesse Rounds}{Handful of water}{1d6 damage}{Any handful of water touched turns into a sparkling nugget of elemental ice in the caster's hand, which may then be thrown at any target (50 + 10/Finesse Feet) away in their LOS (roll "to strike") once each round for the duration of the spell and as long as there is at least a handful of water.}

\spellentry{Gills}{5}{1 Round}{4d6 vs WIL negates}{Touch}{2d20 + 1/Finesse Minutes}{1 creature}{Breathe water}{Enable the recipient to extract oxygen from water. It conveys no movement or other benefits to a submerged creature.}

\spellentry{Hot or Cold}{6}{2 Rounds}{None}{LOS 60 + 15/Finesse Feet}{12 + 3/Finesse Minutes}{4 + 2/Finesse Feet Radius}{Control water temp.}{Ccontrol the temperature of the indicated volume of any aqueous substance, making it boil or freeze, or any temperature between. The desired temperature is maintained for the duration, after which the temperature normalizes according to the ambient air temperature.}

\spellentry{Rain Dance}{7}{1 Minute}{None}{Caster}{1 + 0.25/Finesse Hours}{1 + 0.5/Finesse met Radius}{See text}{Increase or decrease the level of precipitation by up to 1/2 inch per hour. The form will be as rain or snow, depending upon the prevailing air temperature. The area is centered on the caster and with them.}

\spellentry{Hail}{8}{2 Rounds}{4d6 vs HEA for 1/2}{60 + 20/Finesse Feet}{5 + 1/Finesse Rounds}{10 + 5/Finesse Feet Radius}{2d6 damage}{A hail storm strikes the area.}

\spellentry{Dam}{9}{3 Rounds}{None}{LOS 100 + 30/Finesse Feet}{6 + 3/Finesse Minutes}{40 + 20/Finesse Feet Radius}{Stop flowing water}{Create an invisible wall of force which stops the flow of water in a stream or currents in a lake or sea. If the dam can touch land at both ends, it will hold back rising water. If land cannot be touched the water will flow around the Dam with increased force on either side. When the spell ends, any pent-up water is released.}

\spellentry{Riverman}{10}{4 Rounds}{4d6 vs PWR negates}{Multitouch}{1d4 + 1/Finesse Minutes}{Targeted creatures}{Walk on water}{Those under the spell can walk on water without sinking.}

\spellentry{Demolish}{11}{3 Rounds}{None}{LOS 20 Feet}{Instantaneous}{20 + 5/Finesse Feet Radius}{Destruction}{Cause latent moisture within the area of effect to suddenly freeze, totally collapsing and destroying most wooden structures. Living creatures are not affected.}

\spellentry{Ocean Cold}{12}{2 Rounds}{4d6 vs HEA for 1/2}{0}{5 + 1/Finesse Rounds}{80 x 10 Foot Wide cone}{3d6 damage}{Create a cone-shaped area of the chill of the deepest oceans extending from the caster outstretched hands and becoming 10 feet wide at the limit of the range. Creatures in the area of effect suffer 3d6 DP per round and haver their movement speed reduced by 1/2x due to the shock of the intense cold.}

\spellentry{Torrent}{13}{1 Minute}{None}{0}{4 + 1/Finesse Hours}{2 + 1/Finesse met Radius}{Torrential downpour}{Cause the sky to cloud over and unleash 2 inches of rain per hour. Such a massive downpour of rain may cause structures to be washed away or damaged (GM's discretion).}

\spellentry{Divert}{14}{10 Rounds}{None}{120 + 30/Finesse Feet}{See text}{100 + 50/Finesse Feet wide}{Change course}{Change the course of a stream, redirecting the flow in any direction, even uphill. The new channel is permanent, although directional changes (reversing normal flow direction, etc.) will only last one day.}

\spellentry{Akvovoki}{15}{4 Rounds}{None}{LOS 40 + 10/Finesse Feet}{4 + 1/Finesse Minutes}{Special}{Call water elementals}{Summon 1d6+1 water elementals to any point within the caster's range from any water source (even a cup of water is sufficient). The elementals will serve the caster, even in combat. They will not turn on them, but will make note of who summoned them. There is a 1d10 chance on each successive summoning that the elementials will attempt to take the caster with them at the end of the spell.}

\spellentry{Tidal Wave}{16}{2 Rounds}{Special}{LOS 240 + 60/Finesse Feet}{Instantaneous}{Special}{Create tidal wave}{Create a vast wall of moving water on a river, lake, or ocean within range. The wall height will range from 1d4 x 10 Feet for a river or lake to 1d20 x 10 Feet for the ocean. Once cast, the wave will strike in 1d12 rounds. All trees and structures are washjed away (GM's discretion on exceptionally strong stone structures). Living creatures must make either an RC of 5d6 vs Swimming or a 7d6 vs STR or drown.}

\section{Wardings}

%Wardings
\spellentry{Forfend}{1}{4 Rounds}{3d6 vs PWR negates}{Touch}{2d6 + 2/Finesse Hours}{See text}{Area protection}{The magician must touch at least three points during the casting time to set the ward. Any person or creature attempting to enter the Forfended area must make the RC to do so. The area could be a pouch, a door, an opening, etc.}
%Wardings
\spellentry{Fountain}{2}{1 Minute}{None}{Touch}{2d8 + 2/Finesse Hours}{5 + 3/Finesse Feet Radius}{Audible alarm}{Fountain may be cast anywhere, though it is most effective when cast on a dry fountain. When an intruder enters the protected area the fountain starts to flow with a gentle burbling noise that awakens the caster, even from a magical sleep.}
%Wardings
\spellentry{Poisonward}{3}{4 Rounds}{4d6 vs WIL negates}{Touch}{1d8 + 2/Finesse Rounds}{1 creature}{See text}{Poisonward reveals the presence of poisonous creatures or items in the target's LOS. If the effect is finessed four times, the target is allowed to subtract one die from their RC if hit by a poisonous attack during the spell's duration.}
%Wardings
\spellentry{Stay}{4}{4 Rounds}{3d6 vs PWR negates}{Touch}{2d6 + 2/Finesse Hours}{See text}{Prevent exit}{The magician must touch four corners of the area to be warded (the four cardinal points if a circular area) within the casting time. Anyone can enter the warded area, but cannot leave unless the RC is made.}
%Wardings
\spellentry{Weapon Ward}{5}{4 Rounds}{4d6 vs WIL negates}{Touch}{20 + 8/Finesse Minutes}{1 creature}{Reveal weapons}{The target of Weaponward is immediately forewarned of the location of weapons anywhere within 120 feet of their LOS. Anyone carrying weapons cannot approach within melee range of the target unless the RC is made.}
%Wardings
\spellentry{Snugbug}{6}{4 Rounds}{3d6 vs PWR negates}{Touch}{2d4 + 1/Finesse Hours}{See text}{Provide safe sleep}{The caster must touch at least four corners of the area to be warded during the casting time. Anyone within the area at the end of the casting time immediately falls into a restful sleep that can only be interrupted by inflicting damage. Snugbug also ward}
%Wardings
\spellentry{Monitor}{7}{4 Rounds}{None}{Touch}{4 + 1/Finesse Hours}{Caster}{Check wards}{Monitor enables the caster to remotely monitor the condition of any wards he has set, finding out if they have expired or if they have been tested or breached. The caster establishes Monitor by touching the perimeter of the warded area or object and casts}
%Wardings
\spellentry{Debar}{8}{4 Rounds}{4d6 vs PWR negates}{Touch}{1d6 + 1/Finesse Hours}{See text}{Bar 1 + 1/Finesse backgrnds}{Debar gives the caster control over who may enter the warded area, by background. The ward is set by touching at least four corners of the area during the casting time while stating which backgrounds are excluded. People of those backgrounds must make the}
%Wardings
\spellentry{Claw Ward}{9}{4 Rounds}{4d6 vs PWR negates}{Touch}{8 + 4/Finesse Minutes}{1 creature}{Shield natural weaponr}{Creatures with claws, fangs, or other natural weapons attacking the recipient of Claw Ward must make the RC before attempting the blow. If the RC is failed, no attack can be made by that creature that round. y}
%Wardings
\spellentry{Water Tongue}{10}{4 Rounds}{None}{Touch}{Special}{Up to 4 liters}{20 + 5/Finesse words}{Water Tongue enables the caster to endow a volume of water with the ability to deliver a short message (25 words or less) when a preset triggering event occurs. The trigger must be detailed by the caster at the time of casting. The spell remains on the ob}
%Wardings
\spellentry{Magic Ward}{11}{4 Rounds}{4d6 vs PWR negates}{Touch}{1d6 + 1/Finesse Hours}{See text}{Block spells}{The magician must touch at least four corners of the area to be warded during the casting time. A spell cast into the warded area must roll a number of d6 equal to its rank and score higher than the ward's 11d6 roll to be successful.}
%Wardings
\spellentry{Master Ward}{12}{7 Rounds}{5d6 vs PWR}{Touch}{1d8 + 2/Finesse Hours}{See text}{See text}{Master Ward enables the caster to set up any spell he can cast as a penalty for breaching a ward. The ward is set by touching the four corners of the area, casting Master Ward, and then the penalty spell (including any allowable finesses). The indicated Rounds}
%Wardings
\spellentry{Spirit Ward}{13}{4 Rounds}{5d6 vs PWR negates}{Touch}{1d8 + 2/Finesse Hours}{See text}{Ward against spirits}{Spirit Ward is set by the caster by touching at least four corners of the area during the casting time. Spirits cannot enter the warded area unless the RC is made.}
%Wardings
\spellentry{Bliss}{14}{4 Rounds}{4d6 vs PWR negates}{Touch}{Permanent}{See text}{Make items invisible}{The caster sets Bliss by touching at least four corners of the area during the casting time while designating 1 + 1/Finesse inanimate items in the area. At the end of the casting time those items become invisible, and can only be seen from within the area by ma}
%Wardings
\spellentry{Spongelungs}{15}{4 Rounds}{5d6 vs PWR negates}{Touch}{2d6 + 2/Finesse Hours}{See text}{Deadly ward}{Spongelungs is set as described for previous wards. If the RC is failed the creature's lungs fill with water, causing death in four rounds. If drowning is inappropriate for the victim, death may not result (GM discretion).}
%Wardings
\spellentry{Multicheck}{16}{1 Round}{None}{Touch}{See text}{See text}{1 + 1/Finesse RC}{Multicheck is cast immediately after another warding spell. It forces additional RC checks (up to 5) to be made to breach the ward. The duration and area of Multicheck matches the original ward.}

\section{Changings}

%Changings
\spellentry{Distill}{1}{1 Round}{See text}{Touch}{Instantaneous}{1 + 0.5/Finesse liters}{Change liquids}{Distill allows the caster to change any aqueous liquid into another aqueous liquid. The magical properties of the liquid (if any) are unaltered. Magical liquids get an RC of 9 or less on 3d6 to be affected. The caster needs to have tasted the created liqu}
%Changings
\spellentry{Drought}{2}{1 Round}{See text}{Touch}{Instantaneous}{1 + 0.5/Finesse Feet Radius}{Liquid to earth}{Drought changes the indicated amount of aqueous liquid into sand, clay, and/or stone. Magical properties (if any) of the material are not altered. Magical liquids get an RC of 9 or less on 3d6 to be affected.}
%Changings
\spellentry{Memorize Profile}{3}{10 Minutes}{None}{Touch}{Special}{1 creature}{Memorize profile}{This spell allows the caster to memorize the profile of a target creature, animal or person. The profile is a detailed view of the physical characteristics of the target and is used in other changings spells. The caster must be touching the target for the}
%Changings
\spellentry{Condense}{4}{2 Rounds}{None}{LOS 20 + 10/Finesse Feet}{Instantaneous}{5 + 2/Finesse Feet Radius}{Air to water}{Condense changes the indicated volume of air to water, the resultant volume decreasing to 10\% of its original size. The magical properties (if any) are not affected. Magical gasses get an RC of 9 or less on 3d6 to be affected. The GM will adjudicate any d}
%Changings
\spellentry{Dissolve}{5}{2 Rounds}{Special}{Touch}{Instantaneous}{1 + 0.25/Finesse Feet Radius}{Earth to water}{Dissolve changes any sand, clay and/or stone objects up to the volume indicated into water. The magical properties (if any) are unaltered. Magical objects (e.g.  magical gems, etc.) get an RC of 9 or less on 3d6 to be affected.}
%Changings
\spellentry{Dowse}{6}{2 Rounds}{Special}{LOS 10 + 3/Finesse Feet}{1d20 + 4/Finesse Rounds}{1 + 0.5/Finesse Feet Radius}{Fire to water}{Dowse changes any normal fire up to the diameter indicated into water. Magical fires get an RC of 9 or less on 3d6 to be affected.}
%Changings
\spellentry{Evaporate}{7}{1 Round}{Special}{Touch}{Instantaneous}{5 + 2/Finesse Feet Radius}{Liquid to air}{Evaporate changes the indicated amount of aqueous liquid into air. The volume increases by a factor of 10. The magical properties (if any) are unaffected. The GM will adjudicate any pressurization effects. Magical liquids get an RC of 9 or less on 3d6 to }
%Changings
\spellentry{Disguise}{8}{4 Rounds}{4d6 vs PWR negates}{Touch}{1d12 + 2/Finesse Hours}{See text}{Change appearance}{Disguise allows the caster to alter the appearance of one target, of the caster's own species, to roughly the same height and weight (within 10\%). Sex, hair and eye color, etc. may be freely changed.}
%Changings
\spellentry{Endwarf}{9}{4 Rounds}{4d6 vs PWR negates}{Touch}{1d8 + 1/Finesse Hours}{See text}{See text}{Endwarf enables the caster to shrink another of their own race, including equipment, down to 10\% of the target's original size, or 1\% if the effect is finessed once (a 6 foot tall being could be made 7" tall, or 7/10" with one finesse). The target's STR is }
%Changings
\spellentry{Transmute}{10}{4 Rounds}{4d6 vs PWR negates}{Touch}{1d12 + 2/Finesse Hours}{See text}{Change appearance}{Similar to Disguise, this spell allows the caster alter the race of the target, if desired, as well as the appearance. The height and weight limitations are the same, and the target must be of the caster's species.}
%Changings
\spellentry{Enmass}{11}{4 Rounds}{4d6 vs PWR negates}{Touch}{1d8 + 1/Finesse Hours}{See text}{Up to 300 + 100/Finesse\%}{Enmass is the opposite of Endwarf, and affects the target (who must be of the caster's species) and their equipment. If the target is originally 6' tall, he could be made up to 18' with the base spell, or up to 42' tall with four finesses. The target's STR }
%Changings
\spellentry{Elemental Mastery}{12}{3 Rounds}{None}{Touch}{1 day + 12 hours/Finesse}{1 + 0.5/Finesse Feet Radius}{Change any element}{Elemental Mastery enables the magician to convert the indicated volume of one element to another as he desires. Outlining the area to be converted with their hand, he casts the spell and the area changes to the indicated material(s) for the duration of the }
%Changings
\spellentry{Age}{13}{2 Rounds}{4d6 vs PWR negates}{LOS 20 + 5/Finesse Feet}{Permanent}{1 creature}{Age 1d20 + 4/Finesse years}{The target of Age who fails the RC, instantly and permanently becomes older.}
%Changings
\spellentry{Transmogrify}{14}{1 Hour}{None}{Touch}{Permament}{1 creature}{Modify other}{This spell allows the caster to make a physical change in another person or creature. The target of the spell must be forcibly immobilized for the entire casting time of the spell.  The caster uses a memorized profile of the physical attribute to guide th}
%Changings
\spellentry{Mutate}{15}{1 Hour}{None}{Self}{Instantaneous}{Self}{Modify caster/target}{Mutate empowers the caster to make physical changes to themselves (or another single immobilized target with one finess). Having memorized the profile of the features he wishes to acquire, he casts this spell. Guiding the changes to their own body, extra arms }
%Changings
\spellentry{Rebirth}{16}{1 Round}{Special}{Self}{Permanent}{Caster}{Totally transform}{The caster uses this spell to completely renew their physical form. Once started, the rebirth can not be aborted. For 2d6 rounds, in a shower of pyrotechnics, the caster changes. The player rerolls all the initial physical charactistics as if they were crea}

\section{Charms}

%Charms
\spellentry{Rockabye}{1}{1 Round}{3d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{Until awakened}{1 creature}{Put INT >3 asleep}{The intelligent target of this spell must not be in combat, and must be unaware that the caster is attempting a spell. If the spell succeeds, the target slips into a natural sleep, but will awaken normally. If the spell fails, the target is unaware of the}
%Charms
\spellentry{Drowse}{2}{1 Round}{3d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{Until awakened}{1 creature}{Put INT <3 asleep}{The effect of Drowse is the same as Rockabye, but the target must be non-intelligent.}
%Charms
\spellentry{Jackfrost}{3}{1 Round}{3d6 vs WIL negates}{20 + 5/Finesse Feet}{Instantaneous}{5 + 2/Finesse Feet Radius}{Inactivate plants}{Plants or plant-like creatures in the area of effect which fail the RC are blasted with a light ice, which melts away in one round. The plants then turn brown and lie flat on the ground, just as if frosted. The plants will return to their original conditi}
%Charms
\spellentry{Frostfeet}{4}{1 Round}{4d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{2d6 + 2/Finesse Rounds}{1 + 1/Finesse creatures}{Freeze in place}{The intelligent targets of this spell are prevented from moving, as if their feet were frozen to the ground.}
%Charms
\spellentry{Snare}{5}{1 Round}{4d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{2d6 + 2/Finesse Rounds}{1 + 1/Finesse creatures}{Halt INT <3 movement}{The non-intelligent targets of Snare which fail the RC feel incapable of movement, as if bound or caught in a snare, for the duration of the spell.}
%Charms
\spellentry{Shock}{6}{1 Round}{4d6 vs WIL negates}{LOS 20 + 5/Finesse Feet}{2d6 + 2/Finesse Rounds}{10 + 3/Finesse Feet Radius}{Bind plants}{If the RC is failed, plants in the targeted area are prevented from moving, as if bound with rope like shocks of grain in a field, for the spell's duration.}
%Charms
\spellentry{Witchy Sleep}{7}{2 Rounds}{4d6 vs WIL negates}{LOS 25 + 5/Finesse Feet}{1d6 + 1/Finesse Hours}{1 + 1/Finesse creatures}{Induce magical sleep}{If the intelligent targets of Witchy Sleep fail their RC, they slip into a deep, magical sleep. They will only awaken when the spell is cancelled by the caster, Revoked, or the duration expires.}
%Charms
\spellentry{Hibernate}{8}{2 Rounds}{4d6 vs WIL negates}{LOS 25 + 5/Finesse Feet}{1d6 + 1/Finesse Hours}{1 + 1/Finesse creatures}{Induce magical sleep}{This spell is similar to Witchy Sleep, except it works only on non-intelligent creatures.}
%Charms
\spellentry{Dormant}{9}{2 Rounds}{4d6 vs WIL negates}{25 + 5/Finesse Feet}{1d6 + 1/Finesse days}{10 + 3/Finesse Feet Radius}{Halt plant processes}{Dormant causes all plants in the targeted area which fail the RC to immediately suspend all normal plant functions such as growth, photosynthesis, fruit production, etc.}
%Charms
\spellentry{Siren}{10}{2 Rounds}{4d6 vs WIL negates}{LOS 25 + 5/Finesse Feet}{2d6 + 2/Finesse Rounds}{1 + 1/Finesse creatures}{Impose will on INT >3}{Siren enchants any intelligent creature to do the caster's bidding. If the target disobeys the caster, the spell is broken, but the target looses one point of WIL permanently. When the spell ceases, the target will be completely aware of the enchantment a}
%Charms
\spellentry{Drover}{11}{2 Rounds}{4d6 vs WIL negates}{LOS 25 + 5/Finesse Feet}{2d6 + 2/Finesse Rounds}{1 + 1/Finesse creatures}{Impose will on INT <3}{Drover is similar to Siren, except it only works on non-intelligent creatures. Other aspects of the spells are the same.}
%Charms
\spellentry{Greenthumb}{12}{2 Rounds}{4d6 vs WIL negates}{25 + 5/Finesse Feet}{2d6 + 2/Finesse Rounds}{10 + 3/Finesse Feet Radius}{Impose will on plants}{Greenthumb allows the caster to coerce plants in the targeted area, which fail the RC, to do their bidding. Such actions must be within the abilities of the plants, i.e., ground movement is impossible unless the plant has that ability. Accelerated growth, e}
%Charms
\spellentry{Nightmare}{13}{2 Rounds}{5d6 vs WIL negates}{Touch}{Permanent}{Single target}{Withdraws mind}{The intelligent target of this spell has their mind thrust into their subconscious, where he has non-ending and terrifying nightmares. His body seems vacant and vegetable-like. This spell stays in effect until cancelled by the caster or Revoked.}
%Charms
\spellentry{Davy Jones}{14}{4 Rounds}{5d6 vs WIL negates}{Touch}{Permanent}{1 creature}{Confinement}{The target must be within 20 feet of a 200 square foot or larger body of water for the spell to work. This powerful Water magic causes any flesh-and-blood creature to be drawn into the water. The creature does not die, and may be brought back by Revoking }
%Charms
\spellentry{Mesmerize}{15}{3 Rounds}{5d6 vs WIL negates}{LOS 5 + 1/Finesse Feet}{Permanent}{1 creature}{Hypnotize}{At the end of the casting time, scintillating water droplets stream from the caster's fingertips and weave a seductive pattern before the targeted (any flesh-and-blood) creature. If the RC is failed the creature is totally under the caster's control and c}
%Charms
\spellentry{Decant}{16}{2 Minutes}{5d6 vs WIL negates}{Touch}{Permanent}{Single target}{Move life force}{The caster can pour the immersed target's life force into a decanter filled with water, or to return it to a vacant body. If the RC is failed, the life force is transferred. No more than one life force can inhabit a body, unless all parties are willing. A}
